---
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: no
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    latex_engine: xelatex
    keep_tex: yes
    #template: svm-latex-ms.tex
# output: 
#   pdf_document:
#     citation_package: natbib
#     keep_tex: true
#     fig_caption: true
#     latex_engine: pdflatex
#     template: svm-latex-ms.tex
header-includes:
  # - \usepackage{setspace}\doublespace
  # - \usepackage{endnotes}
  #- \let\footnote=\endnote
  #- \usepackage{endfloat}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \setlength{\headheight}{13.6pt}
  - \rhead{\textit{Hoffmann and Velasco}}
editor_options: 
  chunk_output_type: console
  
citeproc: no
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
indent: yes
link-citations: yes
linkcolor: blue
bibliography: zotero.bib  
csl: apa.csl

title: "Supplementary Material"
subtitle: "Making Migration Sexy"
date: "`r format(Sys.time(), '%B %e, %Y')`"
# author:
# - name: Nathan I. Hoffmann
#   affiliation: Department of Sociology, University of California, Los Angeles
# - name: Kristopher Velasco
#   affiliation: Department of Sociology, University of Texas at Austin
 
author:
- Nathan I. Hoffmann, Department of Sociology, University of California, Los Angeles
- Kristopher Velasco, Department of Sociology, University of Texas at Austin

---

<!-- Turn off hyphenation -->
<!-- \usepackage[none]{hyphenat} -->


```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, results = 'asis',
                      cache.lazy = F)

library(janitor)
library(priceR)
library(kableExtra)
library(sjstats)
library(imputeTS)
library(mvtnorm)
library(ordinal)
library(lme4)
library(MASS)
library(countrycode)
library(knitr)
library(RColorBrewer)
library(haven)
library(cowplot)
library(srvyr)
library(lmtest)
library(sandwich)
library(here)
library(plm)
library(stargazer)
library(tidyverse)

# knitr::opts_chunk$set(error=TRUE) 

options("yaml.eval.expr" = TRUE)

# theme_set(theme_cowplot(font_family = 'Palatino') + 
theme_set(theme_classic(base_family = 'Palatino') + 
            theme(legend.title=element_blank(), 
                  panel.grid.major.y = element_line('grey80'),
                  legend.background = element_rect(fill = "transparent")))
ggplot <- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Dark2") + 
  scale_fill_brewer(palette="Dark2")

options(scipen=1, digits=2)
knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")
})
knitr::opts_chunk$set(error=TRUE) 


```

```{r load, include = F}
acs_wide <- read.csv(here('data', 'acs_wide.csv'))
acs_couple_policy <- readRDS(here('data', 'acs_couple_policy.rds'))
acs_prop_yrimmig_policy <- read.csv(here('data', 'acs_prop_yrimmig_policy.csv'))
acs_dyad_policy <- read.csv(here('data', 'acs_dyad_policy.csv'))

# acs_oneimm <- read.csv(here('data', 'acs_oneimm.csv'))
# acs_coupled_imms <- read.csv(here('data', 'acs_coupled_imms.csv'))
# acs_prop_yrimmig <- read.csv(here('data', 'acs_prop_yrimmig.csv'))
# acs_dyad <- read.csv(here('data', 'acs_dyad.csv'))




acs_wide_survey <- acs_wide %>% as_survey_design(ids = cluster, weights = hhwt)
acs_ind_survey <- acs_couple_policy %>% as_survey_design(ids = cluster, weights = perwt)


# Smaller dataset for initial ordered logit analyses
set.seed(1859)
acs_couple_policy_small <- bind_rows(
  filter(acs_couple_policy, same_sex == T),
  sample_n(filter(acs_couple_policy, same_sex == F), 1e5)
)


```

# Additional descriptive
```{r}
acs_couple_policy %>%
  filter(same_sex == T)%>%
  group_by(bpld) %>%
  count() %>%
  arrange(desc(n))

acs_couple_policy %>%
  filter(same_sex == F)%>%
  group_by(bpld) %>%
  count() %>%
  arrange(desc(n))
```

```{r}
hwsei <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(hwsei, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Hauser and Warren Socioeconomic Index')

nchild <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(nchild, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Number of Children')

years <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(years_in_us, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Years in U.S.')
distance <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(distw, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Distance to Country of Origin (KM)')

unemp <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(unemp_dif, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Difference in Unemployment Rates')
state_income <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(state_income, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'U.S. State of Residence Per Capita Income')


bind_rows(hwsei, nchild, years, distance, unemp, state_income) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = var, color = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = var_low, ymax = var_upp, fill = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var_name, scales = "free") +
  ylab('') +
  theme(legend.justification=c(1,0),
      legend.position=c(.22,.83))
```


# Alternate country proportion specifications
```{r country-props}
# acs_prop_yrimmig_policy <- acs_prop_yrimmig_policy %>%
#   filter(yrimmig >= 1992)

# ag <- lm(I(log(n_same_sex+1)) ~ origin_score + I(log(n_total+1)), 
#          data = acs_prop_yrimmig_policy) 
# 
# 
# ag_country_orig <- lm(I(log(n_same_sex+1))  ~ distw + contig + comlang_off + comlang_ethno + colony +
#                    wage_dif + unemp_dif + polity5 + origin_score + stock_prop+ I(log(n_total+1)), 
#          data = acs_prop_yrimmig_policy)
# 
# ag_country_orig_fe <- lm(I(log(n_same_sex+1))  ~ distw + contig + comlang_off + comlang_ethno + colony +
#                    wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld+ I(log(n_total+1)), 
#          data = acs_prop_yrimmig_policy)


# year of immigration fixed-effects
ag_country_time <- lm(prop_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld + yrimmig,
         data = acs_prop_yrimmig_policy)

# Add binary indicator for post-2013
ag_country_orig_2013 <- lm(prop_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop +
                     post_2013,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))

# Interaction, no fixed effects
ag_2013_int <- lm(prop_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5  + stock_prop + bpld +
                     post_2013*origin_score,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))

# Spouse, no fixed effects
ag_2013_spouse1 <- lm(prop_spouse_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + 
                     post_2013,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))

# Spouse, no interaction
ag_2013_spouse2 <- lm(prop_spouse_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld +
                     post_2013,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))

# Spouse + interaction
ag_2013_spouse_int <- lm(prop_spouse_same_sex ~ distw + contig + comlang_off + 
                           comlang_ethno + colony + bpld +
                   wage_dif + unemp_dif + polity5  + stock_prop +
                     post_2013*origin_score,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))


se_list <- list()
mod_list <- list(ag, ag_country_orig, ag_country_orig_fe, 
                 ag_country_orig_fe_2013, ag_2013_int_fe,
                 ag_2013_spouse, ag_2013_spouse_int)
for(mod_num in 1:length(mod_list)){
  se_list[[mod_num]] <- coef(summary(mod_list[[mod_num]], cluster = c("bpld")))[, 2]
}


stargazer(mod_list,
          header = F, 
          model.names = F,
          keep = c('origin_score', 'polity5', 'post_2013'),
          # omit = 'bpld',
          order = c('origin_score', 'polity5', 'post_2013', 'post_2013TRUE:origin_score'),
          # covariate.labels = c('Country LGBT policy score', 'Polity5 democratization',
          #                      'Post-2013', 'Country score × Post-2013'),
          dep.var.labels   = c("Percent in same-sex couples", "Married"),
          add.lines = list(c('Country controls?', 'no', rep('yes', 6)),
                           c('Country FEs?', 'no', 'no', rep('yes', 5))),
          se = se_list,
          keep.stat = c('n'),
          star.cutoffs = c(0.1, 0.05, 0.01, 0.001),
          star.char = c('†', '*', '**', '***'),
          notes = c("Note: †p<0.1; *p<0.05; **p<0.01; ***p<0.001", 
                    "Source: American Community Survey 2008-2019"), 
          notes.label = '',
          notes.align = 'l',
          notes.append = F,
          label = 'tab:country-props',
          title = 'OLS regressions of percent of immigrants in same-sex couples by year of immigration and country of origin. Country-clustered standard errors shown in parentheses. Country controls include population-weighted distance, contiguous border, common official language, common ethnic language, colonial relationship, wage differential, unemployment differential, proportion same-country stock, and Polity 5 measure of democratization.')

# One immigrant
ag_oneimm <- lm(prop_same_sex_oneimm ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld,
         data = acs_prop_yrimmig_policy)

ag_oneimm_2013 <- lm(prop_same_sex_oneimm ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld + post_2013*origin_score,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))
```

# Adjusting proportions based on empirical mismatch rates
```{r country-props-2}
# acs_prop_yrimmig_policy <- acs_prop_yrimmig_policy %>%
#   filter(yrimmig >= 1992)

# ag <- lm(I(log(n_same_sex+1)) ~ origin_score + I(log(n_total+1)), 
#          data = acs_prop_yrimmig_policy) 
# 
# 
# ag_country_orig <- lm(I(log(n_same_sex+1))  ~ distw + contig + comlang_off + comlang_ethno + colony +
#                    wage_dif + unemp_dif + polity5 + origin_score + stock_prop+ I(log(n_total+1)), 
#          data = acs_prop_yrimmig_policy)
# 
# ag_country_orig_fe <- lm(I(log(n_same_sex+1))  ~ distw + contig + comlang_off + comlang_ethno + colony +
#                    wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld+ I(log(n_total+1)), 
#          data = acs_prop_yrimmig_policy)



ag <- lm(prop_same_sex_adj ~ origin_score,
         data = acs_prop_yrimmig_policy)


ag_country_orig <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop,
         data = acs_prop_yrimmig_policy)

ag_country_orig_fe <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld,
         data = acs_prop_yrimmig_policy)

ag_oneimm <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld,
         data = acs_prop_yrimmig_policy)

# Add binary indicator for post-2013
# ag_country_orig_2013 <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
#                    wage_dif + unemp_dif + polity5 + origin_score + stock_prop +
#                      post_2013,
#          data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))

ag_country_orig_fe_2013 <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld +
                     post_2013,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))


ag_2013_int_fe <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5  + stock_prop + bpld +
                     post_2013*origin_score,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))


ag_2013_spouse_int <- lm(prop_spouse_same_sex_adj ~ distw + contig + comlang_off + 
                           comlang_ethno + colony + bpld +
                   wage_dif + unemp_dif + polity5  + stock_prop +
                     post_2013*origin_score,
         data = mutate(acs_prop_yrimmig_policy, post_2013 = (yrimmig > 2013)))

se_list <- list()
mod_list <- list(ag, ag_country_orig, ag_country_orig_fe,
                 ag_country_orig_fe_2013, ag_2013_int_fe, ag_2013_spouse_int)
for(mod_num in 1:length(mod_list)){
  se_list[[mod_num]] <- coef(summary(mod_list[[mod_num]], cluster = c("bpld")))[, 2]
}


stargazer(mod_list,
          header = F,
          model.names = F,
          keep = c('origin_score', 'polity5', 'post_2013'),
          order = paste0("^", c('origin_score', 'polity5', 'post_2013TRUE', 
                                'post_2013TRUE:origin_score') , "$"),
          covariate.labels = c('Country LGBT policy score', 'Polity5 democratization',
                                'Post-2013', 'Country score × Post-2013'),
          dep.var.labels   = c("Percent in same-sex couples", "Married"),
          add.lines = list(c('Country controls?', 'no', rep('yes', 5)),
                           c('Country FEs?', 'no', 'no', rep('yes', 4))),
          se = se_list,
          keep.stat = c('n'),
          star.cutoffs = c(0.1, 0.05, 0.01, 0.001),
          star.char = c('†', '*', '**', '***'),
          notes = c("Note: †p<0.1; *p<0.05; **p<0.01; ***p<0.001",
                    "Source: American Community Survey 2008-2019"),
          notes.label = '',
          notes.align = 'l',
          notes.append = F,
          label = 'tab:country-props',
          title = 'ADJUSTED: OLS regressions of percent of immigrants in same-sex couples by year of immigration and country of origin. Country-clustered standard errors shown in parentheses. Country controls include population-weighted distance, contiguous border, common official language, common ethnic language, colonial relationship, wage differential, unemployment differential, proportion same-country stock, and Polity 5 measure of democratization.')


# mod_list <- list(ag, ag_country_orig, ag_country_orig_fe, 
#                  ag_country_orig_fe_2013, ag_2013_int_fe, ag_2013_spouse_int)
# for(mod_num in 1:length(mod_list)){
#   mod_list[[mod_num]] <- coef(summary(mod_list[[mod_num]], cluster = c("bpld")))
# }
# 
# keep_coefs <- c('Country LGBT policy score' = 'origin_score', 
#                 'Polity5 democratization' = 'polity5', 
#                 'Post-2013' = 'post_2013TRUE', 
#                 'Country score × Post-2013' = 'post_2013TRUE:origin_score')
#    
# 
# export_summs(mod_list,
#        coefs = keep_coefs,
#        statistics = c('N' = 'nobs'),
#        stars = c(`†`= 0.1, `*`= 0.05, `**` = 0.01, `***` = 0.001),
#        note = '{stars}. 
#                Source: American Community Survey 2008-2019') %>%
#   set_caption('OLS regressions of percent of immigrants in same-sex couples by year of immigration and country of origin. Country-clustered standard errors shown in parentheses. Country controls include population-weighted distance, contiguous border, common official language, common ethnic language, colonial relationship, wage differential, unemployment differential, proportion same-country stock, and Polity 5 measure of democratization.') %>%
#   add_rows(matrix(c('Country controls?', 'no', rep('yes', 5),
#                 'Country FEs?', 'no', 'no', rep('yes', 4)),
#            byrow = T, nrow = 2) , 
#            after = nrow(.) - 1)

```

# Accounting for mismatch

```{r mismatch}
tibble(`Study Year` = c(2010, 2010, 2016, 2016),
           Relationship = c('Married', 'Unmarried Partner', 'Married', 'Unmarried Partner'),
           `Mail` = c('59%', '7%', '47.4%', '5.6%'), 
           `Internet` = c('NA', 'NA', '22.5%', '2.4%'), 
           `CAPI/CATI` = c('46%', '13%', 'Unknown', 'Unknown'))
```


```{r respmode}
tabyl(acs_couple_policy, respmode, year, same_sex) %>%
  adorn_percentages('col')
  

tabyl(filter(acs_couple_policy, same_sex == T), respmode, year) %>%
  adorn_percentages('col')
```


In a Census Bureau working paper, @kreider_2015 use personal information such as names and addresses match same-sex couples from the 2010 ACS to Social Security administrative data. They find that 58 percent of married couples coded as same-sex in the ACS are coded as different-sex in the administrative data. The corresponding sex mismatch rate for unmarried same-sex couples is 7 percent. (Our data include `r sum(acs_wide$related_partner == 'Spouse' & acs_wide$same_sex == T)` married and `r sum(acs_wide$related_partner == 'Unmarried Partner' & acs_wide$same_sex == T)` unmarried same-sex immigrant couples.) A follow-up study [@kreider_2017] shows that these mismatch rates appear to have fallen: In a 2016 ACS test module that included explicit categories for different- and same-sex spouses and partners, 31 percent of married and 3 percent of unmarried same-sex couples had inconsistent sex responses. This decreasing mismatch rate may be due to the greater numbers of same-sex couples openly identifying themselves as well as the growing popularity of responding to the ACS via Internet (see Supplementary Material), a response mode introduced in 2013 which is now the default [@u.s.censusbureau_2017]. In the 2016 test of the ACS, the mismatch rate for mail-in responses was 47 and 6 percent for married and unmarried same-sex couples, respectively, whereas for Internet responses they were only 22 and 2.4 percent [@kreider_2017].^[An interview-based CATI/CAPI response mode is sometimes administered as well (to `r with(filter(acs_couple_policy, same_sex == T), round(sum(respmode == 'CATI/CAPI')/nrow(filter(acs_couple_policy, same_sex == T)),2)*100)` percent of our sample of immigrants in same-sex couples), but the 2016 study did not assess its error rate. In the 2010 ACS, Kreider and Lofquist (2015) find CATI/CAPI sex reporting mismatch for 46 and 13 percent for married and unmarried same-sex couples, respectively.] In our sample of immigrants in same-sex couples, `r sum(acs_couple_policy$respmode == 'Mail' & acs_couple_policy$same_sex == T)` responded by mail and `r sum(acs_couple_policy$respmode == 'Internet' & acs_couple_policy$same_sex == T)` responded by Internet survey.   

Despite these difficulties, many researchers have relied on ACS data to draw demographic conclusions about same-sex couples in the U.S., usually without making accommodations for measurement error . In several ways, we test the sensitivity of our results, especially to the possibility that a large proportion of apparently same-sex couples misreported a partner's sex. However, such adjustments assume that sex response error occurs randomly throughout the immigrant population. Beginning in 2019, the ACS added explicit "opposite-sex" and "same-sex" categories to the relationship questions, so future work on same-sex couples in the ACS will be more reliable [@goodnature_2021].

## Sensitivity Analysis
We first assess how much our results would change in the presence of significant measurement error. @kreider_2015 found that in the 2010 ACS, 57 percent of married and 7 percent of unmarried same-sex couples may have been different-sex couples that misreported a partner's sex. By 2016 the mismatch rate had declined to 31 percent of married and 3 percent of unmarried same-sex couples [@kreider_2017]. 

Figure \ref{fig:sens} takes Model 3 form Table \ref{tab:country-props} and reduces the proportions of same-sex couples in the data. It varies the percentage of misreported same-sex married couples from 0 to 90 percent and of unmarried couples from 0 to 30 percent. Highlighted in blue bars are the empirical mismatch rates found in the two studies by Kreider et al. [-@kreider_2015; -@kreider_2017]. We see that extremely high rates of misreporting are necessary to reduce this effect to 0. 

```{r sens, eval = F, fig.cap = 'Coefficient for sending-country LGBT policy context, adjusted for possible misreporting of same-sex couples. Ribbon shows 95 percent confidence intervals and blue bars show estimated misreporting from the 2010 and 2016 U.S. Census Bureau tests on the ACS.'}

acs_prop_yrimmig_policy <- acs_prop_yrimmig_policy %>%
  mutate(prop_same_sex_adj = (prop_spouse_same_sex*(1-.573) + prop_partner_same_sex*(1-.07)))

lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld,
         data = acs_prop_yrimmig_policy) %>%
  summary()
#.35 .034


sens_df <- list()
mis_df <- data.frame(spouse = sort(c(.573, .35, seq(from = 0, to = .9, length.out = 8))),
           partner = sort(c(.07, .034, seq(from = 0, to = .3, length.out = 8))))

for(i in 1:nrow(mis_df)){
  sens_mod <- lm(prop_same_sex_adj ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + polity5 + origin_score + stock_prop + bpld,
         data = acs_prop_yrimmig_policy %>%
           mutate(prop_same_sex_adj = (prop_spouse_same_sex*(1-mis_df[i,1]) + 
                                         prop_partner_same_sex*(1-mis_df[i,2])))
         )
  
  coef_os <- coef(summary(sens_mod, cluster = c("bpld")))['origin_score',]
  sens_df[[i]] <- data.frame(est = coef_os[[1]],
                             se = coef_os[[2]],
                             mis_spouse = mis_df[i,1],
                             mis_partner = mis_df[i,2])
}
sens_df <- bind_rows(sens_df) %>%
  mutate(lower = est - 1.96*se, 
         upper = est + 1.96*se,
         highlight = mis_spouse %in% c(.573, .35))

ggplot(sens_df, aes(x = (mis_spouse+mis_partner)/2, y = est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .2)  +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_errorbar(aes(ymin = lower, ymax = upper, width = .02, color = highlight)) +
  theme(legend.position = "none") +
  scale_color_manual(values = c('transparent', RColorBrewer::brewer.pal(3, 'Dark2')[3])) +
  labs(y = 'Coefficient for LGBT policy context')
```

We next text the robustness of our results to omitted variable bias.

# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent


