---
output:
  # bookdown::pdf_document2:
  bookdown::word_document2:
    reference_docx: "word-template.docx"
    toc: no
    number_sections: no
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    latex_engine: xelatex
    keep_tex: no
    keep_md: true
    #template: svm-latex-ms.tex
# output: 
#   pdf_document:
#     citation_package: natbib
#     keep_tex: true
#     fig_caption: true
#     latex_engine: pdflatex
#     template: svm-latex-ms.tex
header-includes:
  # - \usepackage{setspace}\doublespace
  # - \usepackage{endnotes}
  #- \let\footnote=\endnote
  #- \usepackage{endfloat}
  # - \usepackage{fancyhdr}
  # - \pagestyle{fancy}
  # - \setlength{\headheight}{13.6pt}
  # - \rhead{\textit{Hoffmann and Velasco}}
  # - \lhead{\textit{`r format(Sys.time(), '%B %e, %Y')`}}
editor_options: 
  
  chunk_output_type: console
citeproc: no
# fontfamily: mathpazo
mainfont: "Times New Roman"
fontsize: 12pt
geometry: margin=1in
indent: yes
link-citations: no
linkcolor: black
bibliography: zotero2.bib  
csl: apa.csl
title: "Sexuality, Migration, and LGB Policy: A Portrait of Immigrants in Same-Sex Couples in the United States"
# title: "How Immigrants in Same-Sex Couples Change our Understanding of Migration"
# title: "Repressed or Empowered? Immigrants in Same-Sex Couples in the U.S."
# title: "Making Migration Sexy: How LGB Policies Influence International Migration"
# date: "`r format(Sys.time(), '%B %e, %Y')`"
# author:
# - name: Nathan I. Hoffmann
#   affiliation: Department of Sociology, University of California, Los Angeles
# - name: Kristopher Velasco
#   affiliation: Department of Sociology, Princeton University
 
# author:
abstract: "Both internationally and in the U.S., the policy landscape for same-sex couples is changing rapidly, and surveys report swiftly increasing numbers of immigrants in same-sex couples in the U.S. Yet few researchers have examined immigrant-containing lesbian, gay, and bisexual (LGB) couples on a large scale, especially regarding their relationship to LGB policy. Are these immigrants disadvantaged and fleeing anti-LGB contexts, or empowered to migrate by progressive origin-country LGB policy? Using the American Community Survey from 2008 to 2019 and original datasets indexing LGB policy changes in 121 countries and all U.S. states, this study assesses and characterizes the scale of LGB migration to the U.S. as well as the role of LGB policy. Compared to those in different-sex couples, immigrants in same-sex couples come from richer, more democratic countries that are less represented among immigrants in the U.S. They also tend to be more highly educated, work in more prestigious occupations, and have higher incomes. Contrary to previous work focusing on LGB immigrants from repressive contexts, fixed-effect models show that higher proportions of these immigrants come from LGB-friendly countries, and they are more likely to live in progressive U.S. states. These findings highlight how sexuality as well as state policies seemingly unrelated to migration can shape migratory pathways."
  
# keywords: "migration, same-sex couples, LGB policy, sexuality"   
# thanks: "The authors contributed equally to this paper. We thank Phillip Ayoub, Filiz Garip, Margaret Peters, Roger Waldinger, the Migration Working Group at UCLA, and participants at the Annual Meetings of the American Sociological Association and the Population Association of America for their helpful feedback."
---

<!-- Turn off hyphenation -->
<!-- \usepackage[none]{hyphenat} -->


```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, # results = 'asis',
                      cache.lazy = F, dpi = 600)


library(MASS, exclude = 'select')
library(janitor)
library(priceR)
library(kableExtra)
library(huxtable)
library(jtools)
library(sjstats)
library(imputeTS)
library(mvtnorm)
library(ordinal)
library(lme4)
library(countrycode)
library(knitr)
library(RColorBrewer)
library(haven)
library(cowplot)
library(srvyr)
library(lmtest)
library(sandwich)
library(here)
library(plm)
library(stargazer)
library(patchwork)
library(rworldmap)
require(maps)
library(flextable)
library(tidyverse)

# knitr::opts_chunk$set(error=TRUE) 

options("yaml.eval.expr" = TRUE)

uclablue = '#2774AE'
gray = '#808080'
black = '#000000'
uclagold = '#FFB81C'
ucla_palette = c(uclablue, black, gray)

# theme_set(theme_cowplot(font_family = 'Palatino') + 
theme_set(theme_classic(base_family = 'Palatino') + 
            theme(legend.title=element_blank(), 
                  panel.grid.major.y = element_line('grey80'),
                  legend.background = element_rect(fill = "transparent")))
ggplot <- function(...) ggplot2::ggplot(...) + scale_color_manual(values = ucla_palette) +
  scale_fill_manual(values = ucla_palette)
# ggplot <- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Dark2") +
#   scale_fill_brewer(palette="Dark2")

# library(colorspace)
# colorspace_palette = 'RdPu'
# ggplot <- function(...) ggplot2::ggplot(...) +
#   scale_color_discrete_sequential(palette=colorspace_palette) +
#   scale_fill_discrete_sequential(palette=colorspace_palette)

options(scipen=1, digits=2)
knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")
})
knitr::opts_chunk$set(error=TRUE) 

set_flextable_defaults(font.family = 'Times New Roman',
                       padding.bottom = 1,
                       padding.top = 1,
                       padding.left = 3,
                       padding.right = 1,
                       theme_fun = 'theme_booktabs',
                       digits = 3, 
                       big.mark = ',',
                       line_spacing = 1)
```



```{r load, include = F}
acs_wide <- read.csv(here('data', 'acs_wide.csv'))
acs_ind <- readRDS(here('data', 'acs_ind.rds')) 
acs_prop <- read.csv(here('data', 'acs_prop.csv')) %>%
  mutate(post_2013 = (yrimmig > 2013)) 
acs_prop_state <- read.csv(here('data', 'acs_prop_state.csv')) %>%
  mutate(post_2013 = (year > 2013)) 



# acs_oneimm <- read.csv(here('data', 'acs_oneimm.csv'))
# acs_coupled_imms <- read.csv(here('data', 'acs_coupled_imms.csv'))
# acs_prop_yrimmig <- read.csv(here('data', 'acs_prop_yrimmig.csv'))
# acs_dyad <- read.csv(here('data', 'acs_dyad.csv'))


acs_wide_survey <- acs_wide %>% as_survey_design(ids = cluster, weights = hhwt)
acs_ind_survey <- acs_ind %>% as_survey_design(ids = cluster, weights = perwt)

# Smaller dataset for initial ordered logit analyses
set.seed(1859)
acs_ind_small <- acs_ind
# acs_ind_small <- bind_rows(
#   filter(acs_ind, same_sex == T, yrimmig >= 1991),
#   sample_n(filter(acs_ind, same_sex == F), 1e5)
# ) 

total_df <- acs_wide_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, imm_couple, same_sex) %>%
  summarize(n = survey_total(vartype = "ci")) 
```




```{r intro}
same_2013 <- total_df %>%
  filter(year == 2013, same_sex == 'Same-Sex') %>%
  pull(n) %>% 
  sum()

same_2019 <- total_df %>%
  filter(year == 2019, same_sex == 'Same-Sex') %>%
  pull(n) %>% 
  sum()

dif_2013 <- total_df %>%
  filter(year == 2013, same_sex == 'Different-Sex') %>%
  pull(n) %>% 
  sum()

dif_2019 <- total_df %>%
  filter(year == 2019, same_sex == 'Different-Sex') %>%
  pull(n) %>% 
  sum()
```


# Introduction
In 2013, the U.S. Supreme Court overturned the Defense of Marriage Act and required the U.S. government to recognize marriages between same-sex spouses. Among many consequences, this decision radically changed the immigration landscape: For the first time, same-sex spouses of U.S. citizens and lawful permanent residents were eligible to file a spousal petition for an immigrant visa [@edwards_2013]. Since, the U.S. population of immigrants in same-sex couples has grown rapidly [@redpath_2022_spousal; @ruggles_2021]. According to American Community Survey data, numbers of different-sex couples including immigrants increased by `r round((dif_2019/dif_2013 - 1)*100)` percent from 2013 to 2019 (from `r round(dif_2013 / 1e6,1)` million to `r round(dif_2019 / 1e6,1)` million), while those of corresponding same-sex couples grew from `r round(same_2013 /1e3, 1)` thousand to `r round(same_2019 /1e3, 1)` thousand in the same period, an increase of `r round((same_2019/same_2013 - 1)*100)` percent. While some cursory descriptions exist [@gates_2013; @goldberg_2021], there is now a pressing need to understand the contours of this burgeoning population and the relevant forces shaping their migratory pathways.

While immigrants in same-sex couples may resemble their heterosexual co-nationals and respond to similar forces (e.g., economic benefits and network effects), there is good reason to suspect distinction. As the Supreme Court decision indicates, this population may be particularly responsive to changes in state policy. Indeed, the Supreme Court decision occurred against a backdrop of rapidly changing laws concerning same-sex couples -- and lesbian, gay, and bisexual (LGB) communities, generally -- both in the U.S. and abroad. As some countries expanded rights and social recognition, others imposed new forms of repression [@hadler_2018_world]. Emerging qualitative work demonstrates that changing LGB policy landscapes are salient factors influencing migration decisions to the U.S., especially for asylees seeking refuge from more repressive contexts [@ahmad_2013; @carrillo_2018; @mai_2009; @gorman-murray_2009; @vogler_2016] but also for elite queer migrants [@choi_2022_global; @difeliciantonio_2016_situating]. However, there has yet to be a large-scale, representative investigation into how changing policy environments influence the migration patterns of individuals in same-sex couples into and across the U.S. Through studying the migration of same-sex couples, then, this study allows us to gain a fuller understanding of migratory pathways by underscoring the importance of political and "lifestyle" considerations within migration decisions [@benson_2012; @fitzgerald_2014].  

How does the U.S. population of immigrants in same-sex couples differ from their different-sex counterparts? And how does country-of-origin and U.S. state LGB policy relate to their relative numbers? To address our research questions, 
<!-- we evaluate how LGB policies at country of origin and U.S. state of residence relate to the migratory patterns for immigrants cohabiting with a same-sex partner. We do so by  -->
we integrate two types of data. First, we rely on American Community Survey (ACS) data from 2008 to 2019 [@ruggles_2021], which allows the identification of individuals in cohabiting same-sex couples as well as their country of origin, U.S. state of residence, education, income, and other individual characteristics. Second, we merge these data with country- and state-level datasets on variables relevant to these migrants, most notably an original dataset indexing LGB policy changes in `r length(unique(acs_prop$bpldid))` countries and all U.S. states from 1991 to 2019.  

Our analytic strategy proceeds in two parts. First, we descriptively understand who these cohabiting LGB immigrants are. This first step is important because little is known about this growing population. Second, we focus on country-of-origin and U.S.-state effects, modeling how representation of immigrants in same-sex partnerships changes over time in relation to the LGB policy in both contexts. We control for numbers of immigrants in different-sex couples by modeling the proportion of immigrants from a given country that are in cohabiting same-sex couples, out of all cohabiting immigrants from that country.

<!-- Our analytic strategy proceeds in two parts. First, we descriptively understand who these cohabiting LGB immigrants are. This first step is important because little is known about this growing population. Second, we use fixed effects analyses at the country and U.S. state levels to adjudicate between two competing theories of queer migration. Research on queer asylum suggests repressive policies are more likely to spur migration to the U.S. [@luibheid_2008; @vogler_2016]. However, another set of scholarship, extending from research on "lifestyle migration," suggests affirming policy environments may serve as a migratory launchpad [@benson_2012].       -->
<!-- Theories of migration often focus on economic and networked dynamics to explain migratory patterns. To the extent social policies are factored into the equation, previous scholarship largely centers policies governing immigration itself. Though understandable, inattention to other social policies obscures how the state can structure migration by conditioning qualities of life that motivates one to leave their country-of-origin. One such important quality of life is sexuality. Indeed, sexuality and how the state governs it is a key motivator for leaving one's home country.   -->

Our investigation offers several novel insights. First, we find that immigrants in same-sex couples differ markedly compared to co-nationals in different-sex relationships. Specifically, those in same-sex relationships are often more highly educated, have higher incomes, and come from countries with relatively few immigrants in the U.S. Second, we find that origin countries with more LGB-friendly policies send higher proportions of immigrants in same-sex couples into the U.S. 
<!-- In line with the aspirations-capabilities framework [@dehaas_2021], affirming policies appear to enable the migration projects of LGB individuals and provide the necessary material and symbolic resources, whereas oppressive policies hinder such ambitions.  -->
This finding is unexpected given existing queer migration scholarship which largely focuses on asylum seekers leaving repressive contexts and seeking entry into the U.S. and other countries in the Global North [@murray_2014_real; @sam_2015_teacher; @akin_2017_queer; @dhoest_2019_learning; @giametta_2020_new; @saleh_2020_queer; @karimi_2020]. Yet, the DOMA decision appears to have opened an alternative pathway for sexual migration -- a pathway being utilized by relatively privileged migrants from progressive contexts at much greater scales. Our findings also indicate that immigrants in same-sex couples are more likely to reside in U.S. states with progressive policies, especially if they come from countries with more supportive policies as well. By showing how policies seemingly unrelated to migration can shape migration decisions and actions, this projects shows how identity -- and the state's governance of it -- can interact with broader institutional contexts to yield unexpected results.


<!-- This is an important intervention as analyses into how sexuality motivates migratory decisions or how migration reimagines sexual behaviors and understandings are limited [@carrillo_2018]. As such,  -->
<!-- This project contributes to emerging migration scholarship by incorporating an aspirations-capabilities framework into traditional models of migration. As such, we highlight the important interaction between sexuality and political context for shaping migration decisions. Additionally, our research responds to calls to consider factors in both the origin and destination societies concurrently when studying migration [@fitzgerald_2008; @luthra_2018] and underscores the importance of incorporating political and "lifestyle" variables into typical understandings of migration [@benson_2012; @fitzgerald_2014].  -->


# Background: Changing Policy Landscapes and Same-Sex Immigrant Couples

The U.S. continues to undergo significant shifts in the policies governing LGB populations at both state and federal levels. Since 2003, the U.S. Supreme Court has ruled sodomy laws and the Defense of Marriage Act unconstitutional, federally recognized same-sex marriages, and curtailed employment discrimination. In response, though, several U.S. states have implemented new policies hindering LGB communities on top of existing discriminatory practices [@kazyak_2018]. These dynamics create a varied landscape in which state lines significantly demarcate the types of rights and legal environments LGB people experience. Now, a burgeoning area of scholarship exists to understand the causes of these transformations [@lax_2009; @soule_2004] and their distinct consequences on the lives and well-being of LGB people [@boertien_2019; @carpenter_2020; @kail_2015; @levy_2017; @cantu_2009].  

Although this changing policy landscape affects LGB populations of all types, particular sub-groups within this broad umbrella are differentially impacted. Same-sex immigrant couples represent a population especially vulnerable to recent changes. This is because prior to being able to experience recognized rights like marriage or non-discrimination protections, immigrants in same-sex couples must first be able to enter into the U.S. While single queer migrants could potentially enter the U.S. through some existing visa pathways (e.g., employment-sponsored, family-sponsored, asylum), federal U.S. law historically hindered LGB couples’ ability to enter the country due to the government’s lack of recognition of their relationship [@humanrightswatch_2006]. While, in theory, queer asylum seekers could enter with a same-sex spouse through the "derivative asylee status," lack of spousal recognition by sending countries and invasive requirements to “prove” one's sexuality rendered this avenue effectively inaccessible [@humanrightswatch_2006; @ritholtz_2021_queer].  

The federal environment governing immigration significantly changed after 2013. The U.S. Supreme Court decision ruling the Defense of Marriage Act (DOMA) unconstitutional opened the door for same-sex immigrant couples to enter the U.S. through the same process long governing different-sex couples [@edwards_2013]. Now, couples could enter together, or one partner already in the U.S. could sponsor their same-sex fiancé(e) or spouse. As Figure \@ref(fig:total-pop) highlights, the number of same-sex immigrant couples in the U.S. grew significantly following this ruling -- especially when compared to different-sex couples. Aside from allowing gay and lesbian families to remain unified, this national opening creates an important moment for the scholarly community to begin investigating the factors motivating this emerging population to migrate to the U.S.  

<!-- Broad legal exclusions of queer migrants, and especially those in existing relationships, often carry over to academic scholarship as well. Citizenship and migration research usually presents migrants as heterosexual [@luibheid_2008], and studies that do acknowledge the realities of queer migrants are overwhelmingly qualitative and often center either asylum seekers or the subjective consequences of migration for one's sexuality [@fournier_2018_lesbian; @vogler_2016]. Such studies provide important insights into the interactions between law, sexuality, and migratory experiences, especially for some of the world's most marginalized. One consequence of this focus, however, is that queer migration research presently offers a narrow understanding of the broader population of who queer migrants are and the factors influencing their migratory patterns, especially for those in same-sex relationships. This omission is especially stark considering that family-related migration tends to overshadow work-related and other categories overall [@kandel_2018_familybased]. Another restriction on understanding queer migrants is the dearth of high-quality, representative statistics within the U.S. -- making interview-based studies a necessity in order to gain any understanding at all. Therefore, the present study seeks to add a necessary contribution to scholarship on queer migration by tapping into underutilized data to provide representative insights on this growing segment of the queer migrant population. -->

<!-- Finally, it is important to situate the scope of this study. Namely, we focus on migrants in same-sex relationships in light of the changing legal environment due to DOMA decision. Consequently, we do not thoroughly investigate how policy environments influence single queer migrants. Because scholarship on queer migration is limited, however, we pull on existing studies of all queer migrants, including those who are single, and of non-migrant same-sex couples to ground our theorization and motivate our hypotheses below. In the Discussion, we outline potential implications of our findings for single migrants and discuss how their incorporation within future analyses may alter our present contribution. -->

<!-- Even broad, descriptive understandings of migrants in same-sex relationships within the U.S. are absent within academic literature.  -->

<!-- This leaves a significant blindspot. -->

**[FIGURE \@ref(fig:total-pop) ABOUT HERE]**

The U.S. is not alone in seeing rapid policy change related to LGB couples. Recent changes in LGB rights around the world are increasing the salience of one's sexuality as an important decision point to leave one's home country [@mole_2018a; @murray_2016]. This is partly driven by the transnational flow of information and general visibility of those with diverse sexualities that accompanies policy change [@ayoub_2016; @ayoub_2017]. Greater circulation of cultural content makes one cognizant of how they are treated in their home country due to their sexuality, for better or worse, compared to these outside contexts [@carrillo_2018; @karimi_2020; @vuckovicjuros_2022_lesbian]. Indeed, international organizations contribute to this increased salience by constructing sexuality as a legitimate basis for leaving. For example, in 2008 the United Nations High Commissioner for Refugees issued a new guidance note for why and how countries should consider sexual orientation and gender identity when granting asylum claims [@unhcr_2008]. The note guides various authorities to consider discriminatory domestic policies when evaluating asylum claims as such policies “can create or contribute to an oppressive atmosphere of intolerance and generate a threat of prosecution” [@unhcr_2008, p. 8].  International organizations such as the European Union and several countries now incorporate the U.N. guidance, though not the U.S. [@giametta_2020]. As the DOMA decision demonstrates, too, these LGB policies are important as they can condition initial access to a country and the quality of life once there.

Current research on the types of policy environments likely to influence the migration of those in same-sex couples is both limited in scope and mixed in outcomes. This is partly due to emerging scholarship at the intersection of sexuality and migration being overwhelmingly qualitative. Only recently have representative samples that allow identification of LGB immigrants been available, yet these have thus far been underutilized. Thus a broader portrait of LGB migrants, and how policy environments influence them, is urgently needed. 
<!-- Below, we  summarize competing hypotheses for how country-of-origin policies shape migration before getting to the U.S. state context. -->

# Empirical Expectations

We combine insights from three longstanding theories of migration with newer scholarship on queer migration to help inform our expectations for migrants in same-sex couples. We begin by describing these theories here in brief but then extend our discussions of each below. First, neoclassical economic theory posits that potential migrants engage in cost-benefit analyses when deciding whether and where to migrate [@sjaastad_1962_costs; @borjas_1989_economic]. Second, the New Economics of Labor Migration (NELM) conceives of migration as a family-level decision; even when the individual migrant may not see a wage gain, migration can help the family by offsetting local economic uncertainty [@stark_1985; @taylor_1999]. And, third, cumulative causation theory provides additional reasons why migration may occur in the absence of a wage difference [@massey_1990_social]. Once a migration stream begins, immigrant networks share information and resources to lower the cost of migration and settling in the destination country [@massey_1987], and institutions arise to ease entry and settlement [@hernandez-leon_2013]. Hence the first "pioneer" migrants are often more positively selected than later migrants, who benefit from these lower barriers to migrating [@hatton_2005a]. If immigrants in same-sex couples operate similarly to those in different-sex couples, then we should see similarly processes at play. Yet, it is clear that LGB-specific policy may introduce a distinct element that can either disrupt or, potentially, attenuate these patterns.  

<!-- Previous demographic research suggests how these theories might be relevant. In the labor market, men in same-sex couples tend to earn less than their heterosexual counterparts, but women tend to earn more [@klawitter_2015] -- though the gap for the latter has decreased in recent years [@badgett_2021_lgbtq]. On average, married same-sex couples are older and have higher earnings than their different-sex counterparts [@fisher_2018], while results for unmarried couples are mixed [@baumle_2009; @badgett_2021_lgbtq]. If immigrants in same-sex couples similarly have higher earnings than heterosexual immigrants, than their relative costs of migration may be lower. These higher earnings may open pathways for lifestyle factors to influence their migration decisions [@benson_2009; @dixon_2020; @choi_2022_global]. In addition, although same-sex romantic relationships between men generally have higher levels of instability than between two women or different-sex relationships [@joyner_2017], the dissolution rates for cohabiting same- and different-sex couples are similar [@manning_2016]. If LGB immigrants experience higher rates of union dissolution, then our estimates of the proportion of immigrants who are in same-sex couples represents a lower bound.   -->

With these theories in mind, are LGB immigrants in same-sex couples coming from generally repressive or progressive contexts? Though representative research on this population is lacking, existing evidence from qualitative studies document cases supporting both hypotheses. Much queer migration research supports the former perspective, reflected in Adur's suggestion that [-@adur_2018, p. 321, emphasis theirs], “LGBTI immigrants relocate in pursuit of spaces that they *imagine* will be safer and more liberal.” Indeed, numerous studies characterize migrants in same-sex relationships as largely escaping repressive contexts [@murray_2014_real; @sam_2015_teacher; @akin_2017_queer; @dhoest_2019_learning; @giametta_2020_new; @saleh_2020_queer]. One reason for this prominent characterization of LGB migrants within the U.S. is due to the heavy emphasis on asylum seekers and the asylum process itself -- especially since the U.N. guidance note in 2008 and then-President Obama's move to make queer refugees a "population of concern" for the U.S. in 2011 [@lewis_2014; @luibheid_2008; @vogler_2016]. Although the U.S. is less progressive and inviting compared to many other Western states, high-profile developments such as marriage equality can contribute to an imagined openness relative to many locations around the world. For example, access to gay content in film and on the Internet contributed toward Iranian refugees' decision to seek sexual freedom in the West [@karimi_2020]. Additional research documents how people in comparatively oppressive contexts seek out partners in more equitable locations, like the U.S., who can then sponsor them through the immigration process [@corey-boulet_2019; @carrillo_2018; @kong_2010_chinese]. This pathway fits within the neoclassical economic theory as LGB individuals seeking asylum into the U.S. or wanting to leave such conditions, generally, often face economic difficulties and may expect great returns upon migrating to progressive destinations [@luibheid_2008]. Moreover, under NELM, LGB individuals can potentially position themselves as a family's "insurance-policy" migrant, strategically aligning their desire for sexual freedom with their family's economic security.  

Although the studies described above support the pathway of immigrants in same-sex couples are largely fleeing repressive contexts, Luibhéid [-@luibheid_2005, p. xxv] cautions against overemphasizing queer migration as "a narrative of movement from repression to freedom, or a heroic journey undertaken in search of liberation" [see also @choi_2022_global; @waitt_2011_it]. Even when migrants imagine they are moving to a much more progressive context, actual policy differences may be relatively minor. @carrillo_2018 explores a prominent case of this in his study of the "sexual migration" of gay Mexican men to the U.S. Overall, U.S. LGB policy is somewhat more progressive than Mexico's, but Mexico's large, cosmopolitan cities are more welcoming to queer people than many parts of the U.S. Why, then, do these young men feel like they must cross the northern border in order to live their authentic lives? Carrillo argues that their often rural birthplaces, conservative upbringings, and unsympathetic family members lead these men to believe that only by migrating to the U.S. can they live openly. Since northward migration is a common choice for young Mexicans seeking upward mobility, family members are less likely to question this decision, and steady remittances help ward off prying questions. This observation may generalize to other contexts: LGB immigrants may believe they are moving to a much more progressive country, when actual policy differences are relatively minor.  

<!-- Some recent, small-scale studies also suggest that economic resources can enable the migration of elite gay migrants [@choi_2022_global], and that  -->

On the other hand, immigrants in same-sex couples may come from countries with greater recognition and access to sexuality-related rights and services than the U.S. There are three interrelated reasons for this. First, affirming policy environments are likely to enable people’s capacity to make such an important, expensive move [@dehaas_2021]. Long-standing research on immigrant selection demonstrates that migrants typically occupy stronger social positions -- more formal education, higher incomes, and more prestigious occupations [@feliciano_2020]. Supportive policies like marriage equality and protections against employment discrimination may enable the capacity to migrate by providing more access to necessary social, human, and economic capitals. Second, same-sex coupledom, like marriage, is a culturally contingent artifact [@philpot_2016_gay]. Consequently, the interactive dynamics between policy and culture within an immigrant’s country of origin is likely to influence both their decision to "come out" and then their desire to be part of a couple [@baiocco_2014_desire; @flores_2016_backlash; @ocobock_2020_leveraging; @suen_2021_sexual]. Policies supportive of LGB communities normalize and validate the disclosure of such identities and partnerships [@ocobock_2020_leveraging] -- influencing those with same-sex attractions to imagine and aspire such possibilities for themselves. And, relatedly, being from a country where the state recognizes one’s sexuality and validates these relationships may make survey respondents, once in the U.S., more comfortable disclosing their relationship. And, third, these factors may combine to make those coming from countries with progressive policies more legible to migration officials determining access to visas as the enactment of the sexuality and relationship may align more closely to dominant expressions within the U.S. [@edwards_2013].   

Migration from progressive policy contexts can also support traditional migration theories as well. The cost-benefit analysis inherent in neoclassical economic theory could imply that migrants coming from progressive contexts enjoy greater resources and hence face relatively lower costs to migration. Under NELM, countries that recognize same-sex relationships are more likely to have same-sex couples that make the migration decision together. Finally, since spousal migration for same-sex couples is a new pathway for migration, cumulative causation theory suggests that barriers to migrate will be high and hence "pioneer" migrants will be positively selected.  

<!-- Migration scholars have long accepted that migration is more likely when economic gain is likely [@hatton_2005a; @todaro_1980], established migration flows exist [@massey_1999; @massey_1987; @hernandez-leon_2013], and countries are physically and culturally proximate [@karemera_2000; @mayda_2010]. Despite a lack of research applying these theories to migration of same-sex couples, previous demographic research on non-immigrant same-sex couples suggests how they might apply. In the labor market, men in same-sex couples tend to earn less than their heterosexual counterparts, but women tend to earn more [@klawitter_2015] -- though the gap for the latter has decreased in recent years [@badgett_2021_lgbtq]. On average, married same-sex couples are older and have higher earnings than their different-sex counterparts [@fisher_2018], while results for unmarried couples are mixed [@baumle_2009; @badgett_2021_lgbtq]. If immigrants in same-sex couples similarly have higher earnings than heterosexual immigrants, then the typical economic push and pull factors of immigration may affect them differently. As discussed below, one difference is that these higher earnings may open pathways for lifestyle factors to influence their migration decisions [@benson_2009; @dixon_2020]. Same-sex couples also tend to be less homogamous in age, education, and race/ethnicity than their different-sex counterparts [@schwartz_2009]. Finally, although same-sex romantic relationships between men generally have higher levels of instability than between two women or different-sex relationships [@joyner_2017], the dissolution rates for cohabiting same- and different-sex couples are similar [@manning_2016]. If LGB immigrants experience higher rates of union dissolution, then our estimates of the proportion of immigrants who are in same-sex couples represents a lower bound. -->

<!-- On the other hand, factors specific to LGB couples may outweigh factors conventionally included in migration models. Work on the relationship between welfare policies and immigration suggests that social policies not explicitly related to the latter can impact the migration and settlement processes. Historically, welfare and immigration were tightly linked in the U.S. [@fox_2012], and welfare generosity may play a role in shaping choice of destination for prospective emigrants today [@razin_2015; @ferwerda_2021_pull]. Our research builds on these insights to test how non-welfare social policy specific to LGB individuals may influence their migration and settlement.  -->

<!-- Our analysis compares conventional explanations for migration to political ones related to LGB policy. Massey et al. [-@massey_1999, p. 50] provided an influential synthesis of migration theories from sociology, economics, and anthropology, arguing that "causal processes relevant to international migration might operate on multiple levels simultaneously." Our investigation incorporates insights from their work as well as assesses how sexuality and LGB policy complicate migration research. At one level, neoclassical economic theories underscore that promise of material gain is a frequent motivation to migrate [@hatton_2005a; @todaro_1980], predicting that migrants will follow wage and unemployment differentials across countries. To account for migration costs, these models often adjust for distance: Immigrants are more likely to migrate between proximate, especially bordering, countries. At another level, migration streams often continue in a process of cumulative causation even after wage differentials decrease [@massey_1990_social]; this is due in large part to immigrant networks that share information and resources to lower the cost of migration and settling in the destination country [@massey_1987] as well as institutions that arise to ease entry and settlement [@hernandez-leon_2013].  -->
<!-- @garip_2016 argues that different theories of migration are more relevant for different types of migrants; as the most prevalent migrant types shift, theories arise to better explain them.  -->
<!-- The theories synthesized by @massey_1999 remained largely materialistic.  Our study heeds Garip's [-@garip_2016] call to explore migrant heterogeneity, assessing what the theories of migration synthesized by @massey_1999 might miss when accounting for the migration of this LGB subpopulation. Recent research has shown that not only physical but cultural proximity can matter to the migration process; scholars have shown that shared language, colonial history, and democracy matter in the sending country [@karemera_2000; @mayda_2010], and immigrant rights matter in the receiving country [@fitzgerald_2014]. Work on the relationship between welfare policies and immigration suggests that social policies not explicitly related to the latter can impact the migration and settlement processes. Historically, welfare and immigration were tightly linked in the U.S. [@fox_2012], and welfare generosity may play a role in shaping choice of destination for prospective emigrants today [@razin_2015; @ferwerda_2021_pull]. Our research builds on these insights to test how non-welfare social policy specific to LGB individuals may influence their migration and settlement. -->

<!-- Previous scholarship has not explicitly applied these analyses and insights to the migration of same-sex couples on a large scale in the U.S. context, but  -->

 <!-- Overall, sexuality may interact with processes of migration in ways not anticipated by the conventional migration theories. We seek to intervene within migration scholarship by explicitly considering how sexuality, and the governance of it, expand our comprehension of migration atop this previous work. -->

<!--  @garip_2016 argues that different theories of migration are more relevant for different types of migrants; as the most prevalent migrant types shift, theories arise to better explain them.  --> 

<!-- Whereas @massey_1999 sought to integrate migration theories at different causal levels, more recent sociological work seeks to uncover heterogeneity between different types of migrants [@garip_2012; @drouhot_2021_what].Our study builds on this work by probing migrant heterogeneity, assessing what the theories of migration synthesized by @massey_1999 might miss when accounting for the migration of this LGB subpopulation.

<!-- We harness an aspirations-capabilities framework [@dehaas_2021; @schewel_2020; @carling_2018_aspiration], which allows consideration of the conventional migration theories discussed above as well as additional factors more relevant to LGB individuals.  -->

<!-- ## Sexuality in an Aspirations-Capabilities Framework
We incorporate the conventional migration theories discussed above into a broader aspirations-capabilities framework [@dehaas_2021; @schewel_2020; @carling_2018_aspiration]. De Haas [-@dehaas_2021, p. 17] defines migration as “a function of aspirations and capabilities to migrate within given sets of *perceived* geographical opportunity structures” (emphasis added). This framework blends individual, agentic motivations for migration (e.g., subjective desires and imaginations of well-being) with structural factors that condition these aspirations and the capabilities to act upon them (e.g., policy). Thus, we use this framework to situation our intervention: to demonstrate why sexuality, and the state’s governance of it, meaningfully contributes toward understandings of migration [@carrillo_2018; @cantu_2009; @suen_2021_sexual]. --> 

<!--The state plays an important role setting structural conditions that influence migratory pathways. One direct route is by establishing law governing migration itself. Indeed, significant scholarship details how such laws and policies are used to target distinct populations -- whether based on labor market needs [@hahamovitch_2014; @castles_2006], assistance following local disaster [@bekaert_2020; @hunter_2015], or familial ties [@kofman_2004_family]. However, @dehaas_2021 offers a more expansive view of the state. He contends that the state influences migration through policy that shapes individual’s life aspirations and well-being -- whether real or imagined -- and by conditioning the capacity to migrate based on Berlin's [-@berlin_1969_four] negative and positive “liberties." For example, explicit migration policy represents a “negative liberty,” whereby the state externally imposes opportunities and constraints, but liberal telecommunications policy that allows for the transnational flow of news, media, and information is a “positive liberty” that influences the capacity to migrate -- in this case by increasing knowledge of alternatives and the “capacity to aspire” [@dehaas_2021]. Therefore, the aspirations-capabilities framework is not a rejection of previous theories of migration, but a more expansive framework that acknowledges the subjective role of individuals to determine if migration advances their well-being within given conditions. -->

<!-- Sexuality, and the state’s governance of it, squarely fits within this framework. Sexuality is a salient quality of well-being and, as demonstrated by the DOMA decision in the U.S., conditions the capacity to migrate. Moreover, recent global expansions of LGB rights allows us to leverage varied policy landscapes and investigate beyond migration-oriented policy to theorize the role of the state within these processes. -->

<!-- ## Sexuality and Migration -->
<!-- Sexuality has long factored into migratory decisions [@carrillo_2018; @cantu_2009; @suen_2021_sexual]. Typically, however, there is an assumed heterosexuality when researchers investigate partner selection, family formation, lifestyle, etc. within migration [@luibheid_2008]. If those in same-sex relationships have similar profiles and descriptive characteristics as their heterosexual counterparts, then previous theories are likely to maintain. There are important qualitative distinctions between these populations, though, like differential access to rights and recognition through LGB policies. -->

<!-- Although this provides rich insights into the motivations and consequences of migration, differing effects of policies emerge. Research on queer asylum, one key line of research within this area, suggests repressive policies are more likely to spur migration to the U.S. However, another set of scholarship, extending from research on "lifestyle migration," suggests affirming policy environments may serve as a migratory launchpad.  -->

<!-- Prior to developing our theorization, we clarify our target population. Given the DOMA ruling's direct influence on couples as well as data limitations, we do not consider individuals who do not cohabit with a romantic partner. We control for this by considering only couples, whether same- or different-sex. Four types of couples are included in this population: (1) two-immigrant couples who came to the U.S. together; (2) two-immigrant couples that formed once in the U.S.; (3) mixed status couples where an immigrant migrated to be with their U.S.-born partner; and (4) mixed status couples that formed in the U.S. We theorize general mechanisms that influence all composition types but, below, acknowledge the potential or differential effects depending on couple composition. We elaborate on the implications of these scope conditions in the Discussion. -->

<!-- As such, immigrants from countries without this cultural and political background may be more reticent to desire or disclose same-sex partnership.  -->

<!-- Indeed, these arguments follow from scholarship on lifestyle migration, which @dehaas_2021 integrates into the aspirations-capabilities framework. -->

<!-- Scholarship on lifestyle migration supports these arguments. Benson and O'Reilly [-@benson_2009, p. 608] refer to lifestyle migration as the "relocation of relatively affluent people within the developed world searching for a better way of life." Lifestyle migration is conceptualized as a highly individualized decision-making process as conceptualizations of "better way of life" differ drastically [@benson_2016]. Supportive LGB policies may offer a structural opening by which same-sex couples have the opportunity and bandwidth to consider these individualistic choices. -->

<!-- Of course, as mentioned, there is little previous research on how same-sex immigrant couples fare socioeconomically. However, if these migrants mirror their high-earning U.S.-based counterparts [@badgett_2021_lgbtq], their better standing will likely enable lifestyle migration processes. -->


<!-- ## How Policies at U.S. State of Residence Influence Migration -->

<!-- While these country-of-origin policies may enable the movement of same-sex couples out of their home country, the varied policy environments across U.S. states are likely to differentially encourage the entry of these couples.  -->
 
<!-- Though research often focuses on either country of origin or destination, part of our intervention is to study these in tandem to provide a more comprehensive understanding of migration [@fitzgerald_2008; @luthra_2018].  -->
<!-- Indeed, especially coming from an aspirations-capabilities framework where subjective perceptions are meaningful, this dual investigation is necessary to understand relative contrasts.   -->
<!-- Typically, a strong predictor of where migrants locate within the U.S. is networks, bringing them to the areas where their family, friends, and other social relations live [@massey_1987; @portes_1998; @palloni_2001].  -->
Unlike our mixed expectations for origin-country policy, we are more confident that LGB immigrants will locate to U.S. states with progressive policy contexts. Gay and lesbian couples within the U.S. were likely to leave states without marriage equality prior to national recognition [@beaudin_2017], and queer migrants often have strong cross-national networks for relaying information [@stella_2020]. These factors likely results in a greater concentration of coupled LGB immigrants in states with marriage equality and other protective policies. Additionally, if migrants are coming from a country with greater legal protections, they are unlikely to want to relocate to a state where such rights are no longer recognized -- rendering the political environment acutely important. Of course, this is predicated on the assumption that migrants take such distinct sub-national variations into account -- which they very well may not. Consequently, the "pull" to individual states may operate independently from specific state laws affirming LGB people and their relationships.  

<!-- Though limited, existing demographic research does give some insights into how immigrants in same-sex couples might choose their state of residence once in the U.S. Cohabiting same-sex couples already within the U.S. are more likely to reside in states in the Northeast and West, such as Vermont, Massachusetts, California, and Oregon [@gates_2013a], that have often been at the forefront in safeguarding LGB rights. The same-sex population is growing most rapidly, however, in the Midwest and South (*ibid.*). Regardless of state, same-sex couples are more concentrated in urban areas, although this is truer for men than women [@baumle_2009]. This evidence suggests that if immigrants in same-sex couples are like their U.S.-born counterparts, they are likely to choose progressive states and cities as their place of residence. -->

<!-- In sum, it is evident that sexuality and the policies governing it are salient factors driving migratory decisions -- either enabling the opportunity and flexibility for same-sex couples to make decisions that are best for them or by erecting an environment so repressive that it forces migrants to flee to where imagination of opportunity awaits. Despite qualitative examinations into queer migrants, especially asylum seekers, there is no large-$N$ investigation into how significant transformation of LGB policies -- both globally and across U.S. states -- influence migration into the U.S. Therefore, this research seeks to fill this gap in the literature by providing such an analysis and to further understand how the changing policy landscapes are differentially influencing the lives of queer people depending on their social positions. -->


<!-- Prior to getting to the data and results of this analysis, it is first important to contextualize the present article within existing demographic scholarship.  -->


<!-- Note from Kris: I'm wondering if we should actually incorporate the "labor market" paragraph into the "Conventional Explanations" section? Put in there why we think conventional explanations apply/differ to same-sex couples? And then,the first paragraph could kinda go into theorizing why state policies may matter for same-sex couples but not different sex? -->

# Data and Methods
## Identifying Same-Sex Couples in the ACS
We merge individual-level data on immigrants in the U.S. with state- and country-level variables from a variety of datasets. The individual data come from the 2008 to 2019 ACS [@ruggles_2021]. Each year, the ACS surveys a 1-percent representative sample of the U.S. population about their education, occupation, income, family structure, immigration status, country of origin, location, and a variety of other individual and household attributes. We define a same-sex couple as two individuals of the same sex in the same household who report their relationship as "spouse" or "unmarried partner." We limit the sample to individuals who immigrated at the age of 18 or older and in 1991 or later.^[The exception is for Figure \@ref(fig:total-pop), where we include those who immigrated in any year, at age 18 or older.] Hence this analysis considers four types of couples: (1) two-immigrant couples who came to the U.S. together; (2) two-immigrant couples that formed once in the U.S.; (3) mixed status couples where an immigrant migrated to be with their U.S.-born partner; and (4) mixed status couples that formed in the U.S. In the Online Appendix, we present results separately for one- and two-immigrant couples. We are unable to differentiate between couples that partnered or married abroad and those that did so in the U.S. We elaborate on the implications of these scope conditions in the Discussion.  

The 12 years of survey data contain `r nrow(filter(acs_wide, same_sex == T & (yrimmig_main >= 1991 | yrimmig_partner >= 1991)))` same-sex couples that include at least one immigrant, for a total of `r nrow(filter(acs_ind, same_sex == T))` immigrants in same-sex couples with complete data. These immigrants are compared to `r nrow(filter (acs_wide, same_sex == F & (yrimmig_main >= 1991 | yrimmig_partner >= 1991)))` corresponding different-sex couples containing `r nrow(filter(acs_ind, same_sex == F))` individual immigrants.   

We use "LGB" to refer to all individuals who may be in romantic relationships with members of the same sex. We recognize that we are not able to identify bisexual (or pansexual, multisexual, etc.) individuals with different-sex partners. In addition, measuring the prevalence of same-sex couples in the U.S. is difficult [@michaels_2013]. As in most nationally representative demographic work on same-sex couples [@baumle_2013; @baumle_2019], we are able to identify only LGB couples that cohabit; unpartnered LGB individuals and those who do not live with their partner are not included in the analysis [@baumle_2009, p. 6]. In addition, LGB individuals who do not feel comfortable with the partner labels of the ACS are not in the sample. Another pitfall is measurement error: Misreporting may result when different-sex couples accidentally misspecify the gender of one of the partners [@gates_2009; @goodnature_2021]. Beginning in 2008 the Census Bureau made changes to ACS gender and partnership questions in order to prevent such errors [@u.s.censusbureau_2013], so we rely on data only from 2008 onward, but difficulties remain. If even a small number of different-sex couples misreport one partner's sex, the counts of same-sex couples will be inflated. Following @gates_2009, we remove all respondents that had either their relationship or sex variable allocated by the Census Bureau, which results in dropping `r 7683 - sum(acs_ind$same_sex == T)` immigrants in same-sex couples and `r 949775 - sum(acs_ind$same_sex == F)` in different-sex couples, or `r round(1-nrow(acs_ind) / 957458, 4) * 100` percent of the sample. This is the strategy used by most studies of same-sex couples in the ACS [e.g. @boertien_2019; @gates_2013; @goldberg_2021; @christafore_2019; @martell_2020]. In the Online Appendix we also include robustness checks to test the sensitivity of our results to high rates of misreporting.  


## Analytic Strategy 
Many of our analyses entail descriptive statistics of ACS data. For these and for the reshaping described below, we apply survey weights from the ACS. 

One of our goals is to isolate the effects of country-of-origin LGB policy on the immigration of immigrants in same-sex couples. The ideal survey would follow potential immigrants over time and have information about sexual orientation, allowing us to estimate how the probability of migrating and choice of U.S. state of residence vary by sexual orientation. This ideal dataset does not exist, but we attempt to approximate it.  

We reshape the data so that each observation is the proportion of individual cohabiting immigrants in same-sex couples from country $x$ in state $y$ in survey year $z$, then multiply by 100. For example, the observation for France $\times$ New York $\times$ 2015 is `r acs_prop_state %>% filter(state == 'New York', country == 'France', year == 2015) %>% pull(prop_same_sex)`, meaning that among coupled ACS respondents who immigrated from France and resided in New York at the time of the 2015 survey, a weighted `r acs_prop_state %>% filter(state == 'New York', country == 'France', year == 2015) %>% pull(prop_same_sex)` percent were in same-sex couples. This transformation of the data results in `r nrow(acs_prop_state)` country-state-year observations.^[Cells without any cohabiting immigrants from a given country are necessarily dropped from the analysis.] This reshaping acts to control for aspects of migrant settlement common to immigrants from the same country. We regress the country-state-survey year proportion on state and sending-country policy scores and adjust for state and origin-country controls (listed below) using OLS. State-level measures are linked to survey year, while origin-country LGB policy and controls are linked to the mean year of immigration for each country-state-year immigrant group. In our final models, we include state and country-of-origin fixed effects, and we cluster errors at the state and country levels.    



## Variables
Our country- and state-level variables come from a variety of sources. Country-of-origin variables include bilateral distance, contiguous border, common official language, common ethnic language, and whether the country was a former colony of the U.S., come from CEPII's GeoDist dataset [@mayer_2011]. Difference in living standards, calculated as difference in per capita GDP at purchasing power parity (country of origin minus U.S.), come from the RGDP$^\text{e}$ variable of the Penn World Table [@feenstra_2015], and we rely on World Bank data for differences in unemployment rates [@worldbank_2020]. We use Varieties of Democracy (V-Dem) measures of liberal democracy of the country of origin [@coppedge_2021]. In an attempt to proxy network effects, we create a new variable by dividing each country's immigrant stock by the total number of immigrants in the U.S. in a given year, using the UN's Trends in International Migrant Stock report for [@unitednations_2017] for 1990, 1995, 2000, 2005, 2010, 2015, and 2017, linearly interpolating to yield an annual time series from 1990 to 2020. For state controls, we use per capita income by year from the Bureau of Economic Analysis [@bea_2020] and state-level annual unemployment rates from the Bureau of Labor Statistics [@bls_2020]. All monetary variables are adjusted to 1999 U.S. dollars.  

<!-- We use Polity5 measures of democracy of the country of origin [@marshall_2018], with linear interpolation for missing years.  -->

<!-- the World Bank's Global Bilateral Migration Database [@ozden_2011] for 1960, 1970, and 1980 and -->



To examine LGB policies at country of origin and state of destination, we use original datasets. To create the U.S. state policy index, we compile data from the Movement Advancement Project^[https://www.lgbtmap.org/], a leading LGB organization in the U.S. that collects data on a number of relevant policies. Our state index encompasses both progressive policies (full marriage equality, state recognition of civil unions and domestic partnerships, ban on all employment and housing discrimination based on sexual orientation, hate crime protections based on sexual orientation, legal joint adoption by same-sex couples, and a ban on conversation therapy for minors) and regressive policies (criminalization of sodomy, state constitutional bans of marriage equality, religious freedom exemptions to discriminate against same-sex couples in adoption, and state-level bans on local non-discrimination ordinances encompassing sexual orientation). The state index ranges from `r min(acs_ind$state_policy, na.rm = T)` to `r max(acs_ind$state_policy, na.rm = T)`, and the mean state policy score for immigrants in our sample is `r mean(acs_ind$state_policy, na.rm = T)`.

We measure the origin country policy environment using a modified LGBT Policy Index [@velasco_2018]. The index comprises 16 policies, many similar to those above, but including additional policies like the death penalty for homosexual acts, propaganda laws limiting free speech for LGB communities, and equal age of consent between same-sex and different-sex couples. We remove policies targeting gender identity and transgender communities from the original index. Both state and country indices are created by summing the net total of progressive policies (scored $+1$) over regressive policies (scored $-1$). For the `r length(unique(acs_prop$bpldid))` countries of origin for our sample, country index ranges from `r min(acs_ind$origin_score, na.rm = T)` to `r max(acs_ind$origin_score, na.rm = T)`, and the mean score of country of origin for immigrants in our sample is `r mean(acs_ind$origin_score, na.rm = T)`. Immigrants are assigned U.S. state index scores based on their state of residence as reported in the ACS, and they are assigned country-of-origin index scores based on their birthplace and year of immigration.  

We also include a binary variable to indicate the change in national policy environment within the U.S. following the overturning of DOMA in 2013. This represents an important shift opening up traditional pathways of immigration to same-sex couples. We include an interaction term between this post-DOMA indicator and country-of-origin index because we theorize that the effects will be more pronounced following this period.

For individual characteristics, we rely on ACS variables for reported sex, age, education (with categories for less than high school, high school, some college, and college), year of immigration, inverse hyperbolic sign of positive income in thousands in the past year [similar to the natural log transformation; see @burbidge_1988_alternative], and a binary unemployment indicator (for income reported to be 0 or less). Descriptive statistics for variables used in all analyses are included in Section A of the Online Appendix. 


# Results
## Descriptive Trends
We first estimate total numbers of immigrants in same- and different-sex couples, applying survey weights to obtain population-level estimates from the ACS. Recall that Figure \@ref(fig:total-pop) showed that, whereas numbers of different-sex immigrant couples have steadily increased over the period of study, numbers of same-sex immigrant couples have increased much more rapidly, especially since the the 2013 Supreme Court decision overturning DOMA.  

**[FIGURE \@ref(fig:desc-country) ABOUT HERE]**  


How do same- and different-sex immigrant couples differ in their origin-country and individual attributes? Do variables typically used in models for migration differ between the groups? Figure \@ref(fig:desc-country) compares immigrants in same- and different-sex couples on five origin country-level variables. First, microeconomic theory predicts that differences in wages and living standards across countries is one of the most important motivations for migration. Panel A in Figure \@ref(fig:desc-country) shows that per-capita GDP is indeed higher in the U.S. than the average country of origin for both groups of immigrants, but the gap is significantly greater for immigrants in different-sex couples. This means that immigrants in same-sex couples are coming from countries with higher standards of living than those in different-sex couples. Statistics for the unemployment rate differential (Panel B) indicate similar trends: compared to the countries of origin for immigrants in different-sex, immigrants in same-sex couples come from countries with lower unemployment rates. These findings indicate that microeconomic considerations may be less important to the migration of LGB immigrants.  

Models of migration also often use distance between countries as a proxy for migration costs, with closer countries often having more migration between. Panel C of Figure \@ref(fig:desc-country) shows that immigrants in same-sex couples tend to come from somewhat more distant countries. Panel D looks at a measure of network effects: At the time of immigration, what is the proportion of total immigrants in the U.S. from the country of origin? Compared to different-sex couples, immigrants in same-sex couples immigrated from countries that were less represented in the U.S. population at the time of migration. This indicates that the network effects that attract migrants from the same country of origin may be less relevant to LGB immigrants, or that these migration streams are relatively new. Finally, Panel E of Figure \@ref(fig:desc-country) compares V-Dem democracy level for country of origin at time of migration. We see that levels of liberal democracy tend to be higher for immigrants in same-sex couples, indicating that political context may play a more important or different role in their migration decisions.  

**[FIGURE \@ref(fig:desc-ind) ABOUT HERE]**  

Immigrants in same-sex couples appear to come from more advantaged countries, but are they more advantaged on an individual level as well? Previous studies on LGB immigrants show those coming from advantaged countries may still be fleeing individual disadvantage. Figure \@ref(fig:desc-ind) presents results from four individual attributes. Panels A and B show that immigrants in same-sex couples tend to be *more* highly educated than immigrants in different-sex couples, with about 10 percent more having high school and four-year college degrees. Panel C corroborates the macroeconomic findings at the individual level: Not only do immigrants in same-sex couples come from countries with higher per capita GDP, but they individually tend to earn more than immigrants in different-sex couples. Panel D demonstrates that immigrants in same-sex couples also tend to work in professions with higher occupational prestige scores. Together, these findings indicate that LGB immigrants may come from more privileged social origins than their heterosexual counterparts.  

Figure \@ref(fig:desc-country) implies that there is between-country selection of immigrants in same-sex couples, but Figure \@ref(fig:desc-ind) does not necessarily imply within-country selection as well. Section B of the Online Appendix examines within-country selection explicitly, subtracting the individual attributes in Figure \@ref(fig:desc-ind) from country-year averages in the ACS. The steadily positive values for immigrants in same-sex couples mean that not only do immigrants in same-sex couples come from more advantaged countries, but they tend to be among the most advantaged immigrants from a given country, as well.  

**[FIGURE \@ref(fig:policy-desc) ABOUT HERE]**  

Although we see significant differences between same- and different-sex couples on a number of important migration variables, none shows the sudden jump in recent years reflected in Figure \@ref(fig:total-pop). Turning to LGB policy may better explain this surge. Figure \@ref(fig:policy-desc) charts the average country-of-origin and U.S.-state LGB policy score for the immigrants in our sample over time, comparing means for immigrants in same- and different-sex couples. The left panel shows that country-of-origin index at time of migration is generally higher for immigrants in same-sex couples, and since 2013 it has rapidly increased. Immigrants in same-sex couples tend to come from more progressive countries, and this trend tracks closely with the overall population of this group. The right panel indicates less of a difference in U.S. state policies, although states where immigrants in same-sex couples live tend to score somewhat higher.  

 

Figure \@ref(fig:country-map) shows the percentage of immigrants to the U.S. in same-sex couples from each country of origin, averaging over the 12 years of survey data, and Table \@ref(tab:country-tab) presents the top ten of these along with average LGB policy score over these years. The top sending countries include interesting diversity. Although countries with more progressive policies top the list, Malaysia, Zimbabwe, and Singapore make the list with their relatively repressive contexts. Having countries within the top 10 span multiple regions and cultures provides preliminary evidence that LGB policy is not substantially affecting willingness to respond truthfully on the ACS about being in a same-sex couple. Nor does it appear as though responses to the ACS are simply a function of country-of-origin LGB policies, as policy scores vary significantly across the top 10. Figure \@ref(fig:state-map) presents the percentage of immigrants in same-sex couples in U.S. states, averaging over the survey years and possible countries of origin, and Table \@ref(tab:state-tab) ranks the top ten. Although states with progressive policies occupy the top spots, Mississippi, Missouri, and Florida still make the list with less affirming policy environments. (See appendix for full list of country and state rankings).


**[FIGURE \@ref(fig:country-map) ABOUT HERE]**  

**[FIGURE \@ref(fig:state-map) ABOUT HERE]** 
  
**[TABLE \@ref(tab:country-tab) ABOUT HERE]**  
  
  
**[TABLE \@ref(tab:state-tab) ABOUT HERE]**  
  


```{r country-state-calc}
country_tab <- acs_prop_state %>%
  mutate(bpld = if_else(bpld == "United Kingdom, ns", "England", bpld),
         bpld = if_else(bpldid == 46500, 'Russia', bpld),
         bpld = if_else(bpldid == 46510, 'Belarus', bpld)) %>%
  group_by(bpld) %>%
  summarize(prop_samesex = round(100 * sum(n_same_sex, na.rm = T) / 
                                   sum(n_same_sex + n_dif_sex, na.rm = T), 2)) %>%
  arrange(desc(prop_samesex)) %>%
  .[1:10,] %>%
  mutate(prop_samesex = case_when(
    nchar(prop_samesex) == 4 ~ paste0(prop_samesex, ' %'),
    nchar(prop_samesex) == 3 ~ paste0(prop_samesex, '0 %'),
    nchar(prop_samesex) == 1 ~ paste0(prop_samesex, '.00 %'))) %>%
  rownames_to_column() %>%
  left_join(acs_ind %>%
              group_by(bpld) %>%
              summarize(`Mean policy score` = 
                          weighted.mean(origin_score, na.rm = T, w = perwt))) %>%
  select(Rank = rowname, `Country of origin` = bpld, `Proportion same-sex` = prop_samesex,
         `Mean policy score`)

country_vec <- country_tab$`Mean policy score`
names(country_vec) <- country_tab$`Country of origin`


state_tab <- acs_prop_state %>%
  mutate(bpld = if_else(bpld == "United Kingdom, ns", "England", bpld),
         bpld = if_else(bpldid == 46500, 'Russia', bpld),
         bpld = if_else(bpldid == 46510, 'Belarus', bpld)) %>%
  group_by(state) %>%
  summarize(prop_samesex = round(100 * sum(n_same_sex, na.rm = T) / 
                                   sum(n_same_sex + n_dif_sex, na.rm = T), 2)) %>%
  arrange(desc(prop_samesex)) %>%
  .[1:10,] %>%
  mutate(prop_samesex = case_when(
    nchar(prop_samesex) == 4 ~ paste0(prop_samesex, ' %'),
    nchar(prop_samesex) == 3 ~ paste0(prop_samesex, '0 %'),
    nchar(prop_samesex) == 1 ~ paste0(prop_samesex, '.00 %'))) %>%
  left_join(acs_ind %>%
              group_by(state) %>%
              summarize(`Mean policy score` = 
                          weighted.mean(state_policy, na.rm = T, w = perwt))) %>%
  rownames_to_column() %>%
  select(Rank = rowname, `State` = state, `Proportion same-sex` = prop_samesex,
         `Mean policy score`)

state_vec <- state_tab$`Mean policy score`
names(state_vec) <- state_tab$`State`
```










<!-- Figure \@ref(fig:country-map} shows the percentage of U.S. immigrants in same-sex couples based on country of origin, averaging over the 12 years of survey data. The top sending countries include interesting diversity. Though there are a number of Western countries with more progressive policies in year of departure for immigrants in our sample, notably, Malaysia, Zimbabwe, and Singapore are also in the top 10. Having countries within the top sending countries span multiple regions and cultures gives preliminary support that responses to ACS questions are not being significantly altered due to these differences. Nor does it appear as though responses to the ACS are simply a function of country-of-origin LGB policies, as these also vary significantly across the top sending countries. These patterns provide a level of validity to this measure. Table \@ref(tab:state-tab} similarly ranks U.S. state by the proportion of immigrants in same-sex couples, averaging over the period of interest. Although states with progressive policies make the top of the list, Montana, Missouri, and Florida still make the list with less affirming policy environments. -->


## Model Results

```{r state-props-mods}
country_prop <- lm(prop_same_sex ~ origin_score, data = acs_prop_state)

state_prop <- lm(prop_same_sex ~ state_policy, data = acs_prop_state)

state_prop2 <- lm(prop_same_sex ~ origin_score + state_policy, data = acs_prop_state)

# state_prop_fe <- lm(prop_same_sex ~ state_unemploy + state_income + state + 
#                         origin_score + state_policy, 
#                        data = acs_prop_state)

state_prop_fe <- lm(prop_same_sex ~ state_policy + origin_score + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = acs_prop_state)

state_prop_fe2 <- lm(prop_same_sex ~ state_policy*origin_score + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = acs_prop_state)

state_prop_fe2_2013 <- lm(prop_same_sex ~ state_policy*origin_score + post_2013 + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = acs_prop_state)

state_prop_2013 <- lm(prop_same_sex ~ post_2013 + state_policy + origin_score + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = acs_prop_state)



origin_effect <- sd(acs_ind$origin_score, na.rm = T)*coef(state_prop2)[['origin_score']]
state_effect <- sd(acs_ind$state_policy, na.rm = T)*coef(state_prop2)[['state_policy']]

```

Although the trends in Figure \@ref(fig:policy-desc) are suggestive, other factors correlated with LGB policy may be confounding results. To address this, Table \@ref(tab:state-props) presents OLS models of the U.S. state-level proportion of immigrants in same-sex couples, from a given country of origin in a given survey year. We present only coefficients of interest here, but Section D of the Online Appendix contains the full table of regression coefficients.

**[TABLE \@ref(tab:state-props) ABOUT HERE]**  

Although the trends in Figure \@ref(fig:policy-desc) are suggestive, other factors correlated with LGB policy may be confounding results. To address this, Table \@ref(tab:state-props) presents OLS models of the U.S. state-level proportion of immigrants in same-sex couples, from a given country of origin in a given survey year. We present only coefficients of interest here, but Section D of the Online Appendix contains the full table of regression coefficients.

**[TABLE \@ref(tab:state-props) ABOUT HERE]**  

Model 1 contains only one predictor: origin-country LGB policy score, for average year of immigration for immigrants in the country-state-year cell. We see that countries with more progressive LGB policies tend to send more immigrants to the U.S. who end up in same-sex couples. The average percentage of immigrants in same-sex couples is only `r mean(acs_prop_state$prop_same_sex, na.rm = T)` percent, so an increase of `r coef(country_prop)[[2]]` percent per point increase in LGB policy score represents a substantial effect.  

Model 2 turns to the role of U.S. state policy score in a given survey year, adjusting only for this variable. We see that, on average, states with more friendly LGB policies have somewhat higher proportions of immigrants in same-sex couples, on the order of `r coef(state_prop)[[2]]` percentage points per point of LGB policy score. Model 3 includes both country and state LGB scores; results barely change from the first two models, suggesting that LGB policy in both sending country and receiving state have somewhat independent relationships with LGB migration. 

According to the descriptive analysis above, immigrants in same-sex couples tend to have higher incomes and hold more prestigious occupations than immigrants in different-sex couples, and they tend to come from wealthier countries. This implies that immigrants in same-sex couples may be attracted to progressive states for their economic rather than political benefits, so Model 4 adds state and origin-country controls and fixed effects. Intriguingly, the coefficients for state score and country score *grow* in strength. More progressive sending countries are represented by greater proportions of same-sex couples, and they tend to settle in more progressive U.S. states.  

Model 5 adds an interaction between state and country LGB scores. It is positive and significant; progressive states attract higher proportions of same-sex immigrant couples, and this effect is stronger for immigrants from more progressive countries. To aid interpretation, we use simulation to predict representation of same-sex couples. Country and state LGB policy score are set to the same value for the entire sample, and the outcome is simulated and averaged over the entire sample. The left panel of Figure \@ref(fig:state-sim) has country-of-origin LGB policy on the horizontal axis and the percentage of immigrants in same-sex couples depending on vertical axis, with separate prediction lines shown for immigrants living in U.S. states with the minimum or maximum LGB policy score. The right panel switches the role of country and state policy. Both panels show that LGB immigrants tend to originate from more progressive countries and live in more progressive U.S. states. But the relationship is strongest when both country and state policy scores are high; at low levels of either score, the difference in the other score has little effect on the expected percentage of immigrants in same-sex couples.  

Finally, Model 6 augments Model 4 with a dichotomous variable for the post-2013 DOMA decision era, and Model 7 adds the same variable to Model 5. The significant post-2013 variable in both regressions implies that proportions LGB have increased across the country in the past few years. The coefficient for state policy become less important in Model 7, whereas origin-country score retains its effect. This means that, after 2013, the proportion of immigrants in same-sex couples increased across states for those from more progressive countries. With increasingly progressive federal policy, state policy appears to matter less in the post-DOMA era, or to matter only in conjunction with origin-country policy.  

Our results hold up to a variety of robustness checks presented in Section C of the Online Appendix. These include hypothetically high levels of sex misreporting; weighting by the relative size of a country's immigrant population in the U.S.; restricting the sample to married, one-immigrant, or two-immigrant couples; excluding countries with fewer than 150 and U.S. states with fewer than 500 cohabiting immigrants in our sample; and using one-year lagged measures of policy. For most robustness checks, results are substantively the same as in the main analysis. For weighted regressions that include controls and fixed effects, coefficients for origin-country and state policy scores become nonsignificant, but their interaction retains significance. This implies that results apply more at the country level than to absolute numbers of immigrants in the U.S.  One interesting finding is that results are stronger for married and one-immigrant couples than two-immigrant couples, suggesting that marriages between the U.S.-born and immigrants are driving the results. This is likely because the spousal visa pathway is open only to U.S. citizens, and one-immigrant couples are more likely to contain a U.S. citizen than two-immigrant ones.

<!-- ## Robustness Checks -->

<!-- Our results hold up to a variety of robustness checks presented in Section C of the Online Appendix; we outline the major results here. First, as mentioned above, sex misreporting may bias the results by including a non-trivial number of different-sex couples in the counts of same-sex couples. Published papers using the ACS to study same-sex couples overwhelmingly deal with potential misreporting using the method suggested by @gates_2009 that we employ: dropping respondents whose sex or relation to household head were allocated by the Census Bureau. We conduct a more stringent test of the sensitivity of our results to misreporting, using the empirical mismatch rates by survey response mode from two studies conducted by the Census Bureau in 2010 and 2016 [@kreider_2015; @kreider_2017] to reduce the proportions of immigrants in same-sex couples in data before 2019.^[Beginning in 2019, the ACS added explicit "opposite-sex" and "same-sex" categories for spouses and unmarried partners, so data for this year should be in large part purged of sex misreporting [@walker_2021]. See the Online Appendix for more details.]   -->

<!-- **[FIGURE \@ref(fig:sens) ABOUT HERE]**   -->

<!-- Here we show the hypothetical effects on our coefficients in the presence of even very high misreporting rates. Figure \@ref(fig:sens) takes Model 5 (which includes country and state controls and fixed effects) and reduces the proportions of same-sex couples in the data for pre-2019 data. It varies the percentage of misreported same-sex married couples from 0 to 90 percent and of unmarried couples from 0 to 14 percent; the horizontal axis shows a weighted average of misreporting between these two groups. Highlighted in dashed bars are the empirical mismatch rates found in the two studies by @kreider_2015 and @kreider_2017. We see that even extremely high misreporting rates in the pre-2019 ACS do not render these coefficients nonsignificant. In the Online Appendix, we present the coefficients of interest for all of the models in Tables \@ref(tab:country-props) and \@ref(tab:state-props) with proportions reduced to levels implied by the studies by @kreider_2015 and @kreider_2017. Results are substantively the same.   -->

<!-- Second, although our analysis for the most part has been at the country or state levels, there is the question of whether the aggregate trends in Tables \@ref(tab:country-props) and \@ref(tab:state-props) are driven by smaller, progressive countries that send relatively few immigrants. Hence in the Online Appendix we re-specify these models using Weighted Least Squares (Tables C.5 and C.6), weighting by the relative size of the immigrant stock in the year of immigration. Results are substantively the same; for country proportions models, the coefficient for country LGB policy score is more precisely estimated, while effects are somewhat weaker for the state proportions models. This implies that, for the typical immigrant, associations with LGB policy at country of origin are stronger than those with U.S. state LGB policy.   -->

<!-- Third, we assess whether the results in Table \@ref(tab:country-props) are driven by trends for married couples (Table C.7) or those with one U.S.-born partner (Table C.9). Results are stronger for married couples and those with no U.S.-born partner. This implies that married immigrant couples may be immigrating from countries with more progressive LGB policies, while U.S. citizens tend to partner with immigrants from somewhat less progressive countries.   -->
<!-- <!-- Fourth, we re-specify individual models with the state policy outcome treated as continuous and estimate linear models using OLS. Results are substantively the same as in Table \@ref(tab:ord}.   --> 

<!-- Fourth, if rates of mortality and return migration differ between immigrants in same- and different-sex couples, our results for same-sex representation in year of immigration could be biased. Since there are few studies on immigrants in same-sex couples, we cannot know whether or to what extent these rates might differ. However, research on the health of people in same-sex couples more generally shows similar [@cochran_2015_mortality] or slightly higher [@fenelon_2020_samesex] mortality compared to those in different-sex couples. To the extent that these trends extend to immigrants, the ACS may undercount immigrants in same-sex couples, resulting in conservative coefficient estimates. But whether and in which direction rates of return migration differ between immigrants in same- and different-sex couples is unknown. Like other scholars limiting the bias from retrospective surveys [@riosmena_2016_potential], we replicate Table \@ref(tab:country-props) with only immigrants who migrated in the year prior to each ACS survey, for years 2007 to 2018 (Table C.8). The coefficient for sending-country policy remains strong and *more* precisely estimated than in models relying on the full dataset.   -->

<!-- Finally, although we do not aim to causally assess the effect of the 2013 DOMA decision, we perform placebo tests for Models 4 and 5 in Table \@ref(tab:country-props) and Models 5, 6, and 7 in Table \@ref(tab:state-props), replacing the post-2013 dummy variable with other possible years (Tables C.1 and C.2). The coefficients for post-2013 show a clear peak compared to most other years, with a smaller standard error, supporting the idea that 2013 was a watershed moment for LGB migration to the U.S.-->



# Discussion and Conclusion
In 2013, there were `r round(same_2013 /1e3, 1)` thousand same-sex couples that included immigrants in the U.S. By 2019, this number had nearly doubled to `r round(same_2019 /1e3, 1)` thousand. Despite this expansive growth far outpacing overall migration rates, there is little demographic research investigating the characteristics of these couples or the factors influencing their migratory patterns. The research on queer migrants that does exist is largely qualitative and focused on asylum claims. Consequently, we know little about the broader population: who these migrants are, why they are leaving their home countries, or where they are choosing to locate once in the U.S. Answering these questions is important, not just because this represents an increasing number of border crossers, but because this process has the potential to reshape our conceptualization of who immigrants are, their motivations for moving, and how policy unrelated to migration can shape the aspirations and capabilities of potential migrants.  

The rising number of immigrants in same-sex couples coincides with a dramatic change in policy environments governing LGB communities, both within the U.S. and abroad. Thanks to the 2013 Supreme Court decision striking down DOMA, same-sex couples now have a broader legal pathway into the U.S. [@edwards_2013]. This project leverages changing policy landscapes at both country of origin and U.S. state of residence to understand the migratory patterns of these couples. Engaging in such a question adds to emerging demographic research evaluating how recent policy changes are influencing the health, well-being, and lifestyles of LGB people, while also recognizing that these policies differentially impact LGB people based on different social positions [@boertien_2019; @carpenter_2020; @kail_2015; @levy_2017]. In addition, this project contributes to an important area of migration studies that focuses on the interactions between the state and sexuality in conditioning migratory patterns [@carrillo_2018; @fitzgerald_2014; @cantu_2009; @difeliciantonio_2016_situating].  

To address our research questions, we take advantage of an underutilized data source: self-reports of same-sex immigrant couples in the American Community Survey from 2008 to 2019. Despite this being one of the few national surveys to identify immigrants in same-sex couples, these data have been largely untapped for this purpose. In light of possible reporting issues [@gates_2013; @goodnature_2021], we probe the sensitivity of our findings only to find remarkable robustness, even to potentially high levels of misreporting. As such, these data allow for us to make one of the first large-$N$ investigations of same-sex immigrant couples within the U.S. and to make an important addition to this area of scholarship.    

Our findings offer several important contributions to the literature. First, existing scholarship on immigrants in same-sex couples, and queer migration more broadly, is largely qualitative in nature. Consequently, it is unclear which patterns these studies illuminate generalize across the full population of immigrants in cohabitating, same-sex couples within the U.S. Current queer migration scholarship largely focuses on asylum and refugee processes, for example [@luibheid_2008; @vogler_2016]. Even research on non-refugee LGB immigrants tends to select cases from relatively repressive contexts [e.g. @manalansaniv_2003; @carrillo_2018; @vuckovicjuros_2022_lesbian]. Meanwhile, others contrast this by studying LGB individuals and couples of elite status, but similarly originating from repressive contexts and then moving toward progressive domains [@choi_2022_global; @kong_2010_chinese]. Though such studies are not seeking generalizability, this emphasis within queer migration scholarship can leave an impression that migrating from repressive to relatively more progressive contexts (either real or imagined) is typical. As such, while these experiences certainly occur, such an emphasis potentially distorts our understanding of who these immigrants are, the types of environments they are coming from, and their motivations to seek entry into the U.S. Indeed, when comparing the demographics of same-sex immigrant couples to different-sex immigrant couples, we find same-sex couples generally have higher incomes and occupational prestige and are somewhat more educated. Understanding this profile alone is an important insight; although many queer migrants are fleeing repression, the typical LGB, cohabiting migrant at the population-level within the U.S. is of privileged social standing. Knowing this descriptive point constitutes a substantial amendment to the queer migration literature. 

Second, we build upon our understanding of who these migrants are by assessing how LGB policies at their country of origin influence patterns of migration. Despite existing scholarship portraying LGB couples seeking refuge from repression, we find support for Luibhéid's [-@luibheid_2005, p. xxv] caution against this "narrative of movement from repression to freedom." Same-sex couples in our analysis are leaving countries with more progressive policy environments. As results in Table \@ref(tab:state-props) and trendlines in Figure \@ref(fig:policy-desc) reveal, immigrants in couples are coming from environments that are increasingly more open. This is true even after accounting for factors from long-standing migration models. Though more research is needed, these results, in conjunction with the fact that these same-sex couples achieve higher incomes and greater occupational prestige, describe a situation in which perhaps it is precisely due to the supportive policy environment, access to material benefits that come from marriage and employment protections, and cultural and state validation of these family formations that enable immigrants to achieve the resources necessary to migrate, seek out partnerships, and/or successfully navigate the U.S. immigration to acquire a visa. Such an explanation fits within an aspirations-capabilities framework [@dehaas_2021]: Migration is enabled when desires and resources align within an auspicious international opportunity structure.  
<!-- Or, instead of the immediate benefits of the policy itself, these supportive environments may simply encourage LGB people to feel comfortable openly identifying as being in a relationship when responding to surveys. Regardless, it appears that same-sex couples are more apt to follow the lifestyle migration pathways and motivations when coming to the U.S. as opposed to explicitly trying to leave more repressive contexts in hopes of living more open lives. -->  

And, third, after immigrants within same-sex couples migrate to the U.S., we find that they tend to cluster in states with more progressive policy environments, net of confounders. This finding is important as it helps to address a particular puzzle in LGB international migration. As a whole, the U.S. has less affirming and supportive policies for LGB communities than many countries, both in Europe and Latin America. So why would LGB immigrants from supportive environments choose to come to a country that guarantees fewer rights? One possible answer is that the unique federated system creates variation among U.S. states where more progressive and affirming policy environments can exist, even if the U.S. as a whole may be more repressive. The higher incomes, occupational prestige, and education levels of immigrants in same-sex couples -- as well as greater familiarity with progressive policies in their countries of origin -- may endow them with the necessary capital to understand and navigate the federated U.S. system. This pattern matches research finding U.S.-born LGB individuals to also be more likely to live in states with affirming policies [@beaudin_2017] and research on privileged same-sex immigrant couples in other national contexts able to navigate complex policy environments [@suen_2021_sexual]. An alternative answer as to why LGB immigrants would still migrate to the U.S. is that during his presidency, Barack Obama promoted LGBTQ equality in foreign policy, burnishing an LGB-friendly image of U.S. in the imagination of global communities despite realities.

Finally xx how to frame this within the context of previous migration research, especially "pioneer" migrants.

<!-- One unexpected finding is that different-sex couples from more LGB-friendly countries are likelier to reside in states with increasingly less LGB-friendly policies. These patterns in Figure \@ref(fig:sim} are peculiar because we suspected that state LGB policies to have no association with how different-sex couples are making decisions as to where to reside. This is not the case. One immediate question this raises is if this is a case of “straight flight.” White flight is a well-research process in which white people leave diversifying communities. After straight couples experience liberalized conditions for gay couples, do they intentionally seek to leave such conditions by moving to states that are less supportive? Alternatively, it may be that the economic and social forces pulling different-sex immigrants to these particular states may spuriously be associated with less progressive LGB policies. More research is needed to understand this interesting pattern.   -->

These insights provide a number of implications for future research. First, in demonstrating the importance of the state beyond migration-specific policy, the theoretical model presented here can be applicable in other domains. The state may condition other dimensions of well-being that are relevant to migration, such as those affected by climate policy or the salience of the Black Lives Matter movement to Black migrants. Second, finding that immigrants in same-sex couples are both coming from more open policy environments and locating to progressive U.S. states raises an important follow-up question: Are there reciprocal effects? Once rights are granted, LGB communities are instrumental in consolidating new benefits by bringing them to life -- utilizing them, demonstrating supposed negative effects do not occur, and normalizing them to non-LGB communities through interaction [@ocobock_2020_leveraging]. As such, immigrants who once had access to benefits, but no longer do, can impact policy changes by similarly bringing benefits to life and becoming agents of change [@ayoub_2019_migration; @ocobock_2020_leveraging; @suen_2021_sexual].  

Third, do different policy environments have differential effects based on couple composition? While our primary motivation is to understand differences between same- and different-sex couples, same-sex couples are a heterogeneous category. Results in the Online Appendix show stronger results for containing one U.S.-born and one immigrant member, compared to couples with two immigrants. Current data limit our ability to understand partnering processes and modes of entry into the U.S. Future research should seek to understand whether those who come to the U.S. as couples are different from those who form relationships once in the U.S., or the degree to which mixed-nativity couples result from U.S. citizens meeting a partner abroad.   

Fourth, and relatedly, this is an analysis of cohabiting couples. This particular focus has a two-part justification: the DOMA decision itself carries significant relevance to couples and, therefore, the American Community Survey's attention to cohabitating couples is appropriate. Nevertheless, this means our findings are unable to be generalized to non-cohabitating couples nor single migrants. Moreover, while we shed light on some differences between one-immigrants and two-immigrant couples in the Appendix, we do not fully assess the distinctions between these types of LGB couples as our primary emphasis is contrasting between same-sex and different-sex couples originating from the same country. Future incorporation of these different types of LGB immigrants and relationship formations, especially within nationally representative surveys, is thus urgently needed. The incorporation of such individuals may likely augment the findings presented here in a few particular ways. For example, as mentioned, supportive policy environments may both be the result of and further reinforce self-disclosure and same-sex coupledom. Consequently, data on single LGB migrants may include a more diverse set of origin policy environments while reinforcing the current socioeconomic demographics we present [@choi_2022_global]. If single LGB individuals are more likely to come from repressive contexts or may disproportionately access other visa pathways via employment or asylum, then the LGB sample identified by the ACS will systematically exclude migrants from less favorable policy environments. However, given that U.S. immigration prioritizes family unity, two-immigrant couples may disproportionately also come from countries where both immigrants previously achieved legal marriage recognition and, therefore, further reinforce these findings. Finally, although the cultural and geographic variation in responses shown in Table \@ref(tab:country-tab) suggests that responses are not a function of culture alone, future work must attend to whether response rates correlate with country-of-origin policies. As such, our results cannot generalize to U.S. immigrants who remain single or do not cohabit, whether LGB or straight. 

<!-- Fifth, although misreporting is documented in the ACS [@gates_2009; @goodnature_2021], our robustness checks following previous work utilizing these data [@goldberg_2021; @christafore_2019; @martell_2020] find remarkably consistent results. And,  -->

Fifth, although we find compelling patterns, our statistical investigation cannot determine why these couples are deciding to migrate to the U.S. and their motivations for choosing their state of residence. We encourage future scholarship to further investigate these processes and mechanisms. Finally, our results are descriptive, not causal; more rigorous methods of causal inference are necessary to invoke causality.   

<!-- Recent research has argued that sexuality is a salient factor determining immigration decisions. We show that differences between immigrants in same- and different-sex couples cannot be explained solely using classic theories of migration; policy context and sexuality interact in complicated ways to shape migratory flows.   -->

Although our focus is on same-sex couples, this study offers a broader correction to standard models of migration. Theories of lifestyle migration generally describe affluent people moving in search of a better way of life [@benson_2009]. But what the present study contributes is that sexuality shapes how that "better way of life" is conceptualized and motivated, contributing toward understanding these dynamics [@dixon_2020]. Our findings raise additional questions as to how sexuality motivates migration patterns and are (in)directly influencing seemingly economic or network dynamics, even for heterosexual couples. Finally, to borrow the line from Theda Skocpol  [-@skocpol_1985], we are "bringing the state back in." Conventionally, state policies are less integrated into models of migration, especially those policies that do not explicitly govern migration. But what this study demonstrates is that, once the DOMA ruling created a legal opening to migration for same-sex couples, it was policy specific to LGB issues, rather than to migration more broadly, that enabled their entry into the U.S. This opens up questions as to how state policies relative to a particular group, but not explicitly in the domain of immigration, create structural opportunities for certain individuals to leave their home country. This points to the importance of further studying the role of identity, and the state's governance of it, in migratory processes.


# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

<div id="refs"></div>

\newpage

# Tables

```{r tab-var, eval = F}
country_tab_ci <- acs_ind_survey %>%
  filter(yrimmig >= 1991) %>%
  filter(allocated == F) %>%
  group_by(bpld, same_sex) %>%
  summarize(survey_prop(vartype = 'ci'),
            n_same_sex = n(),
            origin_score =  weighted.mean(origin_score, na.rm = T, w = perwt)) %>%
  filter(same_sex == T) %>%
  mutate(across(c(coef, `_low`, `_upp`), function(x){x*100}))

left_join(country_tab_ci, 
          acs_ind %>%
            filter(yrimmig >= 1991) %>%
            filter(allocated == F) %>%
            count(bpld)) %>%
  write_csv(here('descriptive/files', 'country_tab_ci.csv'))

state_tab_ci <- acs_ind_survey %>%
  filter(yrimmig >= 1991) %>%
  filter(allocated == F) %>%
  group_by(state, same_sex) %>%
  summarize(survey_prop(vartype = 'ci'),
            n_same_sex = n(),
            state_policy = weighted.mean(state_policy, na.rm = T, w = perwt)) %>%
  filter(same_sex == T) %>%
  mutate(across(c(coef, `_low`, `_upp`), function(x){x*100}))

left_join(state_tab_ci, 
          acs_ind %>%
            filter(yrimmig >= 1991) %>%
            filter(allocated == F) %>%
            count(state)) %>%
  write_csv(here('descriptive/files', 'state_tab_ci.csv'))

```

```{r country-tab}
country_tab_ci <- read_csv(here('descriptive/files', 'country_tab_ci.csv'))


country_tab_ci %>% 
  mutate(`_low` = ifelse(`_low` < 0, 0, `_low`),
         across(coef:`_upp`, function(x){
           x = round(x, 2) 
           x =  case_when(
            nchar(x) == 4 ~ paste0(x, ' %'),
            nchar(x) == 3 ~ paste0(x, '0 %'),
            nchar(x) == 1 ~ paste0(x, '.00 %')) }),
         `95% CI` = paste0('(', `_low`, ', ', `_upp`, ')')) %>%
  # filter(n >= 150) %>%
  arrange(desc(coef)) %>%
  rownames_to_column('Rank') %>%
  select(Rank,
         `Country of origin` = bpld,
         `Percentage same-sex` = coef,
         `95% CI`,
         `Mean country policy score` = origin_score) %>%
  .[1:10,] %>%
  flextable() %>% 
  set_caption('Sending countries ranked by proportion immigrant couples with same-sex partners') %>%
  add_footer_lines('Source: American Community Survey 2008-2019. Authors\' calculations.') %>%
  flextable::autofit()
  # as_hux() %>%
  # set_bold(1, T)             %>%
  # set_bottom_border(1, everywhere) %>%
  # set_caption('Sending countries ranked by proportion immigrant couples with same-sex partners') %>%
  # add_footnote('Source: American Community Survey 2008-2019. Authors\' calculations.', number_format = 0) %>%
  # set_width(1)
  


  # kable(booktabs = T, 
  #       linesep = '',
  #       caption = 'Sending countries ranked by proportion immigrant couples with same-sex partners') %>%
  #     footnote(general = "American Community Survey 2008-2019. Authors\' calculations.",
  #          general_title = 'Source:',
  #          footnote_as_chunk = T) 





# acs_ind %>%
#   filter(year == 2008) %>%
#   group_by(bpld, same_sex) %>%
#   count(wt = perwt) %>%
#   pivot_wider(names_from = 'same_sex', values_from = 'n') %>%
#   arrange(desc(`TRUE`)) %>%
#   .[1:10,] %>%
#   rownames_to_column() %>%
#   left_join(acs_ind %>%
#               filter(year == 2008) %>%
#               group_by(bpld) %>%
#               summarize(origin_score = mean(origin_score)) %>%
#               select(bpld , origin_score))

```

\newpage

```{r state-tab}
state_tab_ci <- read_csv(here('descriptive/files', 'state_tab_ci.csv'))


state_tab_ci %>% 
  mutate(`_low` = ifelse(`_low` < 0, 0, `_low`),
         across(coef:`_upp`, function(x){
           x = round(x, 2) 
           x =  case_when(
            nchar(x) == 4 ~ paste0(x, ' %'),
            nchar(x) == 3 ~ paste0(x, '0 %'),
            nchar(x) == 1 ~ paste0(x, '.00 %')) }),
         `95% CI` = paste0('(', `_low`, ', ', `_upp`, ')')) %>%
  # filter(n >= 500) %>%
  arrange(desc(coef)) %>%
  rownames_to_column('Rank') %>%
  select(Rank,
         `State` = state,
         `Percentage same-sex` = coef,
         `95% CI`,
         `Mean state policy score` = state_policy) %>%
  .[1:10,] %>%
  flextable() %>% 
  flextable::set_caption('States ranked by proportion immigrant couples with same-sex partners') %>%
  flextable::add_footer_lines('Source: American Community Survey 2008-2019. Authors\' calculations.') %>%
  flextable::autofit()

# state_tab %>%
#   as_hux() %>%
#   set_bold(1, T) %>%
#   set_bottom_border(1, everywhere) %>%
#   set_caption('States ranked by proportion immigrant couples with same-sex partners') %>%
#   add_footnote('Source: American Community Survey 2008-2019. Authors\' calculations.', number_format = 0) %>%
#   set_width(1)

# state_tab %>%
#   kable(booktabs = T, 
#         linesep = '',
#         caption = 'States ranked by proportion immigrant couples with same-sex partners') %>%
#       footnote(general = "American Community Survey 2008-2019. Authors\' calculations.",
#            general_title = 'Source:',
#            footnote_as_chunk = T) 


```



\newpage

```{r state-props}
mod_list <- list(country_prop, state_prop, state_prop2, state_prop_fe, state_prop_fe2, 
                 state_prop_2013, state_prop_fe2_2013)

robust_list <- list()
for(mod_num in 1:length(mod_list)){
  robust_list[[mod_num]] <- coeftest(mod_list[[mod_num]], 
                                     vcov = vcovCL(mod_list[[mod_num]], cluster = ~ state + Country), 
                                     save = T)
}

huxtable::huxreg(robust_list,
       coefs = c('Country LGB policy score' = 'origin_score', 
                 'State LGB policy score' = 'state_policy',
                 'State score × country-score' = 'state_policy:origin_score',
                 'Post-2013' = 'post_2013TRUE'),
       note = "{stars}. Country and state two-way clustered standard errors are shown in parentheses. State controls include unemployment rate and per-capita income. Country controls include population-weighted distance, contiguous border, common official language, common ethnic language, colonial relationship, per-capita GDP differential, unemployment differential, proportion same-country stock, and democracy. Source: American Community Survey 2008-2019. Authors' calculations.",
       stars = c(`†` = 0.1, `*` = 0.05, `**` = 0.01, `***` = 0.001),
       statistics = c('Observations' = 'nobs')) %>%
  insert_row(c('State controls and FEs?', 'no', 'no', 'no', rep('yes', 4)), after = nrow(.) - 2) %>%
  insert_row(c('Country controls and FEs?', 'no', 'no', 'no', rep('yes', 4)), after = nrow(.) - 2) %>%
  huxtable::set_caption('Percent same-sex in by country of origin, U.S. state, and survey year.') %>%
  huxtable::set_width(1) %>%
  huxtable::set_all_padding(0)
```


\newpage

# Figures
```{r total-pop, fig.cap = 'Estimated totals of different- and same-sex couples containing one ("mixed status") or two immigrants, 2008-2019, with 95% confidence intervals. Vertical line placed at the year 2013, when DOMA was overturned.'}

total_df %>%
  mutate(`Total population (thousands)` = n/1000,
         n_low = n_low/1000, n_upp = n_upp/1000,
         Year = year,
         imm_couple = if_else(imm_couple == 'one', 'Mixed-status couples', 'Two-immigrant couples')) %>%
  ggplot(aes(x = Year, y = `Total population (thousands)`, 
             linetype = imm_couple)) +
  geom_line() +
  geom_ribbon(aes(ymin = n_low, ymax = n_upp, group = imm_couple, col = NULL), alpha = .2) +
  geom_vline(xintercept = 2013, linetype = 1) +
  facet_wrap(~same_sex, scales = "free") +
  xlim(2007, 2019) +
  theme(legend.position="bottom") +
  geom_blank(aes(y = 0))

```



\newpage

  

```{r desc-country, fig.height = 6, fig.cap = "Origin country-level descriptive statistics for immigrants in couples 2008-2019, for year of migration, with survey weights and 95% confidence intervals. Currency in thousands of 1999 dollars."}


distance <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(distw, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'C. Distance to Country of Origin (km)')

unemp <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(unemp_dif, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'B. Unemployment Rate Difference with U.S.')

prop_stock <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(stock_prop, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'D. Proportion U.S. Immigrant Stock')

wage_dif <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(wage_dif, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'A. GDP Difference with U.S.')

vdem <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(vdem, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'E. Democracy Score (V-Dem)')


bind_rows(wage_dif, vdem, prop_stock, unemp, distance) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = var, linetype = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = var_low, ymax = var_upp, group = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var_name, scales = "free", ncol = 2) +
  ylab('') +
  theme( # legend.justification=c(1,0),
  #     legend.position=c(.22,.82),
  #     legend.background = element_rect(fill = alpha("white", 0.5))
      legend.position="bottom",
      text = element_text(size = 8))
 
```



\newpage

```{r desc-ind, fig.cap = 'Individual descriptive statistics for immigrants in couples 2008-2019, with survey weights and 95% confidence intervals. Currency in thousands of 1999 dollars.'}
income <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(inctot, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'C. Annual Gross Income (Thousands)')

hs <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex'),
         hs = if_else(educ %in% c('HS', 'some col', 'college'), 1, 0)) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(hs, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'A. High School Completion')

college <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex'),
         college = if_else(educ == 'college', 1, 0)) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(college, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'B. College Completion')

# state_income <- acs_ind_survey %>%
#   mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
#   group_by(year, same_sex) %>%
#   summarize(var = survey_mean(state_income, vartype = "ci", na.rm = T)) %>%
#   mutate(var_name = 'U.S. State of Residence Per Capita Income')

hwsei <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(hwsei, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'D. Hauser and Warren Socioeconomic Index')


bind_rows(income, hs, college, hwsei) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = var, linetype = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = var_low, ymax = var_upp, group = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var_name, scales = "free", ncol = 2) +
  ylab('') +
  theme( # legend.justification=c(1,0),
  #     legend.position=c(.22,.82),
  #     legend.background = element_rect(fill = alpha("white", 0.5))
      legend.position="bottom",
      text = element_text(size = 8))
```

\newpage 

```{r policy-desc, fig.cap = 'Mean country-of-origin and U.S. state policy index score for immigrants in same- and different-sex couples, 2008-2019, with 95% confidence intervals.'}


origin_df <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(`Policy Score` = survey_mean(origin_score, vartype = "ci", na.rm = T)) %>%
  mutate(var = 'Country-of-Origin Index')

state_df <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(`Policy Score` = survey_mean(state_policy, vartype = "ci", na.rm = T)) %>%
  mutate(var = 'U.S. State Index')

bind_rows(origin_df, state_df) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = `Policy Score`, 
             linetype = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = `Policy Score_low`, ymax = `Policy Score_upp`, 
                  group = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var, scales = "free") +
  theme(legend.justification=c(1,0),
        legend.position=c(.26,.6))
```

\newpage

```{r country-map, fig.cap = 'Percentage of cohabiting immigrants in the U.S. in same-sex couples by country of origin, averaging over ACS survey years 2008 to 2019.'}
country_tab <- acs_prop_state %>%
  mutate(bpld = if_else(bpld == "United Kingdom, ns", "England", bpld),
         bpld = if_else(bpldid == 46500, 'Russia', bpld),
         bpld = if_else(bpldid == 46510, 'Belarus', bpld)) %>%
  group_by(bpld) %>%
  summarize(prop_samesex = round(100 * sum(n_same_sex, na.rm = T) / 
                                   sum(n_same_sex + n_dif_sex, na.rm = T), 2)) %>%
  rename(region = bpld)

country_tab$region <- recode(country_tab$region,
                              'Burma (Myanmar)' = 'Myanmar',
                              'England' = 'UK',
                              'Cambodia (Kampuchea)' = 'Cambodia',
                              'Egypt/United Arab Rep.' = 'Egypt', 
                              'Guyana/British Guiana' = 'Guyana',
                              'Korea' = 'South Korea',
                              'Republic of Georgia' = 'Georgia',
                              'South Africa (Union of)' = 'South Africa',
                              'Sri Lanka (Ceylon)' = 'Sri Lanka',
                              'Yemen Arab Republic (North)' = 'Yemen',
                              'Other USSR/"Russia"' = 'Russia'
)


# 
# # Adding immigrant data to the maps package data
# sPDF <- joinCountryData2Map(country_tab,
#                             joinCode = "NAME",
#                             nameJoinColumn = "Country")
# 
# # Dropping Antarctica
# new_world <- subset(sPDF, continent != "Antarctica")
# 
# mapDevice(titleSpace = 0) #create world map shaped window
# 
# mapParams <- mapCountryData(new_world,
#                             nameColumnToPlot="prop_samesex",
#                             catMethod = "fixedWidth",
#                             numCats = 100, 
#                             colourPalette = "white2Black",
#                             addLegend = F,
#                             mapTitle = '')
# 
# do.call(addMapLegend,
#         c(mapParams,
#           legendLabels = "limits",
#           legendIntervals = "page",
#           horizontal = F))


world_map <- map_data('world') %>%
  filter(!(region %in% c('Antarctica', 'Greenland'))) 

left_join(country_tab, world_map, by = 'region') %>%
  bind_rows(left_join(world_map, country_tab, by = 'region') %>%
  filter(is.na(prop_samesex)) %>%
  mutate(prop_samesex = NA)) %>%
  ggplot(mapping = aes(x = long, y = lat, group = group, 
                       fill = prop_samesex)) +
                       #fill = case_when(prop_samsex == 0.0012515612315, 'white', prop_samesex))) +
  #geom_polygon(color = "gray90", size = 0.1) +
  geom_polygon(color = "gray75", size = 0.1) +
  coord_fixed(1.3) +
  theme_map(font_family = 'Palatino') +
  labs(fill = "Same-sex (%)")+ 
  scale_fill_gradient(low = uclablue, high = uclagold, na.value = 'gray95') +
  #scale_fill_gradient(low = "white", high = "black") +
  theme(legend.justification=c(0,0), legend.position=c(.1,.1),
         legend.title=element_text(size=9),
         legend.key.size = unit(0.3, 'cm'),
         legend.text = element_text(size = 8))

```

\newpage

```{r state-map, fig.cap = 'Percentage of cohabiting immigrants in same-sex couples in U.S. states, averaging over ACS survey years 2008 to 2019.'}
state_tab <- acs_prop_state %>%
  mutate(bpld = if_else(bpld == "United Kingdom, ns", "England", bpld),
         bpld = if_else(bpldid == 46500, 'Russia', bpld),
         bpld = if_else(bpldid == 46510, 'Belarus', bpld)) %>%
  group_by(state) %>%
  summarize(prop_samesex = round(100 * sum(n_same_sex, na.rm = T) / 
                                   sum(n_same_sex + n_dif_sex, na.rm = T), 2))
  

state_tab$region <- tolower(state_tab$state)

# Getting the map ready
us_states <- map_data("state")

state_imm <- left_join(us_states, state_tab)


ggplot(data = state_imm,
       mapping = aes(x = long, y = lat, group = group, fill = prop_samesex)) +
  geom_polygon(color = "gray75", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_map(font_family = 'Palatino') + 
  labs(fill = "Same-sex (%)") + 
  scale_fill_gradient(low = uclablue, high = uclagold, na.value = 'gray95') +
  #scale_fill_gradient(low = "white", high = "black") +
  theme(legend.justification=c(0,0), legend.position=c(.05,.0),
    legend.title=element_text(size=9),
    legend.key.size = unit(0.3, 'cm'),
    legend.text = element_text(size = 8))
```

\newpage
```{r state-sim-calc, eval = F}
## Varying origin score while keeping state score constant (at min and max)
var1 <- 'origin_score'
var <- 'state_policy'
sim_country <- list()
for(cat_loop in unique(acs_prop_state[,var1])){
    var_vals <- c(min(acs_prop_state[,var]), max(acs_prop_state[,var]))
    for(val in var_vals){
      print(paste0(cat_loop, ' ', val))
      
        sim_dat <- acs_prop_state
        sim_dat[, var1] <- cat_loop
        sim_dat[, var] <- val

        pred_dat <- predict(state_prop_fe2, newdata = sim_dat, se.fit = T,
                            type = 'response') %>%
          as.data.frame()

       sim_country[[paste0(var,cat_loop, val)]] <- pred_dat %>%
         summarize(fit = mean(fit, na.rm = T),
                   se.fit = sqrt(mean(se.fit^2, na.rm = T)),
                   fit.low = fit - 1.96*se.fit,
                   fit.high = fit + 1.96*se.fit,
                          origin_score = cat_loop,
                          state_policy = val)
    }
}

bind_rows(sim_country) %>%
  write_csv(here('descriptive/files', 'sim_country.csv'))

## Varying state score while keeping origin policy constant (at min and max)
var1 <- 'state_policy'
var <- 'origin_score'
sim_state <- list()
for(cat_loop in unique(acs_prop_state[,var1])){
    var_vals <- c(min(acs_prop_state[,var]), max(acs_prop_state[,var]))
    for(val in var_vals){
      print(paste0(cat_loop, ' ', val))
        sim_dat <- acs_prop_state
        sim_dat[, var1] <- cat_loop
        sim_dat[, var] <- val

        pred_dat <- predict(state_prop_fe2, newdata = sim_dat, se.fit = T, 
                            type = 'response') %>%
          as.data.frame()

       sim_state[[paste0(var,cat_loop, val)]] <- pred_dat %>%
         summarize(fit = mean(fit, na.rm = T),
                   se.fit = sqrt(mean(se.fit^2, na.rm = T)),
                   fit.low = fit - 1.96*se.fit,
                   fit.high = fit + 1.96*se.fit,
                          state_policy = cat_loop,
                          origin_score = val) 
    }
}

bind_rows(sim_state) %>%
  write_csv(here('descriptive/files', 'sim_state.csv'))
  
```

```{r state-sim, fig.cap='Predicted percentage of immigrants in same-sex couples by country-state, using Model 5 from Table 3.'}
plot1 <- read_csv(here('descriptive/files', 'sim_country.csv'))  %>%
  mutate(state_policy = ifelse(state_policy == min(state_policy), 'Min. U.S. state policy', 'Max. U.S. state policy')) %>%
  ggplot(aes(x=origin_score, y=fit, linetype=state_policy)) +
  geom_line() + 
  geom_ribbon(aes(ymin=fit.low, ymax=fit.high, 
                  col = NULL), alpha = .2) +
  xlab('Country-of-origin LGB policy score') +
  ylab('Percentage of immigrants in same-sex couples') +
  theme(legend.position=c(.4,.8))

plot2 <- read_csv(here('descriptive/files', 'sim_state.csv')) %>%
  mutate(origin_score = ifelse(origin_score == min(origin_score), 'Min. country policy', 'Max. country policy')) %>%
  ggplot(aes(x=state_policy, y=fit, linetype=origin_score)) +
  geom_line() + 
  geom_ribbon(aes(ymin=fit.low, ymax=fit.high, 
                  col = NULL), alpha = .2) +
  xlab('U.S. state-of-residence LGB policy score') +
  ylab('') +
  theme(legend.position=c(.4,.8))


plot1 + plot2 +  plot_layout(guides = 'keep') & theme(text = element_text(size = 9))
```



