---
output:
  # bookdown::pdf_document2:
  bookdown::word_document2:
    reference_docx: "word-template.docx"
    toc: no
    number_sections: no
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    latex_engine: xelatex
always_allow_html: true
header-includes:
  # - \usepackage{setspace}\doublespace
  - \setlength{\headheight}{13.6pt}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{Nathan I. Hoffmann}
  - \rhead{`r format(Sys.time(), '%B %e, %Y')`}
editor_options: 
  chunk_output_type: console
citeproc: no
fontfamily: mathpazo
fontsize: 11pt
geometry: margin= 1in
indent: yes
link-citations: yes
linkcolor: blue
lang: 'en-US'
bibliography: '/Users/nihoffm/My Drive/Projects/2020 Same-Sex Immigrant Couples/Same-Sex Immigration.bib'
title: "The Geography of Immigrants in Same-Sex Couples in the United States"
# date: "September 1, 2024"
date: "`r format(Sys.time(), '%B %e, %Y')`"
author: 
- Nathan I. Hoffmann, Department of Sociology, Emory University
- Kristopher Velasco, Department of Sociology, Princeton University
abstract: "The policy landscape for same-sex couples in the United States has changed rapidly in recent years, and the number of immigrants in same-sex couples have markedly increased. Yet little is known about where these lesbian, gay, bisexual, and queer (LGBQ) immigrants settle and experience these shifting policy environments. Using American Community Survey data from 2008-2023 and original datasets, this paper studies the geographic context of immigrants in same-sex cohabiting couples in the U.S., comparing them to them to immigrants in different-sex couples and the U.S.-born in same-sex couples. We find that the distribution of immigrants in same-sex couples is expanding across the U.S. over time. Individually, these immigrants resemble U.S.-born LGBQ Americans, with higher incomes and educational attainment. At the geographic level, they appear more distinct, clustering in areas with more LGBTQ+ nonprofit organizations, greater numbers of immigrants, higher density, and higher property values than the areas that straight immigrants or the U.S.-born in same-sex couples tend to live in. Our findings contribute to a fuller understanding of this rapidly growing population and its unique characteristics."
---

<!-- Turn off hyphenation -->

<!-- \usepackage[none]{hyphenat} -->

```{r setup, include=F}
library(knitr)
library(broom)
library(huxtable)
library(sandwich)
library(lmtest)
library(here)
library(stargazer)
library(patchwork)
library(flextable)
library(kableExtra)
library(maps)
library(tigris)
library(fiftystater)
library(ggthemes)
library(srvyr)
library(tidyverse)

# find . -name "Icon?" -print0 | xargs -0 rm -rf

knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, # results = 'asis',
                      cache.lazy = F, dpi = 1200)


options("yaml.eval.expr" = TRUE)

uclablue = '#2774AE'
gray = '#808080'
black = '#000000'
uclagold = '#FFB81C'
rust = '#C47A2B'
# ucla_palette = c(uclablue, black, gray)
ucla_palette = c(black, uclablue, rust)
palette_dark <- rev(RColorBrewer::brewer.pal(n=3,"Dark2"))

# theme_set(theme_cowplot(font_family = 'Palatino') + 
theme_set(theme_classic() + 
            theme(legend.title=element_blank(), 
                  axis.line = element_line(size = .3), 
                  axis.ticks = element_line(size = .3),
                  panel.grid.major.y = element_line('grey80', size = .3),
                  legend.background = element_rect(fill = "transparent")))
ggplot <- function(...) ggplot2::ggplot(...) + scale_color_manual(values = ucla_palette) +
  scale_fill_manual(values = ucla_palette)

# theme_set(theme_classic() + 
#             theme(axis.line = element_line(size = .3), 
#                   axis.ticks = element_line(size = .3), 
#                   legend.title=element_blank(), 
#                   panel.grid.major.y = element_line('grey80', size = .3),
#                   legend.background = element_rect(fill = "transparent")))
# ggplot <- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Dark2") +
#   scale_fill_brewer(palette="Dark2")

# library(colorspace)
# colorspace_palette = 'RdPu'
# ggplot <- function(...) ggplot2::ggplot(...) +
#   scale_color_discrete_sequential(palette=colorspace_palette) +
#   scale_fill_discrete_sequential(palette=colorspace_palette)

options(scipen=1, digits=2)
knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")
})
knitr::opts_chunk$set(error=TRUE) 

set_flextable_defaults(font.family = 'Times New Roman',
                       # padding.bottom = 1,
                       # padding.top = 1,
                       # padding.left = 3,
                       # padding.right = 1,
                       theme_fun = 'theme_booktabs',
                       digits = 3, 
                       big.mark = ',',
                       line_spacing = 1)

```

```{r load, include = F}
# acs_ind <- readRDS(here('data', 'acs_ind.rds')) 


# acs_count_base <- read_csv(here('data', 'acs_count_base.csv')) 

all_geo <- read_csv(here('data', 'immigrants_geo.csv')) %>% 
  mutate(group = case_when(
    imm_couple == 'none' ~ 'Same-sex, non-immigrant',
    imm_couple != 'none' & same_sex == T ~ 'Same-sex, immigrant',
    imm_couple != 'none' & same_sex == F ~ 'Different-sex, immigrant'
  )) %>%
  mutate(children = ifelse(nchild > 0, 'has children', 'no children'),
         ethnicity = case_when(
           hispan != 'not hispanic' ~ 'Hispanic',
           race == 'white' ~ 'White',
           race %in% c("other asian or pacific islander", "chinese", "japanese") ~ "Asian",
           race == 'black/african american' ~ 'Black',
           !is.na(race) ~ 'Multiracial or Other'),
         lgbt_nonprofits = ifelse(year >= 2022, NA, lgbt_nonprofits),
         immigrant_nonprofits = ifelse(year >= 2022, NA, immigrant_nonprofits),
         married = ifelse(year < 2012 & same_sex == T,
                                 NA, married))

all_geo_survey <- all_geo %>% 
  as_survey_design(weights = perwt)

immigrants <- all_geo %>%
  filter(imm_couple != 'none') 

immigrants_survey <- immigrants %>% 
  as_survey_design(weights = perwt)

acs_geo <- read_csv(here('data', 'acs_geo_survey.csv')) 



  # left_join(select(read_csv(here('geo/files', 'nonprofits_puma.csv')), -statefp), by = c()) %>%
  # mutate(lgbt_nonprofits = ifelse(is.na(lgbt_nonprofits), 0, lgbt_nonprofits)) %>%
  # left_join(select(read_csv(here('geo/files', 'imm_nonprofits.csv')), -statefp)) %>%
  # mutate(immigrant_nonprofits = ifelse(is.na(immigrant_nonprofits), 0, immigrant_nonprofits))

# acs_geo <- read_csv(here('data', 'acs_geo_survey.csv')) %>%
#   mutate(across(c(bachelors_puma, black_puma, hispanic_puma, owned_puma, immigrant_puma, poverty_100_puma,
#                   poverty_200_puma, unemployed_puma), 
#                 ~.x*100))

prop_tab <- acs_geo %>%
  select(state, puma) %>%
  distinct() %>%
  left_join(immigrants %>%
    mutate(same_sex = ifelse(same_sex == T, 'same_sex', 'dif_sex')) %>%
    # filter(year == 2019) %>%
    group_by(state, puma, same_sex) %>%
    summarize(n = sum(perwt)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'same_sex', values_from = 'n')) %>% 
  replace(is.na(.), 0) %>%
  # complete(year, state, puma, same_sex, fill = list('n' = 0)) %>%
  mutate(prop = same_sex/(same_sex + dif_sex)*100) %>%
  filter(!is.nan(prop))


prop_tab_year <- acs_geo %>%
  select(state, puma, year) %>%
  left_join(immigrants %>%
    mutate(same_sex = ifelse(same_sex == T, 'same_sex', 'dif_sex')) %>%
    # filter(year == 2019) %>%
    group_by(year, state, puma, same_sex) %>%
    summarize(n = sum(perwt)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'same_sex', values_from = 'n')) %>% 
  replace(is.na(.), 0) %>%
  # complete(year, state, puma, same_sex, fill = list('n' = 0)) %>%
  mutate(prop = same_sex/(same_sex + dif_sex)*100) %>%
  filter(!is.nan(prop)) %>%
  select(state, puma, year, prop) %>%
  left_join(acs_geo) 


gay_prop_tab_year <- acs_geo %>%
  select(state, puma, year) %>%
  left_join(all_geo %>%
    filter(group != 'Different-sex, immigrant') %>%
    # filter(year == 2019) %>%
    group_by(year, state, puma, group) %>%
    summarize(n = sum(perwt)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'group', values_from = 'n')) %>% 
  replace(is.na(.), 0) %>%
  # complete(year, state, puma, same_sex, fill = list('n' = 0)) %>%
  mutate(prop = `Same-sex, immigrant`/(`Same-sex, immigrant` + `Same-sex, non-immigrant`)*100) %>%
  filter(!is.nan(prop)) %>%
  select(state, puma, year, prop) %>%
  left_join(acs_geo) 

  # left_join(immigrants %>% group_by(year, state, puma) %>%
  # summarize(across(c(density, state_policy, lgbt_nonprofits, immigrant_nonprofits), ~mean(.x, na.rm = T))))


count_tab_year <- bind_rows(
  select(acs_geo, state, puma, year) %>% mutate(group = 'Different-sex, immigrant'),
  select(acs_geo, state, puma, year) %>% mutate(group = 'Same-sex, immigrant'),
  select(acs_geo, state, puma, year) %>% mutate(group = 'Same-sex, non-immigrant')) %>%
  left_join(all_geo %>%
    # filter(year == 2019) %>%
    group_by(year, state, puma, group) %>%
    summarize(n = sum(perwt)) %>%
    ungroup()) %>% 
    replace(is.na(.), 0) %>%
  # complete(year, state, puma, same_sex, fill = list('n' = 0)) %>%
  left_join(acs_geo) 
  
  
variable_names <- c('Bachelor\'s degree (%)' = 'bachelors_puma', 
                    'Black (%)' = 'black_puma',
                    'Hispanic (%)' = 'hispanic_puma', 
                    'Annual personal income (thousands of 2023 USD)' = 'inctot_puma',
                    # 'Log mean personal income' = 'log_inctot_puma', 
                    'Individuals own home (%)' = 'owned_puma', 
                    'Mean age' = 'age_puma', 
                    'Immigrant (%)' = 'immigrant_puma',
                    'Individuals under 100% of poverty line (%)' = 'poverty_100_puma', 
                    'Individuals under 200% of poverty line (%)' = 'poverty_200_puma', 
                    'Mean HWSEI occupation score' = 'hwsei_puma', 
                    'Unemployed (%)' = 'unemployed_puma', 
                    'Mean individual\'s value of home ($1000s)' = 'valueh_puma',
                    'Mean individual\'s rent' = 'rent_puma', 
                    'Mean cost of electricity' = 'costelec_puma', 
                    'Density (persons per sq. mile)' = 'density', 
                    'State LGBQ policy score' = 'state_policy', 
                    'Mean number of LGBT nonprofits' = 'lgbt_nonprofits',
                    'Mean number of immigrant nonprofits' = 'immigrant_nonprofits',
                    'Annual Gross Income (Thousands)' = 'inctot',
                    'Hauser and Warren Socioeconomic Index' = 'hwsei',
                    'High school graduate' = 'hs',
                    'Four-year college graduate' = 'college',
                    'Number of children' = 'nchild',
                    'Has children' = 'has_children',
                    'Married' = 'married',
                    'White' = 'white',
                    'Black' = 'black',
                    'Asian' = 'asian',
                    'Hispanic' = 'hispanic',
                    'Multiracial or Other' = 'other',
                    'Male' = 'male',
                    'Female' = 'female',
                    'Age' = 'age',
                    'n' = 'n',
                    'n (weighted)' = 'n_weighted')


variable_names_swap <- names(variable_names)
names(variable_names_swap) <- as.vector(variable_names)


sample_size <- all_geo %>%
  group_by(group) %>%
  summarize(n = n(),
            n_weighted = sum(perwt))
```



# Introduction

Residential equity and the geographic distribution of people are linked to a variety of important demographic and social outcomes such as health and well-being, exposure to environmental and social hazards, access to quality education, and social capital, among others. Legal landscapes are salient factors influencing these settlement patterns. This is especially true for LGBTQ+ migrants; recent policy changes in the U.S. and elsewhere have opened pathways for asylees seeking refuge from repressive contexts, elite LGBTQ+ migrants selecting ideal cosmopolitan destinations, and those wanting to be reunified with their binational partner [@carrillo_2018; @mai_2009; @gorman-murray_2009; @vogler_2016;@choi_2022_global; @difeliciantonio_2016_situating; @hoffmann_2024_policy; @yang_2024_rethinking; @jones_2025_jigsaw]. A notable policy development relevant to this latter group came in 2013, when the U.S. Supreme Court overturned the Defense of Marriage Act and required the U.S. government to recognize marriages between same-sex spouses. This decision radically changed the immigration landscape: For the first time, same-sex spouses of U.S. citizens and lawful permanent residents were eligible to file a spousal or fiancée petition for an immigrant visa [@edwards_2013]. In the years since, the U.S. population of lesbian, gay, bisexual, and queer (LGBQ) immigrants has grown rapidly. As @hoffmann_2024_policy show, numbers of different-sex couples (whether cohabiting or married) containing immigrants increased by 22 percent from 2008 to 2019 (from 7.8 million to 9.5 million), while those of corresponding same-sex couples grew by 140 percent in the same period (from 44 thousand to 107 thousand). 

Despite increased interest on this growing population [@seo_2025_sexualities], especially in the wake of the DOMA decision, little is known about *where* LGBQ migrants settle and experience these policy shifts after entering the U.S. Do they gravitate toward LGBTQ+-friendly cities and states or are they more concerned with job opportunities and cost of living? How have these patterns changed over time, especially in the context of different shifts in local policies that either expand or restrict opportunities relevant to both LGBTQ+ people and immigrants? Research makes clear that LGBTQ+ migrants can often feel caught between their sexual identity and migrant community [@acosta_2008_lesbianas]. As such, do patterns of settlement align with their fellow migrants or, does their dispersion mirror their U.S-born LGBQ counterparts? Addressing this question offers important insights into how social processes ultimately affect populations caught in potentially countervailing forces, both in the U.S. and other countries where legal changes enable migration and partnering.  

This paper implements descriptive analyses to study how settlement patterns of immigrants in same-sex cohabiting couples in the U.S. respond to local and national policy changes as well as other local conditions. To tease apart the unique characteristics of our focal group -- immigrants in same-sex couples in the U.S. -- we compare them to two other groups: immigrants in different-sex couples and U.S.-born people in same-sex couples (where the partner is also U.S.-born). We compare individual characteristics of these three groups before turning to their geographic distribution and characteristics of their places of residence. We leverage data from the American Community Survey, spanning 2008 to 2023, to do so.  

<!-- We find that the distribution of immigrants in same-sex couples is expanding across the U.S. over time. Individually, these immigrants resemble U.S.-born LGBQ Americans, with higher incomes and educational attainment. At the geographic level, they appear more distinct, clustering in areas with more LGBT nonprofit organizations, greater numbers of immigrants, higher density, and higher property values than U.S.-born Americans in same-sex couples. Our findings contribute to a fuller understanding of this rapidly growing population and its unique characteristics. -->

Results suggest that the distribution of immigrants in same-sex couples expanded across the U.S. during this time period. Moreover, in individual characteristics, LGBQ migrants resemble the U.S.-born in same-sex couples more than fellow immigrants. Although their post-DOMA rates of marriage are higher than U.S.-born LGBQ individuals (likely due to the immigration benefits of marriage), their high incomes, occupational prestige, and educational attainment as well as few children closely match the characteristics in U.S.-born same-sex couples. In terms of geographic settlement, however, they behave as more of a distinct group. In some respects they resemble other immigrants, residing in areas with high proportions of Hispanic and immigrant individuals. But their levels of segregation from these other groups is high; they concentrate in localities with more LGBTQ+ nonprofit organizations, higher population density, and higher property values than areas where either immigrants more broadly or the U.S.-born in same-sex couples tend to live. Our findings contribute to a fuller understanding of this rapidly growing population and in which aspects they appear more similar to their fellow migrants, their fellow sexual minorities, or as a distinct population.  



# Background
We bridge two streams of literature -- one on immigrant settlement patterns and the other on LGBTQ+ residential dispersion -- to understand where immigrants in same-sex relationships live in the U.S. Migration scholars are particularly interested in the distribution of migrants to understand processes of spatial integration and assimilation [@waters_2005; @gordon_1964_assimilation]. But this question of spatial distribution is also relevant to the residential patterns of LGBTQ+ populations. Research finds that LGBTQ+ individuals, particularly gay men, historically concentrate in urban neighborhoods, i.e., "gayborhoods." But this concentration is declining -- perhaps indicating greater social acceptance as LGBTQ+ individuals integrate into other geographic areas [@spring_2024_new]. As immigrants in same-sex couples continue to rapidly expand in number [@hoffmann_2023_sexuality], this case allows us to understand where these individuals locate amid differing pressures and whether they hew to their co-migrants or, instead, live in areas more aligned with their U.S.-born LGBQ counterparts. Of course, a third option exists, which is that immigrants in same-sex couples, along with LGBTQ+ migrants more broadly, represent a distinct population whose residential patterns are unlike either of their two adjoining counterparts [@acosta_2008_lesbianas; @solveigar-oggudmundsdottir_2024_northtonorth]. Knowing where LGBTQ+ migrants locate and how they subjectively experience place can better determine where services and social support need to be targeted and which social forces are most salient.

## Immigrant Settlement in the United States
There is a long history in migration studies of investigating settlement patterns. Theories of destination choice after entering a country are similar to theories of migration more broadly. At one level, neoclassical economic theories predict that migrants will follow wage and unemployment differentials to places with high labor demand [@hatton_2005a; @todaro_1980]. @newbold_1999_spatial shows how these economic factors help explain settlement patterns in the U.S., also pointing to the importance of government welfare programs and the local immigration rate. This last effect highlights the importance of migrant networks that share information and resources to lower the cost of migration and settling in the destination, promoting settlement in migrant-dense locations [@massey_1987]. Similarly, institutions that arise to recruit workers and ease entry and settlement can promote migration to particular destinations [@hernandez-leon_2013]. Local political factors can also influence migrant settlement. For example, @watson_2013_enforcement shows that U.S. metropolitan areas that adopt 287(g) agreements -- allowing local agencies to enforce Federal immigration law -- see significantly greater propensity of immigrants to leave and relocate to more hospitable parts of the U.S.  

<!-- Indeed, as migration policies become more strict in the U.S., we generally see this having xxeffectxx.   -->

Settlement also varies by individual-level characteristics. @south_2005_migration find greater spatial assimilation for Latino immigrants -- i.e., they are more likely to live among non-Hispanic, White Americans -- when they have greater human and economic capital and greater English proficiency. @iceland_2008_immigrant find similar results, with migrants more likely to live among the U.S.-born when they have been in the U.S. longer, have higher earnings, and own a home. 

Scholars have turned to studying so-called "new" immigrant destinations in the U.S. in recent years. In a trend first noticed in the 1990s and early 2000s [@durand_2000_changing; @zuniga_2005_new], migration to urban areas and California slowed and migrants relocated to rural areas and small towns in the Sun Belt and the Deep South. As an overabundance of labor and increased immigration enforcement pushed migrants out of these traditional destinations, relocation of manufacturing and meat processing plants pulled workers to lower-wage, right-to-work states [@flippen_2021_new]. Due to their lack of established migrant communities, dynamics of settlement differ in these new destinations: Comparing new and established migrant destinations, @hall_2013_residential finds greater segregation in new destinations, even when controlling for other local factors and individual characteristics. If immigrants in same-sex couples are to experience similar migration pressures, then we would expect their spatial patterns to fall along these lines.

## Distribution of Same-Sex Couples in the United States
Concerns of residential distribution, segregation, and equity have also been a prominent line of research for LGBQ populations, generally, and same-sex couples, specifically. Similar to settlement patterns of migrants, LGBQ Americans are also influenced by the role of economic conditions. For example, those in same-sex couples are generally found to be in areas with more socioeconomic advantage, such as with more college-educated individuals, higher property values, and access to better local amenities [@florida_2003_cities; @black_2002_why; @lee_2018_healthrelated]. These patterns reflect same-sex couples' somewhat higher discretionary income and ability to take advantage of local amenities and spend more on housing [@black_2002_why]. We should note that the relative privilege of same-sex couples stands in contrast with other sections of the LGBTQ+ community -- trans individuals and other queer people who do not fit into the categories of gay and lesbian experience higher levels of economic insecurity and issues of personal safety [@zeeman_2019_review}. Similar to heterosexual couples, cohabiting and married same-sex couples are of higher socioeconomic status than single individuals.  

Consistent findings also highlight the role of network effects, with same-sex couples located in areas with high concentrations of other same-sex couples -- similar to ethnic and migrant enclaves. This concentration pattern extends beyond the U.S. as well; LGBQ couples in several European countries and Australia cluster in urban settings and this effect is even stronger for single individuals [@carpio-pinedo_2025_castro; @wimark_2014_city]. Besides network effects, this clustering is also protective for same-sex couples and LGBQ individuals against violence and harassment [@hayslett_2011_out; @lee_2018_healthrelated]. This concentration pattern, however, has encouraged the majority of research investigating residential patterns of same-sex couples to focus on these urban clusters, i.e., "gayborhoods" [@ghaziani_2016_there; @nash_2015_lesbians], including their distinct cultures, economies, and on-going transformations against sexuality-based residential segregation [@ghaziani_2016_there; @spring_2024_new]. Finally, another set of studies examines the role of progressive policies such as same-sex marriage and non-discrimination protections in influencing attitudes toward migration and actual migration practices [@baumle_2023_effects]. @marcen_2022_effect finds that the legalization of same-sex marriage across U.S. states led to an increase in men, not women, in same-sex couples moving to those states, but the authors are not able to provide an explanation for this gendered pattern. 

Taken together, this research emphasizes same-sex couples living in urban gayborhoods with access to more progressive legal environments. Indeed, the imagination of a more liberatory city has characterized motivations for LGBQ individuals to migrate, whether internally within a country or between countries, in research across U.S., Europe, Latin America, and Asia [@difeliciantonio_2020_migration; @wimark_2014_city; @fortesdelena_2022_urban; @luo_2023_geography] But both changes in demographic trends and academic interest have shifted greater attention to LGBQ individuals and same-sex couples living beyond these particular locales. From 2000 to 2010, and continuing to 2020, the concentration of same-sex couples in segregated, urban communities has declined [@spring_2013_declining; @spring_2024_new]. Current research finds that these historic gayborhoods have become gentrified, and rising costs have led many same-sex couples to move to other parts of a city or out of these urban centers altogether [@spring_2021_gentrification]. Meanwhile, others find that discriminatory lending practices also encourage dispersion of same-sex couples to other locations due to simple discrimination on housing loans [@spring_2024_new; @sun_2019_lending]. Furthermore, the residential patterns of queer women have for a long time tended to be more diffuse [@spring_2013_declining]. These trends underscore arguments made by @stone_2018_geography in the U.S. and @waitt_2011_its in Australia: Researchers of LGBTQ+ geography need to look beyond "great cities" and toward more mid-sized cities and rural locales in order to more holistically understand LGBTQ+ geographic patterning [@annes_2012_coming]. Indeed, in the U.S., same-sex couples are increasingly represented in rural communities and states with conservative governments, which may reflect displacement out of cities or increasing acceptance and integration of these types of couples across the country [@brown_2008_urban; @marino_2024_visualizing]. Moreover, @stella_2015_lesbian underscores the importance to this work, as she finds that lesbians in Russia have qualitatively distinct experiences depending on whether they live in Moscow or smaller, regional cities.

<!-- There has also been a bias to examine urban residential patterns despite clear movement toward suburban and rural locations [@marino_2024_visualizing]. More research is needed, then, to understand patterns at lower levels of aggregation -- both geographically and at observation unit [@spring_2024_new].  -->


## Expectations for Immigrants in Same-Sex Couples in the United States
Although the settlement patterns of immigrants in same-sex have not previously received significant study, existing work can help motivate empirical expectations. We anticipate three possible patterns: LGBQ migrants may live in similar areas to other migrants, in areas that other LGBQ people prefer, or areas more unique to LGBQ migrants. Existing research suggests that any of these three scenarios might hold. 

Why might LGBQ migrants live near other migrants? First, the same economic, network, and policy factors that shape migrant settlement more broadly may affect LGBQ migrants. For LGBQ migrants who follow standard migration pathways such as moving for employment or following parents or siblings who have previously migrated, they are likely to reside in areas that are similar to those of other migrants who have chosen the same channels. Second, migrants from non-Western settings or from poorer socioeconomic conditions are often excluded from LGBTQ+ spaces in their countries of destination, facing barriers to their integration into LGBTQ+ residential and social infrastructure [@el-tayeb_2012_gays; @dhoest_2023_intersectional]. For example, while @solveigar-oggudmundsdottir_2024_northtonorth finds privileged migrants, particularly white migrants, capable of integrating into existing LGBTQ+ spaces and patterns of social life in Iceland, LGBTQ+ migrants who are not white or of high SES do not necessarily find themselves welcome in these spaces. If this trend holds more widely, LGBQ migrants may remain in coethnic communities.  

As another possibility, why might LGBQ migrants live near U.S.-born LGBQ people? First, while existing research tends to focus on asylees and individuals seeking refuge in the U.S. and other Western countries [@chossiere_2021_refugeeness; @held_2023_queer], there is evidence that the broader population of LGBQ migrants may be more privileged [@hoffmann_2023_sexuality], which we examine in greater detail below. This may lead to LGBQ migrants living near U.S.-born LGBQ people because more privileged migrants tend to spatially assimilate, living in the same areas as the U.S.-born [@south_2005_migration; @iceland_2008_immigrant]. Second, @hoffmann_2024_policy find that LGBQ migrants are more likely than heterosexual partnered migrants to have U.S.-citizen partners. Emergent research suggests that these U.S.-born partners help to integrate these LGBQ migrants into normative cultures of U.S. life, rather than being pulled toward immigrant communities [@seo_2024_quotidian]. These dynamics may lead them to living near LGBQ Americans. Third, there is some evidence that LGBTQ+ migrants are drawn to progressive local destinations [@marcen_2022_effect]. LGBTQ+ migrants coming to Western countries often see cities as desirable destinations, especially as cities have come to promote themselves as inclusive of LGBTQ+ communities [@binnie_2006_cosmopolitan; @binnie_2014_relational; @klett-davies_2021_border].  

Lastly, why might LGBQ migrants have residential patterns that do not reflect either their conational or U.S. counterparts? Due to their intersectional positioning, LGBQ migrants face unique forms of discrimination from either population [@acosta_2008_lesbianas]. @hoffmann_2024_how show that U.S. citizens tend to see LGB migrants as less deserving of admission and less culturally similar to the U.S. than other migrants. If LGBQ migrants face unique antagonism from the U.S.-born, they may be pushed into their own unique enclaves -- either through individual-level discriminatory experiences or larger, structural-level factors filtering them into distinct destinations. Relatedly, LGBQ migrants may seek refuge from navigating the "ethnic closet" [@cui_2024_queer]; weary of downplaying either their ethnic or queer identity in order to find acceptance, they may find the greatest comfort in areas with other LGBQ migrants. If these migrants tend to be more socioeconomically privileged, they may have the economic power to seek out spaces where they feel most at ease, in line with broader findings around "lifestyle migration" [@dixon_2020; @benson_2012].  



<!-- To date, most quantitative research on same-sex couples implicitly assumes that all respondents are U.S.-born, since nativity status is neither explicitly asked nor specified. Consequently, it is unknown whether the patterns described above will hold for LGBQ immigrants as well. However, existing qualitative work can inform some empirical expectations. Studies demonstrate that LGBTQ+ migrants coming to Western countries such as the U.S. see cities as desirable destinations, especially as cities have come to promote themselves as inclusive of LGBTQ+ communities [@binnie_2006_cosmopolitan; @binnie_2014_relational; @klett-davies_2021_border]. However, migrants from non-Western settings or from lower socioeconomic conditions are often excluded from LGBTQ+ spaces in their countries of destination, facing barriers to their integration into LGBTQ+ residential and social infrastructure [@el-tayeb_2012_gays; @dhoest_2023_intersectional]. For example, @solveigar-oggudmundsdottir_2024_northtonorth finds privileged migrants, particularly white migrants, capable of integrating into existing LGBTQ+ spaces and patterns of social life in Iceland, but not necessarily LGBTQ+ migrants outside of these categories. Similarly, @cui_2024_queer develop the concept of "ethnic closet" to discuss how queer migrants navigate their racialized disadvantage, where these individuals downplay either their ethnic or queer identity depending on the social context. Considering that recent waves of migration into the U.S. consist largely of people from Latin America and Asia, these may be the precise populations that experience exclusion. Consequently, their residential patterns may not necessarily look like their U.S. LGBQ counterparts but, instead, find greater inclusion in areas dominated by their co-migrants. -->

<!-- Existing qualitative research on the residential processes of LGBTQ+ migrants often highlight their exclusion from mainstream LGBTQ+ spaces because these studies tend to focus on asylees and individuals seeking refuge in the U.S. and other Western countries [@chossiere_2021_refugeeness; @held_2023_queer]. Consequently, these samples are more likely to experience social exclusion that the broader population of LGBTQ+ migrants. Moreover, this profile of LGBTQ+ migrant stands in contrast to the typical characteristics of immigrants in same-sex relationships at the country level. @hoffmann_2023_sexuality find that immigrants in same-sex couples, whether married or cohabiting, are quite privileged. Compared to immigrants in different-sex couples, those in same-sex couples come from richer, more democratic, and more LGBQ-friendly countries. They also tend to be more individually privileged than heterosexual migrants from the same country of origin. These are the immigrants, then, that are perhaps best positioned to integrate into existing LGBTQ+ spaces and, in effect, have residential patterns that are more akin to their U.S. counterparts. One factor contributing to the privilege of this group and their potential resemblance to U.S.-born LGBQ people is because a significant number of these immigrants are in binational unions with a U.S.-born counterpart [@hoffmann_2024_policy]. After the end of the Defense of Marriage Act in 2013, U.S citizens could sponsor a spousal and fiance visa for their same-sex partner for the first time -- dramatically increasing the number of immigrants in same-sex couples in the U.S. [@hoffmann_2024_policy]. In addition, emergent research suggests that these immigrants get integrated into the normative cultures of U.S. life rather than their U.S.-born partner being pulled toward immigrant communities [@seo_2024_quotidian].   -->

<!-- Of course, there is also the possibility that immigrants in same-sex couples, and LGBTQ+ migrants more broadly, represent their own distinct population and have residential patterns that do not reflect either their co-national or U.S. counterparts. @hoffmann_2024_how show that U.S. citizens tend to see LGBQ migrants as less deserving of admission and less culturally similar to the U.S. than other migrants. Due to their greater personal privilege as well as this LGBQ migrant-specific antagonism, LGBQ migrants may concentrate in their own unique enclaves.   -->


# Data and Methods

Our main source of data is the American Community Survey (ACS) one-year estimates from 2008 to 2023 [@ruggles_2024_ipums]. We exclude data from the year 2020, when data was not of adequate quality. Each year, the ACS surveys a 1-percent representative sample of U.S. households about their education, occupation, income, family structure, immigration status, country of origin, location, and a variety of other individual and household attributes. We define a same-sex couple as two individuals of the same sex in the same household who report their relationship as "spouse" or "unmarried partner." We limit the sample to individuals ages 18 to 64, and immigrants in the sample must have migrated at age 18 or older.  

We do not focus solely on married individuals for a few reasons. First, there is an issue of data availability; until 2012, the ACS re-labeled same-sex married couples as unmarried partners, even in states where same-sex marriage was legal. Including same-sex couples coded as "unmarried partners" allows us to look at a broader range of years, and some of these couples may in fact have been married before 2012. Second, previous work has shown that the DOMA decision was associated with a rise in same-sex cohabitation for immigrants, regardless of whether they were married or unmarried [@hoffmann_2024_policy]. This suggests that policy around marriage had broader effects not limited to married individuals. Lastly, the end of DOMA opened a pathway to visas for fiancé(e)s (as well as spouses); however, the ACS does not capture whether unmarried partners are engaged to be married, so we cannot measure the prevalence of fiancé(e)s in the data. For these reasons, we include both married and unmarried cohabiting couples in our analyses. As part of these, we examine rates of marriage beginning in the year 2012. In the Online Appendix, we replicate Tables \@ref(tab:desc-ind) and \@ref(tab:desc-geo) with the samples restricted to only married couples, using only data from 2012 onward. Descriptive statistics are very similar to those for the broader sample that includes cohabiting couples.   

<!-- In addition, same-sex marriage was not legal in all 50 states until the 2015 *Obergefell v. Hodges* Supreme Court decision. -->
<!-- ^[In the Online Appendix, we replicate Tables \@ref(tab:desc-ind) and \@ref(tab:desc-geo) with the samples restricted to only married couples. The ACS recoded same-sex married couples as unmarried partners until 2012, so in this section of the Appendix we use only data from 2012 onward. Descriptive statistics are very similar to those for the broader sample including cohabiting individuals.]  -->

We consider the spatial distribution of three groups: immigrants in same-sex couples (unweighted $N$ = `r sample_size$n[sample_size$group == 'Same-sex, immigrant']`), immigrants in different-sex couples ($N$ = `r sample_size$n[sample_size$group == 'Different-sex, immigrant']`), and individuals in same-sex couples where neither individual is an immigrant ($N$ = `r sample_size$n[sample_size$group == 'Same-sex, non-immigrant']`). The weighted sample is equivalent to `r sum(all_geo$perwt)/1e6` million individuals over the 15 years of data.

We use "LGBQ" to refer to all individuals who may be in romantic relationships with members of the same sex, although we recognize that some individuals in same-sex relationships may not identify as lesbian, gay, bisexual, or queer. We also recognize that we are not able to identify bisexual (or pansexual, multisexual, etc.) individuals cohabiting with different-sex partners. Furthermore, measuring the prevalence of same-sex couples in the U.S. is difficult [@michaels_2013]. As in most nationally representative demographic work on same-sex couples [@baumle_2013; @baumle_2019], we are able to identify only LGBQ couples that cohabit; unpartnered LGBQ individuals and those who do not live with their partner are not included in the analysis [@baumle_2009, p. 6]. In addition, LGBQ individuals who do not feel comfortable with the "spouse" or "unmarried partner" labels of the ACS are not in the sample. There may also be some issues with measurement error: Misreporting may result when different-sex couples accidentally misspecify the gender of one of the partners [@gates_2009; @goodnature_2021]. Beginning in 2008, the Census Bureau made changes to ACS gender and partnership questions in order to prevent such errors [@u.s.censusbureau_2013], so we rely on data only from 2008 onward, but difficulties remain. If even a small number of different-sex couples misreport one partner's sex, the counts of same-sex couples will be inflated. Following @gates_2009, we remove all respondents that had either their relationship or sex variable allocated by the Census Bureau. This is the strategy used by most studies of same-sex couples in the ACS [e.g. @boertien_2019; @gates_2013; @goldberg_2021; @christafore_2019; @martell_2020].  

Most of our variables come from the ACS. To place our geographic investigations in context, we rely on some individual-level characteristics. These include annual gross income (in thousands of 2023 U.S. dollars), education (with categories for being a high school or four-year college graduate), the Hauser-Warren Socioeconomic Index for occupational prestige [HWSEI, @hauser_1997_socioeconomic], the number of the respondent's own children in the household, race and ethnicity (with categories for non-Hispanic White and Black, Hispanic, Asian, and Multiraical or Other), age, and two-category sex.^[We are constrained in reporting sex and gender by the questions asked in the ACS, which include only a binary option for sex and no question about gender.]  

For geographic variables, we calculate values at the level of the Public Use Microdata Area (PUMA) -- a geographic unit of analysis that covers a population of at least 100,000 and does not cross state lines. PUMAs partition the entirety of the United States. We consider a variety of variables at the PUMA level in the ACS. First are variables relevant to economic considerations; neoclassical economic theory suggests that immigrants choose to settle in areas with greater economic opportunities. We expect that people want to live in more socioeconomically advantaged areas, as long as the cost of living is not too high. We include the percentage of individuals in the PUMA who hold at least a bachelor's degree and who are unemployed. We also consider the mean PUMA-level values of personal income, Hauser-Warren occupational prestige score, and value of home (in thousands of 2023 dollars)^[The VALUEH home value variable is top-coded at $1,000,000.].

Second, we consider variables relevant to social considerations. Based on theories of immigrant networks and the migration industry, we expect the immigrants will want to live near other immigrants and with organizations that ease migration. Individuals may also take other social aspects into consideration, such as racial composition and density. From the ACS, we calculate the percentage of individuals in the PUMA who are immigrants (born outside the U.S.), who identify as Black, and who identify as Hispanic. We also use a measure from the ACS of the density of the PUMA in persons per square mile.

We also use social data from two other sources. First, to examine LGBQ policies at state of destination, we use original datasets. We predict that LGBQ migrants will prefer to live in states with more progressive LGBQ policy. To create the U.S. state policy index, we compile data from the Movement Advancement Project^[https://www.lgbtmap.org/], a leading LGBQ organization in the U.S. that collects data on a number of relevant policies. A higher score represents more progress state-level policies. Progressive policies include full marriage equality, state recognition of civil unions and domestic partnerships, ban on all employment and housing discrimination based on sexual orientation, hate crime protections based on sexual orientation, legal joint adoption by same-sex couples, and a ban on conversation therapy for minors. For regressive policies, we consider criminalization of sodomy, state constitutional bans of marriage equality, religious freedom exemptions to discriminate against same-sex couples in adoption, and state-level bans on local non-discrimination ordinances encompassing sexual orientation. Due to data unavailability at the time of writing, we use 2019 values for 2021 to 2023. The state index ranges from `r min(all_geo$state_policy, na.rm = T)` to `r max(all_geo$state_policy, na.rm = T)`, and the mean state policy score in this time period is `r mean(all_geo$state_policy, na.rm = T)`.  

Our second outside data source is a measure of the prevalence of LGBTQ+ and immigrant-serving nonprofits in a given PUMA. We expect that LGBQ migrants will choose to live in areas with more LGBTQ+-serving nonprofits, and we expect that migrants more broadly will prefer to live in areas with more immigrant-serving nonprofits. These data come from the Internal Revenue Service's Business Master File (BMF). The BMF collects basic, administrative data on all tax-exempt entities required to submit an annual tax document; commonly referred to as the Form 990. For our purposes, we restrict our sample to just 501c(3) charitable organizations because under U.S. tax code, these are the ones that more most conceptualized as community "nonprofits" as opposed to political political action committees or other entities [@paxton_2020_does]. Additionally, the IRS assigns every organization an activity code from the National Taxonomy of Exempt Entities (NTEE) based on their primary function. NTEE codes are widely used in academic research, or otherwise, to classify an organization. There are 26 major codes, and each major code as accompanying that prvodies greater specificity. For example, the general code "A" stands for Arts and Culture organizations, but "A24" specifically represents folk art organizations. Between major and minor NTEE codes, there are more than 600+ distinct classifications. Consequently, ee classify nonprofits with NTEE codes "Ethnic & Immigrant Centers (code: P84) and "Immigrant Rights" (code: R21) as immigrant serving. While LGBTQ+ nonprofits are those designed as "LGBT Centers" (code: P88) and "Lesbian & Gay Rights" (code: R26). We additionally supplement our classification of LGBTQ+ nonprofits with those used by @velasco_2022_deconstructed, who found LGBTQ+ nonprofits operating under other classification designations. To calculate a PUMA-level measure of these nonprofits, we use their ZIP code to approximate the number of nonprofits in a PUMA in a given year. Data for this variable are available only through 2021. 

Our analyses are descriptive, presenting and comparing the contours of the populations of immigrants in same-sex couples, immigrants in different-sex couples, and U.S.-born in same-sex couples. We present descriptive tables, graphs showing change in variables over time, as well as maps showing the geographic distribution of LGBQ migrants. We also present values of the index of dissimilarity over time [@massey_1990_american, @frey_2018_blackwhite], measuring segregation of immigrants in same-sex couples from immigrants in different-sex couples or from non-immigrants in same-sex couples. The index of dissimilarity ranges from 0 to 1 and can be interpreted as the proportion of a minority group that would have to move to a new area (in this case PUMA) in order to equalize the distribution of the two groups. In all calculations, we use survey weights from the ACS.


# Results


## Individual Characteristics

Table \@ref(tab:desc-ind) presents means of individual variables for immigrants in different-sex couples, immigrants in same-sex couples, and U.S.-born individuals in same-sex couples, while Figure \@ref(fig:desc-ind-figures) shows how the group means of these variables change over time. Beginning with socioeconomic characteristics, individuals in same-sex couples -- regardless of nativity --  are generally more advantaged than immigrants in different-sex couples. This is true for annual income, occupational prestige score, high school completion, or college completion. Income and occupational prestige are nearly indistinguishable between the two groups of LGBQ individuals. Interestingly, compared to U.S.-born LGBQ individuals, LGBQ immigrants have lower rates of high school completion but higher rates of college completion. This implies a bimodal distribution: LGBQ immigrants consist of individuals with high education and those with relatively little education.  

LGBQ individuals in general are much less likely than heterosexual ones to have children. However, before 2015, LGBQ immigrants were more likely to have children than the LGBQ U.S.-born. Children of LGBQ people tend to be from previous heterosexual relationships, which may have been more likely for LGBQ immigrants in the past. In recent years both LGBQ groups are equally likely to have children, at about 20 percent. Marriage is one variable with distinct values for each group. Cohabiting immigrants in different-sex couples are very likely to married, at over 90 percent. LGBQ U.S.-born individuals have relatively low marriage rates, at a little under 50 percent. LGBQ immigrants lie somewhere between these two groups, at about 70 percent married. This higher proportion may reflect the additional benefits that that immigrants receive from marriage to a U.S. resident or citizen, notably visas, residency, and a path to citizenship [@hoffmann_2024_policy].  

Turning to identity characteristics, immigrants in same-sex couples are more likely to be White than those in different sex couples (28 compared to 19 percent), but much less likely to be white than the U.S.-born in same-sex couples (77 percent). Proportions of Black and Hispanic LGBQ immigrants are similar to those for heterosexual immigrants, but the proportion Asian is lower (23 percent for LGBQ compared to 31 percent for heterosexuals). LGBQ individuals are slightly younger than heterosexual immigrants here. As for sex, while there are slightly more women than men among heterosexual immigrants and the LGBQ U.S.-born, there are nearly twice as many men as women among LGBQ migrants.


## Geographic Characteristics

To show broadly how the geography of immigrants in same-sex couples has changed over the past 15 years, Figure \@ref(fig:state-map) presents the percentage coupled immigrants in each state who are in same-sex couples and how these figures have changed over time. The figure shows maps for three time periods: 2008-2012, 2013-2017, and 2018-2023. Most states had greater proportions of immigrants in same-sex couples in 2023 than in 2008. In addition, the pattern of growth is meaningful. Shortly after the end of the Defense of Marriage Act in 2013 -- when U.S. citizens and permanent residents could finally sponsor the visa of a same-sex partner -- their growth was concentrated in the Northeast, where states were relatively early adopters of same-sex marriage, including Massachusetts (2004), Connecticut (2008), Vermont (2009), New Hampshire (2010), New York (2011), Maine (2012), and Maryland (2013). However, in the next period (2018-2023), more of the map shows higher percentages of immigrants in same-sex couples, including more relatively conservative states.  

In the Appendix, we present maps of subgroup distributions. Aggregating all of our years of data, these maps show the percentage of immigrants in same-sex couples in a given state who belong to four types of subgroups: race and ethnicity, sex, whether they have children, and whether they are married (rather than also including unmarried partners). Higher proportions of LGBQ immigrants with children live in the Midwest than the coasts. States close to the Mexican border have higher proportions of LGBQ migrants who are Hispanic, while those in the Midwest and East have somewhat higher proportions that are Asian. Marriage and sex do not seem to be strongly structured by geography.  

```{r segregation-calc, eval = F}
yearly_theil <- all_geo %>%
  group_by(year) %>%
  group_modify(~ segregation::mutual_total(data = .x, 
                                          group = 'group', 
                                          unit = c('puma', 'state'),
                                          weight = 'perwt')
               ) %>%
  filter(stat == 'H') %>%
  mutate(comparison = 'All groups',
         index = 'Theil Index')

seg_imm <- all_geo %>%
  filter(group %in% c('Same-sex, immigrant', 'Different-sex, immigrant')) %>%
  group_by(year) %>%
  group_modify(~segregation::dissimilarity(data = .x, 
                                          group = 'group', 
                                          unit = c('puma', 'state'),
                                          weight = 'perwt',
                                          se = T)
               ) %>%
  mutate(comparison = 'Same-sex, imm. vs. dif.-sex, imm.',
         index = 'Dissimilarity Index')


seg_lgbq <- all_geo %>%
  filter(group %in% c('Same-sex, immigrant', 'Same-sex, non-immigrant')) %>%
  group_by(year) %>%
  group_modify(~segregation::dissimilarity(data = .x, 
                                          group = 'group', 
                                          unit = c('puma', 'state'),
                                          weight = 'perwt',
                                          se = T)
               )   %>%
  mutate(comparison = 'Same-sex, imm. vs. same-sex, non-imm.',
         index = 'Dissimilarity Index')
  
bind_rows(seg_imm, seg_lgbq) |>
  write_csv(here('geo/files/', 'seg.csv'))
```

```{r seg-load}
seg <- read_csv(here('geo/files/', 'seg.csv'))

seg_imm <- filter(seg, comparison == 'Same-sex, imm. vs. dif.-sex, imm.')

seg_lgbq <- filter(seg, comparison == 'Same-sex, imm. vs. same-sex, non-imm.')
```


We next consider segregation at the PUMA level between immigrants in same-sex couples and the other two groups. As shown in Figure \@ref(fig:segregation), the index of dissimilarity is quite high in both cases, though it has declined slightly between 2008 and 2023. In 2008, this index is equal to about `r seg_imm %>% filter(year == 2008) %>% pull(est) %>% round(2)` when comparing to immigrants in different-sex couples, meaning that `r 100*(seg_imm %>% filter(year == 2008) %>% pull(est) %>% round(2))` percent of immigrants in same-sex couples would need to move to a different PUMA to equalize the distribution between these two groups. By 2023, it decreases to `r seg_imm %>% filter(year == 2023) %>% pull(est) %>% round(2)`. As for segregation from the U.S.-born in same-sex couples, the index is even higher, at `r seg_lgbq %>% filter(year == 2008) %>% pull(est) %>% round(2)` in 2008 and `r seg_lgbq %>% filter(year == 2023) %>% pull(est) %>% round(2)` in 2023. To put these numbers in context, @massey_1990_american considers an index of dissimilarity above 0.6 to be "high segregation," and Black-White segregation by census tract in the most segregated U.S. cities ranges from 0.66 to 0.80 [@frey_2018_blackwhite]. This gives support to the idea that LGBQ immigrants are settling in unique enclaves in the U.S., away from both the broader immigrant and LGBQ populations.  

Table \@ref(tab:desc-geo) presents descriptive statistics for PUMA-level characteristics, separated by the three groups of interest: immigrants in different-sex couples, immigrants in same-sex couples, and non-immigrants partnered with same-sex non-immigrants. In the table, all variables are averaged over the full range of survey years. In some ways, immigrants in same-sex couples tend to be more similar to immigrants in different-sex couples than non-immigrants in same-sex couples. For example, immigrants in same- and different-sex couples live in areas with similar proportions of immigrants and Hispanic people. However, immigrants in same-sex couples live in areas with somewhat greater advantage than the other two groups. They live in areas with higher incomes and higher occupational prestige, and the home and rent prices of their areas are somewhat higher than those of immigrants more broadly. They also live in denser areas.    

Clear differences also arise around two variables relevant to LGBTQ+ people. First, although all three groups tend to live in relatively progressive states with a score of at least 3 -- representing at least three progressive state policies -- immigrants in same-sex couples live in states with an average score of nearly 4. Although a difference of 1 might appear small, it may be the difference between employment protection and no employment protection, or between conversion therapy for minors being allowed or forbidden. Also notable are differences around the number of LGBTQ+ nonprofits. On average, immigrants in same-sex couples live in PUMAs with 1.3 LGBTQ+ nonprofits, while LGBQ non-immigrants live in PUMAs with 0.8 and immigrants in different-sex couples with 0.3. They also live in areas with somewhat more immigrant nonprofits, with an average of 1.1. Surprisingly, the number of immigrant nonprofits for the average immigrant in a different-sex couple and average non-immigrant in a same sex couple are quite similar, at about 0.9. 

Figure \@ref(fig:desc-geo-figures) shows how these differences are fairly stable over time. Arguably, the gap in residential characteristics between LGBQ and heterosexual migrants has slightly narrowed over time, with heterosexual migrants somewhat catching up in recent years to the advantaged geographic characteristics enjoyed by LGBQ migrants. Similarly, one thing these descriptive trend lines show is that while the DOMA decision contributed to the rapid expansion of immigrants in same-sex couples in the U.S., the quality of these immigrants appears unchanged. That is, at first look, there is not a clear post-2013 disruption to the trends shown at either the individual or geographic-levels; instead, there is a continuation of patterns already identified. This suggests that the type of immigrant in our sample and where they live are largely stable during this period.


# Discussion and Conclusion
As dozens of countries have come to legally recognize same-sex spouses, these legal changes have been accompanied by increased numbers of and attention on LGBQ migrants in these places [@hoffmann_2024_policy; @luibheid_2018_samesex; @yue_2008_samesex; @chauvin_2021_class; @vuckovicjuros_2022_sexualities]. In 2013, the U.S. followed this trend after Supreme Court overturned the Defense of Marriage Act and required the federal government to begin recognizing marriages between same-sex spouses. Once immigrants in same-sex couples had access to spousal and fiancé(e) visas, their numbers increased rapidly [@hoffmann_2024_policy]. Where have these immigrants settled? Investigating this question is important given the strong connection between residential distribution and a variety of important demographic and social outcomes such as health and well-being, exposure to environmental and social hazards, access to quality of education, and social integration.  

First, we find that immigrants in same-sex couples are increasing their geographic distribution across the United States from 2008 to 2023 and increasingly live in a variety of states and geographic areas outside of particular cities. These trends mimic findings by @marino_2024_visualizing and @spring_2024_new showing more geographic distributions at the county-level and metropolitan-level for all same-sex couples. Moreover, it underscores the call by queer scholars in the U.S. and other Western countries to look beyond "great cities" when investigating the geographies of LGBTQ+ populations [@stone_2018_geography; @waitt_2011_its; @stella_2015_lesbian].  

Second, at the individual level, immigrants in same-sex couples are more similar to U.S.-born counterparts in same-sex couples. That is, both groups generally have higher incomes, socioeconomic status, education rates, and are unlikely to have children in comparison with immigrants in different-sex couples. Notwithstanding the theoretical and human importance of LGBQ asylum seekers and refugees, numbers of these are relatively small. Overall, LGBQ migrants tend to be socioeconomically privileged, a finding supported in other research [@hoffmann_2023_sexuality]. In addition, LGBQ migrants in couples have higher rates of marriage (80 percent) than U.S.-born LGBQ couples (60 percent), likely due to the immigration benefits of same-sex marriage that followed the end of DOMA. Indeed, this marriage mechanism may be an important factor explaining individual-level similarities: When entering these binational marriages, Americans may select partners that more closely resemble their own sociodemographic characteristics.

While this group resembles the LGBQ U.S.-born individually, at the geographic level, there is evidence that immigrants in same-sex couples have their own distinct settlement patterns. At the PUMA level, we find high levels of segregation from immigrants in different-sex couples or non-immigrants in same-sex couples. While these measures of segregation have declined over the past two decades, they remain as high as Black-White segregation in some of the U.S.'s most segregated cities. This suggests that LGB immigrants tend to live in their own distinct areas, away from the broader immigrant and LGBQ populations.  

This distinctiveness is further supported by geographic characteristics of the areas in which LGBQ migrants live. In some ways, the areas hosting these LGBQ immigrants resemble areas where straight immigrants live, with high proportions of Hispanic and immigrant individuals. However, these areas also differ in important ways; LGB migrants tend to live in areas with more LGBTQ+ nonprofit organizations, greater density, and higher property values than areas that host either the broader population of immigrants or the LGBQ U.S.-born. Overall, these immigrants live in diverse, cosmopolitan, and LGBTQ+ supportive locations that have relatively diverse racial and migratory compositions. This is counter to U.S.-born same-sex couples that are increasingly spreading to conservative states and non-urban locations. Furthermore, they do not seem to be following immigrants in different-sex couples' pathways to "new immigrant destinations." Due to their privileged socioeconomic status, they may be in a better position to consider lifestyle migration factors than others [@carrillo_2018; @dixon_2020]. However, much more research is needed to understand the direct mechanisms driving these distributional trends. Our interpretation here is only suggestive for future research. Moreover, it is not immediately clear how these patterns compare to LGBQ migrants in other countries, as limited representative studies have been conducted.

This study is subject to a number of limitations that future research should attempt to overcome. First, ACS data limit analysis to LGBQ migrants in couples who live together, excluding single or non-cohabiting LGBQ migrants. Reliance on data from the decennial census and the American Community Survey has necessarily constrained insights to those in cohabiting, same-sex couples (whether married or unmarried). Representative quantitative data on single queer people is quite limited, due to the rarity of including survey questions about sexual orientation and gender identity. Second, PUMAs are large (including at least 100,000 people), so analysis can not be as fine-grained as would be ideal. While urban areas have many PUMAs (for example, New York City has 55), in rural areas PUMAs can be quite vast; research at lower levels of aggregation is necessary to better understand the settlement of these areas. Third, summary statistics from the ACS are only proxies for the lived spatial experiences of LGBQ migrants in these areas. In-depth qualitative work is necessary to understand how LGBQ migrants consider various aspects of place and space in deciding where to settle. This latter point is particularly important because how individuals experience space can vary widely [@acosta_2008_lesbianas; @solveigar-oggudmundsdottir_2024_northtonorth]. And while existing work details how LGBQ migrants experience different locations, it is less clear how and why they choose the particular destinations they do. Relatedly, our data do not allow us to distinguish between migrants choosing a destination of their own accord and migrants being compelled to settle in a particular area due to reasons beyond their control or structural sorting processes. Further inquiries on this front can help connect our descriptive findings to subjective realities. 

Our study generates new insights on the complex relationship between sexuality, place, and legal landscapes. One important implication relates to the politics of sexuality and migration. Recently, scholars have pointed to the unique position of queer migrants to reshape politics and activism related to both migration and LGBTQ+ rights due to how they alter conceptualizations of the nation [@girard_1987_subversion; @luibheid_2002_entry; @richardson_2018_sexuality; @terriquez_2015_intersectional]. @chavez_2013_queer, in particular, argues that queer migrants are able to construct "coalitional moments" in which their unique social positioning enables them to bridge multiple communities and advance a new type of politics. Moreover, while we find LGBQ migrants in more LGBTQ+-friendly locations, there is some evidence from Europe and Hong Kong that these migrants are the very ones advancing these progressive policy changes -- as opposed to simply selecting into destinations with such policies already [@ayoub_2019_migration; @suen_2021_sexual]. Indeed, as our findings showcase, one reason for this is because the LGBQ migrants in our sample are relatively privileged, perhaps enabling them to leverage their skills and resources to advocate for social change. However, since we find these migrants to primarily locate into their own, distinct communities, how diffuse these legal effects can be may be limited. Moreover, the fact that we are primarily analyzing relatively privileged, married LGBQ migrants suggests these may not be those pushing for more radical politics but, instead, are those advancing rights that adhere to existing social structures of the nation [@puar_2018_terrorist]. Regardless, our study demonstrates that LGBQ migration into the U.S. is rapidly expanding due to new legal recognitions of same-sex marriage. As these migrants settle across the U.S., more research will be needed to understand how they are altering social fabrics and geography.


\newpage

# References

<div id="refs"></div>

\newpage

# Tables


```{r desc-calc, eval = F}
desc_tab <- all_geo_survey %>%
  group_by(group) %>%
  summarize(across(c(bachelors_puma, black_puma, hispanic_puma, inctot_puma,
                     owned_puma, age_puma, immigrant_puma,
                     poverty_100_puma, poverty_200_puma, hwsei_puma, 
                     unemployed_puma, valueh_puma,
                     rent_puma, costelec_puma, density, state_policy, 
                     lgbt_nonprofits, immigrant_nonprofits), 
            ~ survey_mean(.x, na.rm = T)))

write_csv(desc_tab, here('geo/files', 'desc_tab.csv'))

desc_tab_ind <- all_geo_survey %>%
  mutate(hs = if_else(educ %in% c('HS', 'some col', 'college'), 1, 0),
         college = if_else(educ == 'college', 1, 0),
         has_children = nchild > 0,
         white = ethnicity == 'White',
         black = ethnicity == 'Black',
         hispanic = ethnicity == 'Hispanic',
         asian = ethnicity == 'Asian',
         other = ethnicity == 'Multiracial or Other',
         Female = sex == 'female',
         Male = sex == 'male') %>% 
  group_by(group) %>%
  summarize(across(c(inctot, hs, college, hwsei, has_children, nchild, married,
                     white, black, hispanic, asian, other, age, Female, Male), 
            ~ survey_mean(.x, na.rm = T)))

write_csv(desc_tab_ind, here('geo/files', 'desc_tab_ind.csv'))

desc_tab_time <- all_geo_survey %>%
  group_by(group, year) %>%
  summarize(across(c(bachelors_puma, black_puma, hispanic_puma, inctot_puma,
                     owned_puma, age_puma, immigrant_puma,
                     poverty_100_puma, poverty_200_puma, hwsei_puma,
                     unemployed_puma, valueh_puma,
                     rent_puma, costelec_puma, density, state_policy,
                     lgbt_nonprofits, immigrant_nonprofits),
            ~ survey_mean(.x, na.rm = T))) %>%
  replace(. == 0, NA)


write_csv(desc_tab_time, here('geo/files', 'desc_tab_time.csv'))
```

```{r desc-ind}

# desc_tab_se <- immigrants_survey %>%
#   group_by(same_sex) %>%
#   summarize(across(c(bachelors_puma_se, black_puma_se, hispanic_puma_se, inctot_puma,
#                      log_inctot_puma_se, owned_puma_se, age_puma_se, immigrant_puma,
#                      poverty_100_puma_se, poverty_200_puma_se, hwsei_puma_se, 
#                      unemployed_puma_se, valueh_puma,
#                      rent_puma_se, costelec_puma_se, density, pctmetro),
#             ~ sqrt(sum(perwt*(.x^2), na.rm = T)/sum(perwt, na.rm = T)^2)
#   ))
            #~ sqrt(survey_mean(.x^2, na.rm = T)/n())))



desc_tab_ind <- read_csv(here('geo/files', 'desc_tab_ind.csv'))

FitFlextableToPage <- function(ft, pgwidth = 6){
  ft_out <- ft %>% flextable::autofit()
  ft_out <- flextable::width(ft_out, width = dim(ft_out)$widths*pgwidth /
                               (flextable::flextable_dim(ft_out)$widths))
  return(ft_out)
}

# desc_tab_ind %>%
#   rename_with(~ paste0(.x, '_coef'), !ends_with('_se')) %>%
#   pivot_longer(-group_coef, names_sep = '_')


desc_tab_ind %>%
  left_join(sample_size) %>%
  column_to_rownames('group') %>%
  t() %>%
  as.data.frame() %>%
  # mutate(`Different-sex, immigrant` = round(V1, 3), `Same-sex, immigrant` = round(V2, 3),
  #        `Same-sex, non-immigrant` = round(V3, 3)) %>%
  # select(different_sex, same_sex) %>%
  rownames_to_column('Variable') %>%
  # mutate(Variable = ifelse(str_detect(Variable, '_se'), Variable, paste0(Variable, '_coef'))) %>%
  # pivot_wider(names_from = Variable, names_sep = '_')
  filter(!str_detect(Variable, '_se')) %>%
  mutate(Variable = recode(Variable, !!!variable_names_swap)) %>%
  flextable() %>% 
  colformat_double(i = 16:17, digits = 0) %>%
  FitFlextableToPage() %>% # flextable::autofit() %>%
  flextable::set_caption('Mean values for individual variables for immigrants in different- or same-sex couples, 2008-2019 and 2021-2023. For same-sex couples, the married variable is coded as missing before 2012 (in this time period, all same-sex couples are designated as "unmarried parterns"). Differences in means between the three groups are significant in all cases.') 
  # kable(booktabs = T, linesep = '', caption = 'Mean values for geographic variables for immigrants in different- or same-sex couples, 2008-2022')

# desc_tab_ind |>
#   select(group, !ends_with("_se")) |>
#   pivot_longer(!group, values_to = "estimate") |>
#   mutate(estimate = as.character(signif(estimate, 2))) |>
#   left_join(
#     desc_tab_ind |>
#       select(group, ends_with("_se")) |>
#       pivot_longer(!group, values_to = "se") |>
#       mutate(se = paste0("(", as.character(signif(se, 2)), ")"),
#              name = str_remove(name, "_se$"))
#   ) |>
#   mutate(value = paste(estimate, se)) |>
#   select(group, name, value) |>
#   pivot_wider() |>
#   left_join(sample_size) %>%
#   column_to_rownames('group') %>%
#   t() %>%
#   as.data.frame() %>%
#   # mutate(`Different-sex, immigrant` = round(V1, 3), `Same-sex, immigrant` = round(V2, 3),
#   #        `Same-sex, non-immigrant` = round(V3, 3)) %>%
#   # select(different_sex, same_sex) %>%
#   rownames_to_column('Variable') %>%
#   # mutate(Variable = ifelse(str_detect(Variable, '_se'), Variable, paste0(Variable, '_coef'))) %>%
#   # pivot_wider(names_from = Variable, names_sep = '_')
#   mutate(Variable = ifelse(str_detect(Variable, '_se'), '', Variable)) |>
#   # filter(!str_detect(Variable, '_se')) %>%
#   mutate(Variable = recode(Variable, !!!variable_names_swap)) %>%
#   flextable() %>% 
#   colformat_double(i = 17:18, digits = 0) %>%
#   FitFlextableToPage() %>% # flextable::autofit() %>%
#   flextable::set_caption('Mean values for individual variables for immigrants in different- or same-sex couples, 2008-2019 and 2021-2023. For same-sex couples, the married variable is coded as missing before 2012 (in this time period, all same-sex couples are designated as "unmarried parterns").') 

```

\newpage

```{r desc-geo}

# desc_tab_se <- immigrants_survey %>%
#   group_by(same_sex) %>%
#   summarize(across(c(bachelors_puma_se, black_puma_se, hispanic_puma_se, inctot_puma,
#                      log_inctot_puma_se, owned_puma_se, age_puma_se, immigrant_puma,
#                      poverty_100_puma_se, poverty_200_puma_se, hwsei_puma_se, 
#                      unemployed_puma_se, valueh_puma,
#                      rent_puma_se, costelec_puma_se, density, pctmetro),
#             ~ sqrt(sum(perwt*(.x^2), na.rm = T)/sum(perwt, na.rm = T)^2)
#   ))
            #~ sqrt(survey_mean(.x^2, na.rm = T)/n())))



desc_tab <- read_csv(here('geo/files', 'desc_tab.csv'))

FitFlextableToPage <- function(ft, pgwidth = 6){
  ft_out <- ft %>% flextable::autofit()
  ft_out <- flextable::width(ft_out, width = dim(ft_out)$widths*pgwidth /
                               (flextable::flextable_dim(ft_out)$widths))
  return(ft_out)
}

desc_tab %>%
  select(group, inctot_puma, unemployed_puma, hwsei_puma, valueh_puma,
         bachelors_puma, black_puma, hispanic_puma, immigrant_puma, 
         density, state_policy, lgbt_nonprofits, immigrant_nonprofits) %>%
  left_join(sample_size) %>%
  column_to_rownames('group') %>%
  t() %>%
  as.data.frame() %>%
  # mutate(`Different-sex, immigrant` = round(V1, 3), `Same-sex, immigrant` = round(V2, 3),
  #        `Same-sex, non-immigrant` = round(V3, 3)) %>%
  # select(different_sex, same_sex) %>%
  rownames_to_column('Variable') %>%
  # filter(!str_detect(Variable, '_se')) %>%
  mutate(Variable = recode(Variable, !!!variable_names_swap)) %>%
  flextable() %>% 
  colformat_double(i = 13:14, digits = 0) %>%
  FitFlextableToPage() %>% # flextable::autofit() %>%
  flextable::set_caption('Mean values for geographic variables for immigrants in different- or same-sex couples, 2008-2019 and 2021-2023. Differences in means between the three groups are significant in all cases.') 
  # kable(booktabs = T, linesep = '', caption = 'Mean values for geographic variables for immigrants in different- or same-sex couples, 2008-2022')


```


\newpage

# Figures



```{r desc-ind-figures-calc}
income <- all_geo_survey %>%
  group_by(year, group) %>%
  summarize(var = survey_mean(inctot, na.rm = T)) %>%
  mutate(var_name = 'Annual Personal Income (Thousands)')

hs <- all_geo_survey %>%
  mutate(hs = if_else(educ %in% c('HS', 'some col', 'college'), 1, 0)) %>%
  group_by(year, group) %>%
  summarize(var = survey_mean(hs, na.rm = T)) %>%
  mutate(var_name = 'High School Completion')

college <- all_geo_survey %>%
  mutate(college = if_else(educ == 'college', 1, 0)) %>%
  group_by(year, group) %>%
  summarize(var = survey_mean(college, na.rm = T)) %>%
  mutate(var_name = 'College Completion')

hwsei <- all_geo_survey %>%
  group_by(year, group) %>%
  summarize(var = survey_mean(hwsei, na.rm = T)) %>%
  mutate(var_name = 'Hauser and Warren Socioeconomic Index')

has_children <- all_geo_survey %>%
  group_by(year, group) %>%
  mutate(has_children = nchild > 0) %>%
  summarize(var = survey_mean(has_children, na.rm = T)) %>%
  mutate(var_name = 'Have Children')

married <- all_geo_survey %>%
  group_by(year, group) %>%
  summarize(var = survey_mean(married, na.rm = T)) %>%
  mutate(var = ifelse(var == 0, NA, var)) %>%
  mutate(var_name = 'Married')

# age <- all_geo %>%
#   group_by(year, group) %>%
#   summarize(var = weighted.mean(age, perwt, na.rm = T)) %>%
#   mutate(var_name = 'Age')

bind_rows(income, hwsei, hs, college, has_children, married) |>
  write_csv(here('geo/files', 'ind_fig.csv'))
```



```{r desc-ind-figures, fig.cap = 'Individual characteristics over time for immigrants in different- and same-sex couples and non-immigrants in same-sex couples, based on American Community Survey data for 2008-2019 and 2021-2023. Estimates incorporate sampling weights, and ribbons represent 95-percent confidence intervals.'}
read_csv(here('geo/files', 'ind_fig.csv')) %>%
  mutate(var_name = factor(var_name, levels = unique({.$var_name}))) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = var, linetype = group, color = group, fill = group)) +
  geom_line() +
  geom_ribbon(aes(ymin = var - 1.96*var_se, 
                  ymax = var + 1.96*var_se, 
                  color = NULL), alpha = .2) +
  facet_wrap(~var_name, scales = "free", ncol = 2) +
  ylab('') +
  theme(legend.position="bottom",
        text = element_text(size = 9))
```

\newpage

```{r state-props-calc, eval = F}
state_tab <- immigrants_survey %>%
  group_by(state, year, same_sex) %>%
  summarize(survey_prop()) %>%
  filter(same_sex == T) %>%
  complete(state, year) %>%
  replace(is.na(.), 0) 
  
write_csv(state_tab, here('geo/files', 'state_tab.csv'))
```



```{r state-map, fig.cap = 'Percentage of cohabiting immigrants in same-sex couples in U.S. states, averaging over ACS survey years 2008-2019 and 2021-2023.'}
state_tab <- read_csv(here('geo/files', 'state_tab.csv')) %>%
  mutate(year_cat = case_when(
         year < 2013 ~ '2008-2012',
         year %in% 2013:2017 ~ '2013-2017',
         year > 2017 ~ '2018-2023'),
         state = tolower(state)) %>%
  filter(state != 'district of columbia') %>%
  group_by(year_cat, state) %>%
  summarize(prop_samesex = mean(coef)*100) 


state_tab %>%
  # mutate(prop_samesex = ifelse(prop_samesex > 4, 4, prop_samesex)) %>%
  ggplot(aes(map_id = state, fill = prop_samesex)) +
  geom_map(map = fiftystater::fifty_states) +
  expand_limits(x = fiftystater::fifty_states$long, y = fiftystater::fifty_states$lat) +
  coord_map() +
  # scale_fill_gradient(low = "#7570B3", high = "#1B9E77", na.value = 'gray95') +
  scale_fill_gradient(low = uclablue, high = uclagold, na.value = 'gray95') +
                      #breaks = c(0,7), labels=format(c(0,7))) +
  theme_void() + # font_family = 'Palatino') +
  facet_wrap(~year_cat, ncol = 1) +
  labs(fill = "Same-sex (%)") +
  # guides(color = guide_legend(override.aes = list(size = 0.5))) +
  theme(legend.title = element_text(size=8), legend.text=element_text(size=8))
  # theme(legend.title=element_blank(), 
  #       panel.grid.major.y = element_line('grey80'),
  #       legend.background = element_rect(fill = "transparent"))


# (state_tab %>%
#   filter(year_cat == '2008-2013') %>%
#   ggplot(aes(map_id = state, fill = prop_samesex)) +
#   geom_map(map = fiftystater::fifty_states) +
#   expand_limits(x = fiftystater::fifty_states$long, y = fiftystater::fifty_states$lat) +
#   coord_map() +
#   scale_fill_gradient(low = uclablue, high = uclagold, na.value = 'gray95') +
#                       #breaks = c(0,7), labels=format(c(0,7))) +
#   theme_map() + # font_family = 'Palatino') +
#   facet_wrap(~year_cat, ncol = 1) +
#   labs(fill = "Same-sex (%)")) /
#   (state_tab %>%
#   filter(year_cat == '2014-2017') %>%
#   ggplot(aes(map_id = state, fill = prop_samesex)) +
#   geom_map(map = fiftystater::fifty_states) +
#   expand_limits(x = fiftystater::fifty_states$long, y = fiftystater::fifty_states$lat) +
#   coord_map() +
#   scale_fill_gradient(low = uclablue, high = uclagold, na.value = 'gray95') +
#                       #breaks = c(0,7), labels=format(c(0,7))) +
#   theme_map() + # font_family = 'Palatino') +
#   facet_wrap(~year_cat, ncol = 1) +
#   labs(fill = "Same-sex (%)")) /
#   (state_tab %>%
#   filter(year_cat == '2019-2022') %>%
#   ggplot(aes(map_id = state, fill = prop_samesex)) +
#   geom_map(map = fiftystater::fifty_states) +
#   expand_limits(x = fiftystater::fifty_states$long, y = fiftystater::fifty_states$lat) +
#   coord_map() +
#   scale_fill_gradient(low = uclablue, high = uclagold, na.value = 'gray95') +
#                       #breaks = c(0,7), labels=format(c(0,7))) +
#   theme_map() + # font_family = 'Palatino') +
#   facet_wrap(~year_cat, ncol = 1) +
#   labs(fill = "Same-sex (%)"))



```

\newpage

```{r segregation, fig.cap = 'Index of dissimilarity by PUMA for 2008 to 2023, comparing immigrants in same-sex couples to immigrants in different-sex couples or to non-immigrants in same-sex couples.'}

seg %>%
  ggplot(aes(x = year, y = est, color = comparison, fill = comparison,
             linetype = comparison)) +
  geom_line() +
  # geom_ribbon(aes(ymin = est - 1.96*se, ymax = est + 1.96*se, color = NULL),
  #             alpha = 0.2) +
  geom_hline(yintercept = 0, color= 'transparent') +
  theme(legend.position = 'bottom') +
  labs(x = 'Year', y = 'Index of dissimilarity')


# bind_rows(yearly_theil, seg_imm, seg_lgbq) %>%
#   ggplot(aes(x = year, y = est, color = comparison, linetype = comparison)) +
#   geom_line() +
#   geom_hline(yintercept = 0) +
#   facet_wrap(~index, scales = 'free')
```


\newpage

```{r desc-geo-figures, fig.height = 6.5, fig.cap = 'Geographic characteristics over time for immigrants in different- and same-sex couples and non-immigrants in same-sex couples, based on American Community Survey data for 2008-2019 and 2021-2023. Estimates incorporate sampling weights, and ribbons represent 95-percent confidence intervals.'}

desc_tab_time <- read_csv(here('geo/files', 'desc_tab_time.csv')) 

# desc_tab_time %>%
#   select(!ends_with('_se')) %>%
#   select(group, year,
#          inctot_puma, unemployed_puma, hwsei_puma, valueh_puma,
#          bachelors_puma, black_puma, hispanic_puma, immigrant_puma, 
#          density, state_policy, lgbt_nonprofits, immigrant_nonprofits) %>%
#   pivot_longer(-c(group, year), names_to = 'variable') %>%
#   mutate(variable = recode(variable, !!!variable_names_swap)) %>%
#   mutate(variable = factor(variable, levels = unique({.$variable}))) %>%
#   # filter(str_detect(variable, '(%)')) %>%
#   # mutate(same_sex = ifelse(same_sex == T, 'Same-sex', 'Different-sex')) %>%
#   ggplot(aes(x = year, y = value, color = group, linetype = group)) +
#   geom_line() +
#   facet_wrap(~variable, scales = 'free_y', ncol = 2) +
#   labs(x = 'Year', y = '') +
#   theme(legend.position = 'bottom',
#         text = element_text(size = 9))


left_join(
  desc_tab_time %>%
    select(!ends_with('_se')) %>%
    pivot_longer(-c(group, year), names_to = 'variable', values_to = 'estimate'),
  desc_tab_time %>%
    select(group, year, ends_with('_se')) %>%
    rename_with(
    ~ str_remove(., "_se$"), ends_with("_se")) |>
    pivot_longer(-c(group, year), names_to = 'variable', values_to = 'se')
) |>
  mutate(lower = estimate - 1.96*se, upper = estimate + 1.96*se) |>
  mutate(variable = recode(variable, !!!variable_names_swap)) %>%
  mutate(variable = factor(variable, levels = unique({.$variable}))) %>%
  ggplot(aes(x = year, y = estimate, color = group, linetype = group)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = group, color = NULL),
              alpha = .2) +
  facet_wrap(~variable, scales = 'free_y', ncol = 2) +
  labs(x = 'Year', y = '') +
  theme(legend.position = 'bottom',
        text = element_text(size = 9))
  


```




