---
output:
  beamer_presentation:
    keep_tex: yes
    incremental: no
    toc: no
    latex_engine: pdflatex
    slide_level: 2
    include:
      in_header: ucla_beamer.tex
    fig_width: 12.5
    

# header-includes:
#   - \AtBeginDocument{\title[Making Migration Sexy]{Making Migration Sexy: How State and National Policies Influence Migration of Same-Sex Couples}}
#   -\AtBeginDocument{\author[Hoffmann \& Velasco]{ Nathan I. Hoffmann, Sociology, UCLA
#  \\ Kristopher Velasco, Sociology, Princeton}}


title: "Making Migration Sexy: How State and National Policies Influence International Migration of Same-Sex Couples"
subtitle: "ASA 2021 Annual Meeting"


author: |
  | Nathan I. Hoffmann, Department of Sociology, UCLA
  | Kristopher Velasco, Depatment of Sociology, Princeton

date: "UCLA & Princeton"
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, results = 'asis',
                      cache.lazy = F)

library(MASS)
library(janitor)
library(priceR)
library(kableExtra)
library(huxtable)
library(jtools)
library(sjstats)
library(imputeTS)
library(mvtnorm)
library(ordinal)
library(lme4)
library(countrycode)
library(knitr)
library(RColorBrewer)
library(haven)
library(cowplot)
library(srvyr)
library(lmtest)
library(sandwich)
library(here)
library(plm)
library(stargazer)
library(tidyverse)

# knitr::opts_chunk$set(error=TRUE) 

options("yaml.eval.expr" = TRUE)

options(scipen=1, digits=2)
knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")
})
knitr::opts_chunk$set(error=TRUE) 

uclablue = '#2774AE'
uclagold = '#FFB81C'
ucladarkblue = '#003B5C'
uclalightblue = '#8BB8E8'
gray = '#808080'
black = '#000000'
blue = "#0000FF"
ucla_palette = c(uclablue, black, gray)

theme_set(theme_classic(base_size = 21) + 
            theme(legend.title=element_blank(), 
                  panel.grid.major.y = element_line('grey80'),
                  legend.background = element_rect(fill = "transparent"),
                  rect = element_rect(fill = "transparent"),
                  plot.background = element_rect(fill = "transparent", color = NA),
                  panel.background = element_rect(fill = "transparent", color = NA)))
ggplot <- function(...) ggplot2::ggplot(...) + scale_color_manual(values = ucla_palette) +
  scale_fill_manual(values = ucla_palette)
# ggplot <- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Dark2") + 
#   scale_fill_brewer(palette="Dark2")
```

```{r load, include = F}

acs_wide <- read.csv(here('data', 'acs_wide.csv'))
acs_couple_policy <- readRDS(here('data', 'acs_couple_policy.rds')) 
acs_prop_yrimmig_policy <- read.csv(here('data', 'acs_prop_yrimmig_policy.csv')) %>%
  mutate(post_2013 = (yrimmig > 2013))
acs_dyad_policy <- read.csv(here('data', 'acs_dyad_policy.csv'))

# acs_oneimm <- read.csv(here('data', 'acs_oneimm.csv'))
# acs_coupled_imms <- read.csv(here('data', 'acs_coupled_imms.csv'))
# acs_prop_yrimmig <- read.csv(here('data', 'acs_prop_yrimmig.csv'))
# acs_dyad <- read.csv(here('data', 'acs_dyad.csv'))


acs_wide_survey <- acs_wide %>% as_survey_design(ids = cluster, weights = hhwt)
acs_ind_survey <- acs_couple_policy %>% as_survey_design(ids = cluster, weights = perwt)

# Smaller dataset for initial ordered logit analyses
set.seed(1859)
acs_couple_policy_small <- bind_rows(
  filter(acs_couple_policy, same_sex == T, yrimmig >= 1991),
  sample_n(filter(acs_couple_policy, same_sex == F), 1e5)
) 

total_df <- acs_wide_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, imm_couple, same_sex) %>%
  summarize(n = survey_total(vartype = "ci")) 
```

# Introduction
## Introduction
- In 2013, the U.S. Supreme Court struck down the Defense of Marriage Act
- From 2013 to 2019, 76% increase in immigrant same-sex couples, compared to 13% for different-sex
- LGB policy environments, both at country of origin and in U.S. states, are changing rapidly and yet, little is understood how this contributes toward the migration of immigrants in same-sex couples
- More broadly, how can understanding LGB migration help us better understand the links between migration, policy, and identity more broadly?

## Introduction
```{r total-pop}
total_df <- acs_wide_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-sex immigrant couples', 'Different-Sex immigrant couples')) %>%
  group_by(year, imm_couple, same_sex) %>%
  summarize(n = survey_total(vartype = "ci")) 

total_df %>%
  mutate(`Total population (thousands)` = n/1000,
         n_low = n_low/1000, n_upp = n_upp/1000,
         Year = year) %>%
  ggplot(aes(x = Year, y = `Total population (thousands)`, 
             color = imm_couple)) +
  geom_line() +
  geom_ribbon(aes(ymin = n_low, ymax = n_upp, fill = imm_couple, col = NULL), alpha = .2) +
  geom_vline(xintercept = 2013, linetype = 'dashed') +
  facet_wrap(~same_sex, scales = "free") +
  xlim(2007, 2019) +
  theme(legend.justification=c(1,0),
        legend.position=c(.15,.35))

```

# Background
## Conventional Explanations
- Migration theory relies heavily on economic and network theories
- Recently, culture and social policy considered
- Little previous research applying migration theories to same-sex couples. Do they hold up?


## Our Intervention
- Incorporate an aspirations-capabilities framework (de Haas 2021)
- We argue that the state's regulation of sexuality influences migratory patterns through altering the aspirations and capabilities to migrate
- This population is particularly sensitive to changing policy
    + Same-sex relationships were not recognized by the U.S. government before the 2013 DOMA decision
    + Growing awareness of LGBTQ+ rights within global discourses increases the salience of considering such policies to motivate migration
    + Policy may overpower traditional migration theories


## Our Intervention
- We consider policy at both country of origin and U.S. state (FitzGerald 2008)
- Country of origin
    + Do LGB people choose to flee repressive policy contexts?
      + Qualitative approaches to queer migration identify this (Carrillo 2017; Vogler 2016)
    + Or does progressive policy create the capacity for migration?
- U.S. state
    + Do immigrants in same-sex couples, like their U.S.-born counterparts, choose to live in more progressive states?


# Data
## Identifying Same-Sex Couples in the ACS
- 2008 to 2019 American Community Survey (ACS)
    + immigrated at age 18 or older post-1990
- Immigrants in same-sex couples are identified as foreign-born respondents who live with a same-sex married or unmarried partner
    + This necessarily excludes single and non-cohabiting LGB individuals
- Sample of 7,011 immigrants in same-sex couples compared to 898,869 immigrants in different-sex couples


## Variables
- Explanatory variables
    + Country of origin LGBT policy index (sum of 14 policies) (Velasco 2020)
    + U.S. state LGBT policy index (sum of 8 policies) (Movement Advancement Project 2020)
- Controls
    + Factors from standard migration models, including country- and state-level economic, political, and demographic variables from the UN, U.S. government, and other sources
    + The relative size of the co-national immigrant population as a proxy for immigrant networks
    + Individual sociodemographic variables from ACS



# Methods
## Methods
- **Analysis 1**: Country-level percentage of immigrants in same-sex couples, by country and year of immigration
    + OLS regression with country fixed effects
- **Analysis 2**: State-level percentage of immigrants in same-sex couples, by country and year of immigration
    + OLS regression with state and country fixed effects
- **Analysis 3**: Individual-level models predicting state policy environment
    + Ordered logistic regression with survey-year fixed effects
    + See paper for more details.


# Results
## Results
\begin{center}
  \huge{\textcolor{uclablue}{Results}}
\end{center}
  

## Descriptive Statistics
```{r desc}
income <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(inctot, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Individual Income')

prop_stock <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(stock_prop, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Proportion Immigrant Stock in Year of Migration')

wage_dif <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(wage_dif, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'Bilateral Per-Capita GDP Differential in Year of Migration')

vdem <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(vdem, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'V-Dem Score in Year of Migration')


desc_plot <- function(df, legend_pos = c(.8, .8), ylab = ''){
  df %>%
    rename(Year = year) %>%
    ggplot(aes(x = Year, y = var, color = same_sex)) +
    geom_line() +
    geom_ribbon(aes(ymin = var_low, ymax = var_upp, fill = same_sex, col = NULL), alpha = .2) +
    facet_wrap(~var_name) +
    ylab(ylab) +
    theme(legend.justification=c(1,0),
        legend.position= legend_pos,
        legend.background = element_rect(fill = alpha("white", 0.5)))
}

desc_plot(income, legend_pos = c(.5,.82), ylab = '1999 U.S. dollars (thousands)')

# bind_rows(income, wage_dif, vdem, prop_stock) %>%
#   rename(Year = year) %>%
#   ggplot(aes(x = Year, y = var, color = same_sex)) +
#   geom_line() +
#   geom_ribbon(aes(ymin = var_low, ymax = var_upp, fill = same_sex, col = NULL), alpha = .2) +
#   facet_wrap(~var_name, scales = "free") +
#   ylab('') +
#   theme(legend.justification=c(1,0),
#       legend.position=c(.22,.82),
#       legend.background = element_rect(fill = alpha("white", 0.5))) 
```


## Descriptive Statistics
```{r desc2}
desc_plot(wage_dif, c(.3, .83), ylab = '1999 U.S. dollars (thousands)')
```

## Descriptive Statistics
```{r desc3}
desc_plot(vdem, c(.9, .8))
```

## Descriptive Statistics
```{r desc4}
desc_plot(prop_stock, c(.8, .76))
```




## Descriptive Statistics
```{r policy-desc}
origin_df <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(`Policy Score` = survey_mean(origin_score, vartype = "ci", na.rm = T)) %>%
  mutate(var = 'Country-of-Origin Index')

state_df <- acs_ind_survey %>%
  mutate(same_sex = ifelse(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(`Policy Score` = survey_mean(state_policy, vartype = "ci", na.rm = T)) %>%
  mutate(var = 'U.S. State Index')

bind_rows(origin_df, state_df) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = `Policy Score`, 
             color = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = `Policy Score_low`, ymax = `Policy Score_upp`, 
                  fill = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var, scales = "free") +
  theme(legend.justification=c(1,0),
        legend.position=c(.25,.65))
```






## Model Results: Country-of-Origin Effects (Table 3)
```{r country-coefs}
ag_country_orig_fe <- lm(prop_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + vdem + origin_score + stock_prop + bpld,
         data = acs_prop_yrimmig_policy)

ag_2013_int_fe <- lm(prop_same_sex ~ distw + contig + comlang_off + comlang_ethno + colony +
                   wage_dif + unemp_dif + vdem  + stock_prop + bpld +
                     post_2013*origin_score,
         data = acs_prop_yrimmig_policy)

coef_list <- list()
mod_list <- list(ag_country_orig_fe,  ag_2013_int_fe)
mod_names <- c('3: Country controls and FEs', '5: Controls + 2013 Interaction')
for(mod_num in 1:length(mod_list)){
  coef_list[[mod_num]]  <- coef(summary(mod_list[[mod_num]], cluster = c("bpld"))) %>%
    as.data.frame() %>%
    rownames_to_column('variable') %>%
    mutate(lower = Estimate - 1.96*`Std. Error`, 
           upper = Estimate + 1.96*`Std. Error`,
           model = mod_names[mod_num])
}

coef_df <- bind_rows(coef_list) %>%
  filter(variable %in% c('origin_score', 'post_2013TRUE', 'post_2013TRUE:origin_score')) %>%
  mutate(variable = 
           case_when(
              variable == 'origin_score' ~ 'Country LGB policy score',
              variable == 'post_2013TRUE' ~ 'Post-2013',
              variable == 'post_2013TRUE:origin_score' ~ 'Country score × Post-2013'),
         variable = factor(variable, 
                           levels = c('Country score × Post-2013', 'Post-2013', 'Country LGB policy score'))
         )

ggplot(coef_df, aes(x = variable, y = Estimate, 
                    color = ifelse(startsWith(model, '3'), 'black', 'grey'))) +
  geom_point(position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .1, alpha = 1, position = position_dodge(0.2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  facet_wrap(~model) +
  # scale_y_continuous(breaks = c(seq(-1.00,.50,.20))) +
  # scale_y_continuous(breaks = c(seq(-100,50,20))) +
  labs(x = '', y = 'Effect on percentage in same-sex couples') +
  theme(panel.grid.major.y = element_line('transparent')) +
  scale_color_identity()

```

## Model Results: Country-of-Origin Effects (Table 3)
```{r country-coefs2}
ggplot(coef_df, aes(x = variable, y = Estimate, 
                    color = ifelse(startsWith(model, '5'), 'black', 'grey'))) +
  geom_point(position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .1, alpha = 1, position = position_dodge(0.2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  facet_wrap(~model) +
  # scale_y_continuous(breaks = c(seq(-1.00,.50,.20))) +
  # scale_y_continuous(breaks = c(seq(-100,50,20))) +
  labs(x = '', y = 'Effect on percentage in same-sex couples') +
  theme(panel.grid.major.y = element_line('transparent')) +
  scale_color_identity()

```

<!-- ## Model Results: Country-of-Origin Effects (Table 3) -->
```{r country-sim, eval = F}
Mode <- function(x) {
 ux <- unique(x)
 ux[which.max(tabulate(match(x, ux)))]
}

predict.rob <- function(x,clcov,newdata){
  if(missing(newdata)){ newdata <- x$model }
  tt <- terms(x)
  Terms <- delete.response(tt)
  m.mat <- model.matrix(Terms, data=newdata)
  m.coef <- x$coef
  fit <- as.vector(m.mat %*% as.matrix(x$coef))
  se.fit <- sqrt(diag(m.mat%*%clcov%*%t(m.mat)))
  return(list(fit=fit,se.fit=se.fit))
  }

# Jamaica (-1, rank 12), Mexico (3.5, rank 1), France (6.3, rank 33)

sim_dat1 <- acs_prop_yrimmig_policy %>%
  group_by(yrimmig) %>%
  summarize(across(c(comlang_off, comlang_ethno, colony, contig), Mode),
            across(c(distw, wage_dif, unemp_dif, vdem, stock_prop), mean, na.rm = T))


sim_dat_j <- sim_dat1 %>%
  mutate(bpld = 'Jamaica', origin_score = -1)


sim_dat_m <- sim_dat1 %>%
  mutate(bpld = 'Mexico', origin_score = 3.5)

sim_dat_f <- sim_dat1 %>%
  mutate(bpld = 'France', origin_score = 6.3)

sim_dat <- bind_rows(sim_dat_j, sim_dat_m, sim_dat_f) %>%
           mutate(post_2013 = (yrimmig > 2013),
                  bpld = factor(bpld, levels = unique(acs_prop_yrimmig_policy$bpld)))

cbind(sim_dat,
          predict(ag_2013_int_fe, newdata = sim_dat, interval = 'confidence')) %>%
  ggplot(aes(x = yrimmig, y = fit, color = bpld)) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr, ymax = upr, fill = bpld, col = NULL), alpha = .2)



# predict.rob(x = ag_2013_int_fe, 
#             clcov = vcovCL(ag_2013_int_fe, cluster = ~ bpld),
#             newdata = sim_dat)

sim <- list()
for(var in c('educ_short', 'imm15', 'int_lang')){
 sim_dat_loop <- bind_rows(rep(list(select(sim_dat, -!!sym(var))),
                               length(unique(pull(pew, !!sym(var))))))
 sim_dat_loop[, var] <- rep(unique(pull(pew, !!sym(var))), each = 3)
 sim_dat_loop <- sim_dat_loop %>%
   mutate(variable = var,
          value = as.character(!!sym(var)))
 sim_dat_loop <- bind_cols(sim_dat_loop,
                           predict(ind_int,
                                   newdata = sim_dat_loop,
                                   se.fit = T)) %>%
   mutate(fit.low = fit - 1.96 * se.fit,
          fit.high = fit + 1.96 * se.fit) 
 sim[[var]] <- sim_dat_loop
}
sim <- bind_rows(sim) %>%
 mutate(value = case_when(value == '0' ~ 'No',
                          value == '1' ~ 'Yes',
                          T ~ value),
        variable = case_when(variable == 'educ_short' ~ 'Education',
                             variable == 'imm15' ~ '1.5 Generation',
                             variable == 'int_lang' ~ 'Spanish Interview'),
 across(c(fit, fit.low, fit.high), sigmoid))













## Simulation
set.seed(185)
sims <- 10
sim <- list()
sim_mod <- list(sims)
pseudo_mod <- ag_2013_int_fe
k <- length(coef(pseudo_mod))
theta_mat <- MASS::mvrnorm(sims, coef(pseudo_mod), vcovCL(pseudo_mod, cluster = ~ bpld))
var1_vals <- unique(acs_prop_yrimmig_policy$yrimmig)






# acs_prop_yrimmig_policy %>%
#   group_by(bpld) %>%
#   summarize(origin_score = mean(origin_score),
#             stock_prop = mean(stock_prop)) %>%
#   arrange(desc(stock_prop)) %>%
#   as.data.frame()



for (i in 1:sims) {
  if(i %% 10 == 0) print(i)
  theta <- as.numeric(theta_mat[i,])
  pseudo_mod$coefficients <- theta[1:k]
  
  for(var1 in unique(acs_prop_yrimmig_policy$yrimmig)){
    for(val2 in range(acs_prop_yrimmig_policy$origin_score)){
        sim_dat <- acs_prop_yrimmig_policy %>%
          select(all.vars(formula(ag_2013_int_fe)))
        sim_dat[, 'yrimmig'] <- val1
        sim_dat[, var2] <- val2

        pred_dat <- as.numeric(predict(pseudo_mod, newdata = sim_dat))

       sim[[paste0(var,var1, var2, i)]] <-
        data.frame(
          fit = mean(pred_dat),
                          yrimmig = val1,
                          origin_score = as.character(val2))
    }
  }
}

sim_df <- bind_rows(sim) %>%
  group_by(legalstat_imp, variable, value) %>%
  summarize(fit.low = quantile(fit, .025), fit.high = quantile(fit, .975), fit = mean(fit)) %>%
  mutate(across(c(fit, fit.low, fit.high), sigmoid)) %>%
  mutate(value = case_when(value == '0' ~ 'No',
                           value == '1' ~ 'Yes',
                           T ~ value),
         variable = case_when(variable == 'educ_short' ~ 'Education',
                              variable == 'imm15' ~ '1.5 Generation',
                              variable == 'int_lang' ~ 'Interview Language'),
         `Legal status` = legalstat_imp)


write_csv(sim_df, 'ind_sim.csv')

```



## Model Results: State Effects (Table 4)
```{r state-coefs}
state_prop_fe <- lm(same_prop ~ state_policy + origin_score + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = acs_dyad_policy)

state_prop_fe2 <- lm(same_prop ~ state_policy*origin_score + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = acs_dyad_policy)
state_prop_2013_int <- lm(same_prop ~ post_2013*origin_score + state_policy + distw + contig + 
                          comlang_off + comlang_ethno + colony +
                          wage_dif + unemp_dif + vdem + stock_prop +
                           state_unemploy + state_income + Country + state, 
                       data = mutate(acs_dyad_policy, post_2013 = (year > 2013)))

coef_list <- list()
mod_list <- list(state_prop_fe,  state_prop_fe2, state_prop_2013_int)
mod_names <- c('3: Country and state controls + FEs', 
               '4: Controls + Interaction' , 
               '7: Controls + 2013 Interaction')
for(mod_num in 1:length(mod_list)){
  coef_list[[mod_num]]  <- coef(summary(mod_list[[mod_num]], cluster = c("state", "Country"))) %>%
    as.data.frame() %>%
    rownames_to_column('variable') %>%
    mutate(lower = Estimate - 1.96*`Std. Error`, 
           upper = Estimate + 1.96*`Std. Error`,
           model = mod_names[mod_num])
}

coef_df <- bind_rows(coef_list) %>%
  filter(variable %in% c('state_policy', 'origin_score', 
                                'state_policy:origin_score', 'post_2013TRUE', 
                                'post_2013TRUE:state_policy', 
                                'post_2013TRUE:origin_score')) %>%
  mutate(variable = 
           case_when(
              variable == 'origin_score' ~ 'Country LGB policy score',
              variable == 'post_2013TRUE' ~ 'Post-2013',
              variable == 'post_2013TRUE:origin_score' ~ 'Country score × Post-2013',
              variable == 'state_policy' ~ 'State LGB policy score',
              variable == 'state_policy:origin_score' ~ 'State score × country-score'),
         variable = factor(variable, 
                           levels = rev(c('State LGB policy score', 'Country LGB policy score',
                                'State score × country-score', 'Post-2013', 
                               'Country score × Post-2013')))
         )

ggplot(coef_df, aes(x = variable, y = Estimate,
                    color = ifelse(startsWith(model, '3'), 'black', 'grey'))) +
  geom_point(position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .1, alpha = 1, position = position_dodge(0.2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  facet_wrap(~model, labeller = label_wrap_gen(width=20)) +
  # scale_y_continuous(breaks = c(seq(-1.00,.50,.20))) +
  # scale_y_continuous(breaks = c(seq(-100,50,20))) +
  labs(x = '', y = 'Effect on percentage in same-sex couples') +
  theme(panel.grid.major.y = element_line('transparent')) +
  scale_color_identity()
```

## Model Results: State Effects (Table 4)
```{r state-coefs2}
ggplot(coef_df, aes(x = variable, y = Estimate,
                    color = ifelse(startsWith(model, '4'), 'black', 'grey'))) +
  geom_point(position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .1, alpha = 1, position = position_dodge(0.2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  facet_wrap(~model, labeller = label_wrap_gen(width=20)) +
  # scale_y_continuous(breaks = c(seq(-1.00,.50,.20))) +
  # scale_y_continuous(breaks = c(seq(-100,50,20))) +
  labs(x = '', y = 'Effect on percentage in same-sex couples') +
  theme(panel.grid.major.y = element_line('transparent')) +
  scale_color_identity()
```

## Model Results: State Effects (Table 4)
```{r state-coefs3}
ggplot(coef_df, aes(x = variable, y = Estimate,
                    color = ifelse(startsWith(model, '7'), 'black', 'grey'))) +
  geom_point(position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .1, alpha = 1, position = position_dodge(0.2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  facet_wrap(~model, labeller = label_wrap_gen(width=20)) +
  # scale_y_continuous(breaks = c(seq(-1.00,.50,.20))) +
  # scale_y_continuous(breaks = c(seq(-100,50,20))) +
  labs(x = '', y = 'Effect on percentage in same-sex couples') +
  theme(panel.grid.major.y = element_line('transparent')) +
  scale_color_identity()
```


<!-- ## Model Results: Individual Analysis -->
<!-- - Ordered logit to predict three-category LGB policy environment at the U.S. state level corroborate findings from previous models -->
<!-- - At the individual level, immigrants in same-sex couples are more likely to live in progressive states, net of individual characteristics -->
<!-- - Interaction between country of origin and same-sex partnership: LGB immigrants from more progressive origin countries tend to live in more progressive U.S. states -->
<!--     + For immigrants in different-sex couples, interaction term is insignificant -->


# Discussion
## Discussion
- Immigrants in same-sex couples have higher incomes, occupational prestige, and education than those in different-sex couples, and they come from countries with smaller wage and unemployment gaps with the U.S.
- Progressive countries send higher proportions of immigrants in same-sex couples
    + Contrary to existing, mostly qualitative scholarship on queer migration 
    + Patterns hold across one- and two-immigrant couples.
- Immigrants in same-sex couples reside in more progressive U.S. states

## Discussion
- Policies not explicitly related to migration may shape migration flows
- Importance of migration scholars studying the state's governance of identity
- Need to move beyond the traditional economic and network explanations of migration
- Incorporating an aspirations-capabilities framework is fruitful for migration research (de Haas 2021)



## 
\begin{center}
  \huge{\textcolor{uclablue}{Thank You}}
\end{center}

- Nathan I. Hoffmann ([nathanihoff\@ucla.edu](mailto:nathanihoff@ucla.edu))
- Kristopher Velasco ([kvelasco\@princeton.edu](mailto:kvelasco\@princeton.edu))
- Full paper on SocArXiv: https://tinyurl.com/hoffmann-velasco/



``` {=latex}
\end{frame}
\appendix
\begin{frame}<0| handout:0>
```



# Supplemental material

## Descriptive Statistics
```{r country-tab}
country_tab <- acs_couple_policy %>%
  group_by(bpld, same_sex) %>%
  count(wt = perwt) %>%
  pivot_wider(names_from = 'same_sex', values_from = 'n') %>%
  mutate(prop_samesex = round(100*`TRUE` / `FALSE`, 2)) %>%
  arrange(desc(prop_samesex)) %>%
  .[1:10,] %>%
   mutate(prop_samesex = case_when(
    nchar(prop_samesex) == 4 ~ paste0(prop_samesex, ' %'),
    nchar(prop_samesex) == 3 ~ paste0(prop_samesex, '0 %'),
    nchar(prop_samesex) == 1 ~ paste0(prop_samesex, '.00 %'))) %>%
  rownames_to_column() %>%
  left_join(acs_couple_policy %>%
              group_by(bpld) %>%
              summarize(`Mean policy score` = 
                          weighted.mean(origin_score, na.rm = T, w = perwt))) %>%
  select(Rank = rowname, `Country of origin` = bpld, `Proportion same-sex` = prop_samesex,
         `Mean policy score`)

country_tab %>%
  kable(booktabs = T, 
        linesep = '',
        caption = 'Sending countries ranked by proportion immigrant couples with same-sex partners') %>%
      footnote(general = "American Community Survey 2008-2019. Authors\' calculations.",
           general_title = 'Source:',
           footnote_as_chunk = T) %>%
  kable_styling(font_size = 9)


country_vec <- country_tab$`Mean policy score`
names(country_vec) <- country_tab$`Country of origin`


# acs_couple_policy %>%
#   filter(year == 2008) %>%
#   group_by(bpld, same_sex) %>%
#   count(wt = perwt) %>%
#   pivot_wider(names_from = 'same_sex', values_from = 'n') %>%
#   arrange(desc(`TRUE`)) %>%
#   .[1:10,] %>%
#   rownames_to_column() %>%
#   left_join(acs_couple_policy %>%
#               filter(year == 2008) %>%
#               group_by(bpld) %>%
#               summarize(origin_score = mean(origin_score)) %>%
#               select(bpld , origin_score))

```

## Descriptive Statistics
```{r state-tab}
state_tab <- acs_wide %>%
  filter(state != 'District of Columbia') %>%
  group_by(state, same_sex) %>%
  count(wt = hhwt) %>%
  pivot_wider(names_from = 'same_sex', values_from = 'n') %>%
  mutate(prop_samesex = round(100*`TRUE` / `FALSE`, 2)) %>%
  arrange(desc(prop_samesex)) %>%
  .[1:10,] %>%
  mutate(prop_samesex = case_when(
    nchar(prop_samesex) == 4 ~ paste0(prop_samesex, ' %'),
    nchar(prop_samesex) == 3 ~ paste0(prop_samesex, '0 %'),
    nchar(prop_samesex) == 1 ~ paste0(prop_samesex, '.00 %'))) %>%
  left_join(acs_couple_policy %>%
              group_by(state) %>%
              summarize(`Mean policy score` = 
                          weighted.mean(state_policy, na.rm = T, w = perwt))) %>%
  rownames_to_column() %>%
  select(Rank = rowname, `State` = state, `Proportion same-sex` = prop_samesex,
         `Mean policy score`)

state_tab %>%
  kable(booktabs = T, 
        linesep = '',
        caption = 'States ranked by proportion immigrant couples with same-sex partners') %>%
      footnote(general = "American Community Survey 2008-2019. Authors\' calculations.",
           general_title = 'Source:',
           footnote_as_chunk = T)  %>%
  kable_styling(font_size = 9)

state_vec <- state_tab$`Mean policy score`
names(state_vec) <- state_tab$`State`
```

## Model Results: Individual Analysis (Table 5)
```{r ind-coefs}
ord_interact_controls <- polr(state_policy_binned ~ 
                                same_sex*(origin_score + gender + age + educ +
                     nchild + ihs_income + no_income + yrimmig) +
                     distw + contig + comlang_off + comlang_ethno +
                     wage_dif + unemp_dif + vdem + stock_prop + #  bpld + state +
                     state_unemploy + state_income,
                data = rename(acs_couple_policy_small, gender = sex), Hess = T, weights = perwt/12)



coef_df <- coef(summary(ord_interact_controls, cluster =c('bpld', 'state'))) %>%
    as.data.frame() %>%
    rownames_to_column('variable') %>%
    mutate(lower = Value - 1.96*`Std. Error`, 
           upper = Value + 1.96*`Std. Error`,
           model = c('4: Individual and country controls + Interaction')) %>%
  filter(variable %in% c('same_sexTRUE', 'origin_score', 'same_sexTRUE:origin_score')) %>%
  mutate(variable = 
           case_when(
              variable == 'origin_score' ~ 'Country LGB policy score',
              variable == 'same_sexTRUE' ~ 'Same-sex',
              variable == 'same_sexTRUE:origin_score' ~ 'Same-sex × country score'),
         variable = factor(variable, 
                           levels = c('Same-sex × country score', 
                                      'Country LGB policy score', 
                                      'Same-sex'))
         )

ggplot(coef_df, aes(x = variable, y = Value)) +
  geom_point(position = position_dodge(0.2)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .1, alpha = 1, position = position_dodge(0.2)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() +
  facet_wrap(~model) +
  # scale_y_continuous(breaks = c(seq(-1.00,.50,.20))) +
  # scale_y_continuous(breaks = c(seq(-100,50,20))) +
  labs(x = '', y = 'Effect on 3-category state LGB policy score') +
  theme(panel.grid.major.y = element_line('transparent'))

```