---
output:
  beamer_presentation:
    keep_tex: no
    incremental: no
    toc: no
    latex_engine: pdflatex
    slide_level: 2
    include:
      in_header: ucla_beamer.tex
    fig_width: 12.5
    

# header-includes:
#   - \AtBeginDocument{\title[Mixed-Citizenship, Same-Sex Unions]{Policy Effects on Mixed-Citizenship, Same-Sex Unions: A Triple DIfferences Analysis}}
#   -\AtBeginDocument{\author[Hoffmann \& Velasco]{ Nathan I. Hoffmann, Sociology, UCLA
#  \\ Kristopher Velasco, Sociology, Princeton}}


title: "Policy Effects on Mixed-Citizenship, Same-Sex Unions: A Triple-Difference Analysis"
subtitle: "ASA 2022 Annual Meeting"


author: |
  | Nathan I. Hoffmann, Department of Sociology, UCLA
  | Kristopher Velasco, Depatment of Sociology, Princeton

date: "UCLA & Princeton"
---

```{r setup, include=F}
library(knitr)
library(broom)
library(srvyr)
library(huxtable)
library(sandwich)
library(lmtest)
library(here)
library(stargazer)
library(patchwork)
library(flextable)
library(tidyverse)
library(kableExtra)

knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = T, # results = 'asis',
                      cache.lazy = F, dpi = 600)

# knitr::opts_chunk$set(error=TRUE) 

options("yaml.eval.expr" = TRUE)

uclablue = '#2774AE'
gray = '#808080'
black = '#000000'
uclagold = '#FFB81C'
ucla_palette = c(uclablue, black, gray)

# theme_set(theme_cowplot(font_family = 'Palatino') + 
theme_set(theme_classic(base_size = 21) + 
            theme(legend.title=element_blank(), 
                  panel.grid.major.y = element_line('grey80'),
                  legend.background = element_rect(fill = "transparent"),
                  rect = element_rect(fill = "transparent"),
                  plot.background = element_rect(fill = "transparent", color = NA),
                  panel.background = element_rect(fill = "transparent", color = NA)))

# ggplot <- function(...) ggplot2::ggplot(...) + scale_color_brewer(palette="Dark2") +
#   scale_fill_brewer(palette="Dark2")
ggplot <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = ucla_palette) +
  scale_fill_manual(values = ucla_palette)

update_geom_defaults("line", list(size = 1.75))
update_geom_defaults("pointrange", list(size = 1.3))



# library(colorspace)
# colorspace_palette = 'RdPu'
# ggplot <- function(...) ggplot2::ggplot(...) +
#   scale_color_discrete_sequential(palette=colorspace_palette) +
#   scale_fill_discrete_sequential(palette=colorspace_palette)

options(scipen=999, digits=2)
knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")
})
knitr::opts_chunk$set(error=TRUE) 

set_flextable_defaults(font.family = 'Times New Roman',
                       # padding.bottom = 1,
                       # padding.top = 1,
                       # padding.left = 3,
                       # padding.right = 1,
                       theme_fun = 'theme_booktabs',
                       digits = 3, 
                       big.mark = ',',
                       line_spacing = 1)

defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

```{r load, include = F}
acs_count_base <- read_csv(here('data', 'acs_count_base.csv'))

count_func <- function(condition = '!is.na(state)'){
  acs_count_base %>%
    rename(mixed = mixed_citizenship) %>%
    filter(eval(rlang::parse_expr(condition))) %>%
    group_by(state, year, same_sex, mixed) %>%
    summarize(n = sum(perwt), 
              n_unweighted = n(),
              n_ar = sum(perwt[qcitizen != 'Allocated' & qeduc != 'Consistency edit' & allocated == F]), 
              n_unweighted_ar = sum(qcitizen != 'Allocated' & qeduc != 'Consistency edit' & allocated == F)) %>%
    ungroup() %>%
    complete(state, year, same_sex, mixed) %>%
    mutate(across(n:n_unweighted_ar, function(x) ifelse(is.na(x), 0, x))) %>%
    mutate(post_2013 = year > 2013,
           group_fe = paste(state, same_sex, mixed, sep = '_'))
}

# acs_count_base %>%
#   filter(state == 'Alabama', year == 2008) %>%
#   group_by(same_sex, mixed_citizenship, bpldnew) %>%
#     summarize(n = sum(perwt), 
#               n_unweighted = n(),
#               n_ar = sum(perwt[qcitizen != 'Allocated' & qeduc != 'Consistency edit' & allocated == F]), 
#               n_unweighted_ar = sum(qcitizen != 'Allocated' & qeduc != 'Consistency edit' & allocated == F)) %>%
#   group_by(same_sex, mixed_citizenship) %>%
#    summarize(n = sum(n),
#             n_unweighted = sum(n_unweighted),
#             n_ar = sum(n_ar),
#             n_unweighted_ar = sum(n_unweighted_ar))

acs_count_mixed <- count_func()

# acs_count_mixed <- read_csv(here('data', 'acs_count_mixed.csv')) %>%
#   group_by(state, year, same_sex, mixed) %>%
#   summarize(n = sum(n),
#             n_unweighted = sum(n_unweighted),
#             n_ar = sum(n_ar),
#             n_unweighted_ar = sum(n_unweighted_ar)) %>%
#   mutate(post_2013 = year > 2013,
#          group_fe = paste(state, same_sex, mixed, sep = '_'))

acs_prop <- read.csv(here('data', 'acs_prop_yrimmig_policy.csv')) %>%
  mutate(post_2013 = (yrimmig > 2013))
acs_prop_state <- read.csv(here('data', 'acs_dyad_policy.csv'))

lgb_origin_index <- read_csv(here('data', 'lgb_origin_index.csv'))
acs_ind <- read_rds(here('data', 'acs_couple_policy.rds'))  %>%
  mutate(post_2013 = year > 2013
         # mixed =  mixed_citizenship = (citizen_main == 'Not a citizen' & citizen_partner != 'Not a citizen') | 
         #   (citizen_main != 'Not a citizen' & citizen_partner == 'Not a citizen')
         ) %>%
  left_join(lgb_origin_index,  by = c('yrimmig' = 'year', 'bpldid' = 'Code')) 

acs_count_mixed_bpld <- read_csv(here('data', 'acs_count_mixed.csv'))  %>%
  left_join(
    acs_ind %>%
      group_by(year, state, bpld, same_sex) %>% 
      summarize(across(c(equal_age:propaganda, origin_score), mean, na.rm = T))
      #summarize(origin_score = mean(origin_score))
    )

acs_ind_survey <- acs_ind %>% as_survey_design(ids = cluster, weights = perwt)
```

```{r intro-stats}
total_same_mixed <- acs_count_mixed_bpld %>%
    filter(mixed == T & same_sex == T) %>%
    group_by(year) %>%
    summarize(n = sum(n)) %>%
  pull(n)

total_dif_mixed <- acs_count_mixed_bpld %>%
    filter(mixed == T & same_sex == F) %>%
    group_by(year) %>%
    summarize(n = sum(n)) %>%
  pull(n)
```

# Introduction
## Introduction
- In 2013, the U.S. Supreme Court struck down the Defense of Marriage Act
    + Allowing same-sex couples to apply for spousal/marriage visas
- From 2013 to 2019, 156% increase in mixed-citizenship same-sex couples, compared to 8% for different-sex
- But from where do these couples come? 
    + Do LGB policy environments at country of origin contribute toward the migration of immigrants in same-sex couples into the U.S.?


## Introduction
```{r total-pop}

bind_rows(
  acs_count_mixed_bpld %>%
    mutate(cat = case_when(origin_score > 3 ~ 'Progressive',
                           origin_score < 0 ~ 'Regressive')) %>%
    filter(!is.na(cat) & mixed == T & same_sex == T) %>%
    group_by(year, cat) %>%
    summarize(n = sum(n)),
   acs_count_mixed_bpld %>%
    filter(mixed == T & same_sex == T) %>%
    group_by(year) %>%
    summarize(n = sum(n)) %>%
    mutate(cat = 'Full sample')
) %>% 
  ggplot(aes(x = year, y = n/1000, color = cat, linetype = cat)) +
  geom_line() +
  geom_vline(xintercept = 2013, linetype = 2) +
  labs(x = 'Year', y = 'n (thousands)')
```

# Background
## How Country-Origin LGB Policies Influence Migration

- Policies both influence and are constituted by cultural environments  
- Progressive LGB policies (e.g., marriage equality), make desiring same-sex union possible
- Regressive policies may: limit aspirations for same-sex union, public disclosure of such unions, and, ultimately, undermine credibility in visa applications.
- Regressive policies may also limit access to requisite resources to migrate


## Our Contribution

- Migration theory typically emphasizes economic and network theories  
- We aim to demonstrate how culture and identity interact with policy to shape migratory pathways
- Continue advancing research on queer migration; Recognizing populations besides refugees and asylum-seekers


# Data and Method
## Data
- all couples in the 2008 to 2019 American Community Survey (ACS)
    + 18 to 64 at time of survey
    + for immigrants, only those who immigrated at age 18 or older

- Individuals in same-sex couples are identified as respondents who live with a same-sex married or unmarried partner
    + necessarily excludes single and non-cohabiting LGB individuals


## Data
- Data structure: state-year-group counts ($n$ = 2448)
  + groups: same- or different-sex $\times$ same- or mixed-citizenship
  + e.g. number of people in same-sex mixed-citizenship couples in Arizona in 2012
- Explanatory variable: country of origin LGB policy index (sum of 16 policies) (Velasco 2020)
  + ranges from -3 to 10, with a mean of 1.7
  + progressive policies such civil unions, marriage equality, and hate-crime protection
  + regressive policies such as anti-propaganda laws

## Sample sizes
```{r data-tab}
options(scipen=999, digits=2)
acs_count_mixed %>%
  mutate(Composition = if_else(same_sex == T, 'Same sex', 'Different sex'),
         Citizenship = if_else(mixed == T, 'Mixed citizenship', 'Same citizenship')) %>%
  group_by(Composition, Citizenship) %>%
  summarize(`n (unweighted)` = sum(n_unweighted),
            `n (weighted)` = sum(n)) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ",")) %>%
   kable_styling(position = "center")
  # flextable() %>%
  # flextable::autofit() 
  # flextable::set_caption('Unweighted and weighted sample sizes from American Community Survey (ACS) data, 2008-2019')
```



## Method
- Difference-in-differences-in-differences (DDD) using quasi-Poisson regression
  + Following Redpath (2022), model incidence of mixed-citizenship same-sex couples after 2013 relative to those before 2013 and same-citizenship or different-sex couples
- Stratify by policy environment
  + include only country-state-year mixed-citizenship groups with progressive (>3) or regressive (<0) policy scores
  + include only groups with specific policies

## Method
$$\begin{aligned}
y_{gst} = &\exp[\beta_0 + \beta_1 post_t + \beta_2 (M_g \times post_t) + \beta_3 (S_g \times post_t)  \\
&+ \beta_4 ( M_g \times S_g \times post_t) + \alpha_{gs} + \gamma_{t} + \epsilon_{gst}]
\end{aligned}$$  

- Where
  + $y_{gst}$: count of individuals in group $g$ in state $s$ in survey year $t$  
  + $post_t$: indicator  for $t > 2013$  
  + $M_g$: indicator for group $g$ being mixed-citizenship  
  + $S_g$: indicator for group $g$ being same-sex  
  + $\alpha_{gs}$: group-state fixed effects  
  + $\gamma_t$: survey year fixed effects  
  + $\epsilon_{gst}$ is an error term such that $\mathbb E(\epsilon_{gst}) = 0$ 

- Coefficient of interest is $\beta_4$: the three-way interaction between indicators for post-2013, same-sex, and mixed-citizenship. the incidence ratio $\exp(\beta_4)$ estimates the relative increase in mixed-citizenship same-sex couples after 2013, relative to other couples.


# Results
## Results
\begin{center}
  \huge{\textcolor{uclablue}{Results}}
\end{center}
  

## Main Effects
```{r mod-tab-setup, include = F}
mod_mixed <- acs_count_mixed %>%
  glm(n ~ I(post_2013*same_sex*mixed) + I(post_2013*same_sex) + I(post_2013*mixed) + 
        post_2013 + group_fe + as.factor(year), 
                        data = ., 
                        family = 'quasipoisson') %>%
  coeftest(., vcov = vcovCL(., cluster = ~group_fe)) 

# acs_count_mixed %>%
#   glm(n ~ post_2013*same_sex*mixed + group_fe + as.factor(year),
#                         data = .,
#                         family = 'quasipoisson') %>%
# coeftest(., vcov = vcovCL(., cluster = ~group_fe)) %>%
#   tidy() %>%
#   filter(!str_detect(term, 'group'), !str_detect(term, 'year'))

count_origin_high <- # count_func('origin_score > 3 | is.na(origin_score)') 
  acs_count_mixed_bpld %>%
  filter(origin_score > 3 | is.na(origin_score)) %>%
  group_by(state, year, same_sex, mixed, post_2013, group_fe) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  complete(state, year, same_sex, mixed) %>%
  mutate(across(n, function(x) ifelse(is.na(x), 0, x)),
         post_2013 = year > 2013,
         group_fe = paste(state, same_sex, mixed, sep = '_'))

# acs_count_base %>%
#   select(state, year, mixed_citizenship, same_sex, bpld, bpldnew, yrimmig, yrimmigmod, origin_score) %>%
#   sample_n(10)



# count_origin_high <- acs_count_base %>%
#     rename(mixed = mixed_citizenship) %>%
#     filter(origin_score > 3 | mixed == F) %>%
#     # filter(origin_score > 3 | is.na(origin_score)) %>%
#     group_by(state, year, same_sex, mixed) %>%
#     summarize(n = sum(perwt), 
#               n_unweighted = n(),
#               n_ar = sum(perwt[qcitizen != 'Allocated' & qeduc != 'Consistency edit' & allocated == F]), 
#               n_unweighted_ar = sum(qcitizen != 'Allocated' & qeduc != 'Consistency edit' & allocated == F)) %>%
#     ungroup() %>%
#     complete(state, year, same_sex, mixed) %>%
#     mutate(across(n:n_unweighted_ar, function(x) ifelse(is.na(x), 0, x))) %>%
#     mutate(post_2013 = year > 2013,
#            group_fe = paste(state, same_sex, mixed, sep = '_')) 


mod_os_high <- glm(n ~  I(post_2013*same_sex*mixed) + 
                        I(post_2013*same_sex) + 
                        I(post_2013*mixed) + 
                        post_2013 + 
                        group_fe + as.factor(year), 
                        data = count_origin_high, 
                        family = 'quasipoisson') %>%
  coeftest(., vcov = vcovCL(., cluster = ~group_fe))

count_origin_low <- # count_func('origin_score < 0 |  is.na(origin_score)')
  acs_count_mixed_bpld %>%
  filter(origin_score < 0 | is.na(origin_score)) %>%
  group_by(state, year, same_sex, mixed, post_2013, group_fe) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  complete(state, year, same_sex, mixed) %>%
  mutate(across(n, function(x) ifelse(is.na(x), 0, x)),
         post_2013 = year > 2013,
         group_fe = paste(state, same_sex, mixed, sep = '_'))

mod_os_low <- glm(n ~  I(post_2013*same_sex*mixed) + 
                        I(post_2013*same_sex) + 
                        I(post_2013*mixed) + 
                        post_2013 + 
                        group_fe + as.factor(year), 
                        data = count_origin_low, 
                        family = 'quasipoisson') %>%
  coeftest(., vcov = vcovCL(., cluster = ~group_fe)) 
```

```{r mod-tab, eval = T}
bind_rows(
  mutate(tidy(mod_mixed, conf.int = T), name = 'Full sample'),
  mutate(tidy(mod_os_high, conf.int = T), name = 'Progressive (>3)'),
  mutate(tidy(mod_os_low, conf.int = T), name = 'Regressive (<0)')) %>%
  filter(term %in% c('I(post_2013 * same_sex * mixed)', 'I(post_2013 * same_sex)',
                'I(post_2013 * mixed)', 'post_2013TRUE')) %>%
  mutate(term = recode(term, 
                'I(post_2013 * same_sex * mixed)' = 'Post-2013 × Same-sex × Mixed-citizenship' ,
                'I(post_2013 * same_sex)' ='Post-2013 × Same-sex',
                'I(post_2013 * mixed)' = 'Post-2013 × Mixed-citizenship',
                'post_2013TRUE' = 'Post-2013')) %>%
  ggplot(aes(x = term, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip() + 
  facet_wrap(~name, ncol = 3) +
  labs(x = '', y = '')

```

## Dynamic Models
```{r lag-plot-1}
base_plot <- acs_count_mixed %>%
  mutate(year = floor(year/2)*2) %>%
  group_by(state, year, same_sex, mixed, group_fe) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  mutate(year = factor(floor(year/2)*2, 
            levels = c('2018', '2016', '2014', '2012', '2010', '2008'))) %>%
  glm(n ~ as.factor(year):same_sex:mixed + 
                        as.factor(year):same_sex + 
                        as.factor(year):mixed + 
                        as.factor(year) + group_fe,
                        data = ., 
                        family = 'quasipoisson') %>%
  coeftest(., vcov = vcovCL(., cluster = ~group_fe)) %>%
  tidy(conf.int = T) %>% 
  filter(str_detect(term, ':same_sexTRUE:mixedTRUE')) %>%
  mutate(term = str_extract(term, '\\d\\d\\d\\d'),
         year = as.numeric(term)) %>%
  ggplot(aes(x = year, y = estimate)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 2013, linetype = 2) +
    theme(legend.position = "none") +
    labs(x = 'Year', y = 'Estimate', title = 'Full sample')



extra_pois_mixed_high <- glm(n ~ as.factor(year):same_sex:mixed + 
                        as.factor(year):same_sex + 
                        as.factor(year):mixed + 
                        as.factor(year) + group_fe,
                        data = mutate(count_origin_high, 
                                      year = factor(floor(year/2)*2, 
                                                    levels = c('2018', '2016', '2014', '2012', '2010', '2008'))), 
                        family = 'quasipoisson')

extra_pois_mixed_low <- glm(n ~ as.factor(year):same_sex:mixed + 
                        as.factor(year):same_sex + 
                        as.factor(year):mixed + 
                        as.factor(year) + group_fe,
                        data = mutate(count_origin_low, 
                                      year = factor(floor(year/2)*2, 
                                                    levels = c('2018', '2016', '2014', '2012', '2010', '2008'))), 
                        family = 'quasipoisson')

strat_plot <- bind_rows(
  extra_pois_mixed_high %>%
    coeftest(., vcov = vcovCL(., cluster = ~group_fe)) %>%
    tidy(conf.int = T) %>% 
    filter(str_detect(term, ':same_sexTRUE:mixedTRUE')) %>%
    mutate(year = as.numeric(str_extract(term, '\\d\\d\\d\\d')),
           mod = 'Score > 3'),
  extra_pois_mixed_low %>%
    coeftest(., vcov = vcovCL(., cluster = ~group_fe)) %>%
    tidy(conf.int = T) %>%  
    filter(str_detect(term, ':same_sexTRUE:mixedTRUE')) %>%
    mutate(year = as.numeric(str_extract(term, '\\d\\d\\d\\d')),
           mod = 'Score < 0')) %>%
  ggplot(aes(x = year, y = estimate, color = mod)) + #shape = mod)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high), position=position_dodge(width=.5)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 2013, linetype = 2)  + 
  theme(legend.position=c(.18,.9)) +
  labs(x = 'Year', y = 'Estimate', title = 'Stratified')

p_combined <- base_plot + strat_plot
p_ranges_y <- c(ggplot_build(p_combined[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(p_combined[[2]])$layout$panel_scales_y[[1]]$range$range)
base_plot & 
  ylim(min(p_ranges_y), max(p_ranges_y))
```

## Dynamic Models
```{r lag-plot-2}
strat_plot & 
  ylim(min(p_ranges_y), max(p_ranges_y))
```


## Specific Policies: Progressive
```{r policies-plot-1}
policy_list <- list()
variables <- names(subset(acs_count_mixed_bpld, select = equal_age:propaganda))

for(var_name in variables){
  dataset_policy <- acs_count_mixed_bpld %>%
    filter(!!sym(var_name) > 0
           | is.na(!!sym(var_name))) %>%
    group_by(state, year, same_sex, mixed, post_2013, group_fe) %>%
    summarize(n = sum(n)) %>%
    ungroup() %>%
    complete(state, year, same_sex, mixed) %>%
    mutate(across(n, function(x) ifelse(is.na(x), 0, x)),
           post_2013 = year > 2013,
           group_fe = paste(state, same_sex, mixed, sep = '_'))
  
  # dataset_policy <- count_func(paste0(var_name, ' > 0 | is.na(', var_name , ')'))
  
  policy_list[[var_name]] <- glm(n ~  I(post_2013*same_sex*mixed) + 
                        I(post_2013*same_sex) + 
                        I(post_2013*mixed) + 
                        post_2013 + 
                        group_fe + as.factor(year), 
                        data = dataset_policy, 
                        family = 'quasipoisson') %>%
  coeftest(., vcov = vcovCL(., cluster = ~group_fe))
}

prog_policies <- c('equal_age', 'constitution', 'employment_discrim', 'hate_crimes',
                   'incite_hate', 'joint_adoption', 'lgb_military', 'marriage_equality',
                   'samesex_legal', 'civil_unions', 'conversion_therapies')
rep_policies <- c('unequal_age', 'death_penalty', 'lgb_military_ban', 
                  'marriage_ban', 'propaganda')

policy_df <- lapply(policy_list, function(x){
  tidy(x)[2, 2:3]}
  ) %>%
  bind_rows() %>%
  mutate(policy = names(policy_list),
         type = if_else(policy %in% prog_policies, 'Progressive', 'Regressive'),
         policy_name = recode(policy, 
                         equal_age = 'Equal age of consent', 
                         unequal_age = 'Unequal age of consent',
                         constitution = 'Constitutional protection',
                         conversion_therapies = 'Conversion therapy ban',
                         death_penalty = 'Death penalty',
                         employment_discrim = 'Employment protection',
                         hate_crimes = 'Hate-crime protection',
                         incite_hate = 'Incitement to hate banned',
                         joint_adoption = 'Joint adoption',
                         lgb_military = 'LGB military',
                         lgb_military_ban = 'LGB military ban',
                         marriage_equality = 'Marriage equality',
                         marriage_ban = 'Marriage ban',
                         samesex_legal = 'Same-sex acts legal',
                         civil_unions = 'Civil unions',
                         propaganda = 'Anti-propaganda laws'),
         lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error)

policy_df %>% 
  filter(type == 'Progressive') %>%
  # arrange(desc(estimate)) %>%
  # head(10) %>%
  ggplot(aes(reorder(policy_name, estimate), y = estimate)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_pointrange(aes(ymin = lower, ymax = upper)) +
  coord_flip() +
  # facet_wrap(~type, scales = 'free_y') +
  labs(x = '', y = '') +
  ylim(c(-.4, 0.75))
```

## Specific Policies: Regressive

```{r policies-plot-2}
policy_df %>% 
  filter(type == 'Regressive') %>%
  # arrange(desc(estimate)) %>%
  # head(10) %>%
  ggplot(aes(reorder(policy_name, estimate), y = estimate)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_pointrange(aes(ymin = lower, ymax = upper)) +
  coord_flip() +
  # facet_wrap(~type, scales = 'free_y') +
  labs(x = '', y = '') +
  ylim(c(-.4, 0.75))
```


# Discussion

- Rise in mixed-citizenship, same sex couples after the 2013 end of DOMA, but only for immigrants from progressive countries
- stronger effects for civil unions and marriage equality
- culture, identity, and policy can affect migration and unions
- broadening focus of queer migration studies to immigrants from progressive countries


## Next steps
- Adjust for possible state- and group-level confounders
- Stratify by state LGB policy environment
- Incorporate 2020 data


## 
\begin{center}
  \huge{\textcolor{uclablue}{Thank You}}
\end{center}

- Nathan I. Hoffmann ([nathanihoff\@ucla.edu](mailto:nathanihoff@ucla.edu))
- Kristopher Velasco ([kvelasco\@princeton.edu](mailto:kvelasco\@princeton.edu))  


``` {=latex}
\end{frame}
\appendix
\begin{frame}<0| handout:0>
```


# Supplementary Material
## Country-Level Descriptives
```{r desc-country}
distance <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(distw, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'C. Distance to Country of Origin (km)')

unemp <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = -survey_mean(unemp_dif, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'B. Unemployment Rate Difference with U.S.')

prop_stock <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(stock_prop, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'D. Proportion U.S. Immigrant Stock')

wage_dif <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = -survey_mean(wage_dif, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'A. GDP Difference with U.S.')

vdem <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(vdem, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'E. Democracy Score (V-Dem)')


bind_rows(wage_dif, vdem, prop_stock, distance) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = var, linetype = same_sex,
             color = same_sex, fill = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = var_low, ymax = var_upp, group = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var_name, scales = "free", ncol = 2) +
  ylab('') +
  theme( # legend.justification=c(1,0),
  #     legend.position=c(.22,.82),
  #     legend.background = element_rect(fill = alpha("white", 0.5))
      legend.position="bottom")
 
```

## Individual Descriptives
```{r desc-ind}
income <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(inctot, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'C. Annual Gross Income (Thousands)')

hs <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex'),
         hs = if_else(educ %in% c('HS', 'some col', 'college'), 1, 0)) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(hs, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'A. High School Completion')

college <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex'),
         college = if_else(educ == 'college', 1, 0)) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(college, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'B. College Completion')

# state_income <- acs_ind_survey %>%
#   mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
#   group_by(year, same_sex) %>%
#   summarize(var = survey_mean(state_income, vartype = "ci", na.rm = T)) %>%
#   mutate(var_name = 'U.S. State of Residence Per Capita Income')

hwsei <- acs_ind_survey %>%
  mutate(same_sex = if_else(same_sex == T, 'Same-Sex', 'Different-Sex')) %>%
  group_by(year, same_sex) %>%
  summarize(var = survey_mean(hwsei, vartype = "ci", na.rm = T)) %>%
  mutate(var_name = 'D. Hauser and Warren Socioeconomic Index')


bind_rows(income, hs, college, hwsei) %>%
  rename(Year = year) %>%
  ggplot(aes(x = Year, y = var, linetype = same_sex, color = same_sex, fill = same_sex)) +
  geom_line() +
  geom_ribbon(aes(ymin = var_low, ymax = var_upp, 
                  group = same_sex, col = NULL), alpha = .2) +
  facet_wrap(~var_name, scales = "free", ncol = 2) +
  ylab('') +
  theme( # legend.justification=c(1,0),
  #     legend.position=c(.22,.82),
  #     legend.background = element_rect(fill = alpha("white", 0.5))
      legend.position="bottom")
```
